\documentclass[11pt]{article}

% --- Encoding & fonts ---
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{microtype}

% --- Math ---
\usepackage{amsmath,amssymb,amsfonts,amsthm}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}
\newtheorem{remark}{Remark}
\newtheorem{assumption}{Assumption}

% --- Layout ---
\usepackage[margin=1in]{geometry}
\usepackage[section]{placeins}
\setlength{\emergencystretch}{5em}

% Force \FloatBarrier before every \subsection and \subsubsection so floats
% cannot cross into the next (sub)subsection, eliminating orphan headings
% followed by tall whitespace (the \texttt{[section]} option of
% \texttt{placeins} only barriers at \section).
\let\origsubsection\subsection
\renewcommand{\subsection}{\FloatBarrier\origsubsection}
\let\origsubsubsection\subsubsection
\renewcommand{\subsubsection}{\FloatBarrier\origsubsubsection}

% --- Floats, tables, figures ---
\usepackage{graphicx}
\usepackage{float}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{caption}
\usepackage{subcaption}

% --- Aggressive float packing (relaxed defaults to avoid sparse pages) ---
\renewcommand{\topfraction}{0.95}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\textfraction}{0.05}
\renewcommand{\floatpagefraction}{0.75}
\renewcommand{\dbltopfraction}{0.95}
\renewcommand{\dblfloatpagefraction}{0.75}
\setcounter{topnumber}{3}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{5}

% --- Algorithms, lists, color ---
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{enumitem}
\usepackage{xcolor}

% breakablealgorithm: a non-floating algorithm environment that can split
% across pages. Use in place of \begin{algorithm} for any algorithm body
% taller than a single page; \caption / \label / counter increments still
% work as in the floating version.
\makeatletter
\newenvironment{breakablealgorithm}
  {\begin{center}
    \refstepcounter{algorithm}%
    \hrule height.8pt depth0pt \kern2pt%
    \renewcommand{\caption}[2][\relax]{%
      {\raggedright\textbf{\ALG@name~\thealgorithm} ##2\par}%
      \ifx\relax##1\relax%
        \addcontentsline{loa}{algorithm}{\protect\numberline{\thealgorithm}##2}%
      \else%
        \addcontentsline{loa}{algorithm}{\protect\numberline{\thealgorithm}##1}%
      \fi%
      \kern2pt\hrule\kern2pt%
    }%
  }{\kern2pt\hrule\relax%
   \end{center}%
  }
\makeatother

% --- TikZ ---
\usepackage{tikz}
\usetikzlibrary{shapes.geometric,arrows.meta,positioning,fit,calc,decorations.pathreplacing}

% --- Bibliography ---
\usepackage[numbers,sort&compress]{natbib}

% --- Links (load last) ---
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    citecolor=blue,
    urlcolor=blue,
    pdftitle={Continuous Hidden Markov Models for Equity Returns: Heavy-Tail Emission Families and Regime-Conditional Value-at-Risk},
    pdfauthor={Abdulrahman Alswaidan, Cade Jin, Jeffrey D. Varner},
    pdfsubject={CHMM synthetic-data generator for daily US-equity returns},
    pdfkeywords={CHMM, Baum-Welch, Student-t emissions, semi-Markov, stylized facts, copula, Value-at-Risk, Kupiec, synthetic data}
}
\makeatletter
\g@addto@macro\UrlBreaks{\do\-\do\_}
\makeatother

% --- Caption style (V10/V11, 2026-04-22) ---
% Top-of-table labels, bold table/figure tags, small font for captions and subcaptions.
\captionsetup{labelfont=bf, labelsep=period, font=small, justification=justified}
\captionsetup[sub]{labelfont=bf, font=small, labelformat=parens, labelsep=space}

% --- Custom commands ---
\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Prob}{\mathbb{P}}

% --- Project-level numeric macros (V10/V11 bundle, 2026-04-22) ---
% Consistent formatting for recurring numeric constructs.
% Use in new prose and tables; legacy cells may be migrated gradually (see V16
% siunitx migration in plan-equity-paper.md for the bulk pass).
\newcommand{\pct}[1]{\ensuremath{#1\%}}
\newcommand{\Kdisc}[1]{\ensuremath{K_{\text{disc}} = #1}}
\newcommand{\LRuc}[1]{\ensuremath{\text{LR}_{\text{uc}} = #1}}
\newcommand{\NPaths}[1]{\ensuremath{N_{\text{paths}} = #1}}
\newcommand{\TIS}[1]{\ensuremath{T_{\text{IS}} = #1}}
\newcommand{\TOoS}[1]{\ensuremath{T_{\text{OoS}} = #1}}

\title{Continuous Hidden Markov Models for Equity Returns:\\ Heavy-Tail Emission Families\\ and Regime-Conditional Value-at-Risk}

\author{
    Abdulrahman Alswaidan\thanks{Cornell University, Robert Frederick Smith School of Chemical and Biomolecular Engineering, Ithaca, NY, USA. Email: \texttt{aa2725@cornell.edu}.}
    \and
    Cade Jin\thanks{Cornell Ann S.\ Bowers College of Computing and Information Science, Cornell University, Ithaca, NY, USA. Email: \texttt{cj383@cornell.edu}.}
    \and
    Jeffrey D.\ Varner\thanks{Cornell University, Robert Frederick Smith School of Chemical and Biomolecular Engineering, Ithaca, NY, USA. Email: \texttt{jdv27@cornell.edu}.}
}

\date{\today}

\begin{document}

\maketitle

% ========================================================================================= %
\begin{abstract}
We present a continuous hidden Markov model (CHMM) framework for daily US-equity returns. The textbook bilinear identity for the absolute-return autocorrelation function recasts the empirical Ryd\'{e}n et al.\ low-$K$ failure as an effective-rank statement on the centred transition matrix $\mathbf T - \mathbf 1\bar{\boldsymbol\pi}^\top$. An empirical diagnostic on the fitted estimator $\hat{\mathbf T}$ shows the implied $K - 1$ rank bound is non-binding at the cross-ticker median for $K \ge 3$, with the bound still doing visible work on the left tail of the distribution (minimum dominant-mode share $32.6\%$ on NEM).

A unified ECM scaffold supports four emission families (Gaussian, Student-$t$, Laplace, Generalised Error Distribution) with shared forward-backward recursions and quantile-based initialisation; the M-step is the only architectural difference. The empirical study covers SPY 2014 to 2026, a sector-balanced 30-ticker panel, a CRSP cross-decade transfer (1994 to 2004 in-sample, 2004 to 2006 out-of-sample), and a six-asset US-equity copula extension.

Headline operating result: a regime-conditional Value-at-Risk that propagates the one-step-ahead state forecast through the predictive mixture passes the Christoffersen joint conditional-coverage test cleanly at $\alpha = 0.05$ across emission families on the headline out-of-sample window and on $19/24$ rows of a six-fold rolling-origin walk-forward; rejections concentrate on out-of-distribution stress folds. At $\alpha = 0.01$ the Engle-Manganelli dynamic-quantile backstop rejects $K = 18$ on the headline window, so the strict-tail Christoffersen pass is power-driven.

Positioning is honest. The i.i.d.\ bootstrap and a maximum-likelihood hidden semi-Markov model both exceed the CHMM on raw single-window out-of-sample Kolmogorov-Smirnov. The CHMM value-add is the structural use cases that those alternatives cannot serve (regime-conditional VaR, multi-asset copula composition, parametric synthetic-data privacy), together with a unified emission-family ablation. The single-window OoS sits at the upper tail of a six-fold rolling-origin walk-forward whose median CHMM-N OoS KS is $62.1\%$ at $K^\star = 3$, the operationally informative summary. Scope is daily US equities under stationary OoS; a non-equity stress test on GLD collapses under static fitting, and periodic refit is the deployment recipe.
\end{abstract}

\noindent\textbf{Keywords:} Continuous Hidden Markov Model, Baum-Welch, Student-$t$ Emissions, Stylized Facts, Volatility Clustering, Value-at-Risk, Synthetic-Data Generator.

% ========================================================================================= %
\section{Introduction}
\label{sec:introduction}
\input{sections/introduction}

\section{Related Work}
\label{sec:related}
\input{sections/related_work}

\section{Methods}
\label{sec:methods}
\input{sections/model}
\input{sections/estimation}

\section{Spectral Mechanism}
\label{sec:spectral}
\input{sections/theory}

\section{Empirical Study}
\label{sec:results}
\input{sections/results}
\input{sections/var_backtest}

\section{Discussion}
\label{sec:discussion}
\input{sections/discussion}

\section{Conclusion}
\label{sec:conclusion}
\input{sections/conclusion}

% ========================================================================================= %
\bibliographystyle{unsrtnat}
\bibliography{references}

% ========================================================================================= %
\appendix

\section{Supplementary Material}
\label{sec:supplementary}
\input{sections/supplementary}

\end{document}
