\subsubsection{OoS KS Test-Power Calibration}
\label{sec:ks_power}

We calibrated the two-sample KS test power to anchor the OoS KS pass rates of the main generator comparison (Table~\ref{tab:ks_power}).
For each of $1{,}000$ Monte Carlo replications (seed = 20260420), a simulated series of length $T_{\text{OoS}} = 572$ is drawn from each of three reference generators and tested against the observed OoS series using the same $\alpha = 0.05$ threshold as the main-text pass-rate metric.

\begin{table}[!ht]
\centering
\small
\caption{OoS KS pass-rate calibration (1{,}000 replications, seed = 20260420).}
\label{tab:ks_power}
\begin{tabular}{l c}
\toprule
Reference generator & Pass rate (\%) \\
\midrule
I.i.d.\ resamples of $R_{\text{OoS}}$ at length $T_{\text{OoS}} = 572$ (known-correct ceiling) & 100.0 \\
I.i.d.\ resamples of $R_{\text{IS}}$ at length $T_{\text{OoS}} = 572$ (nearly correct)          & 90.0  \\
Gaussian$(\hat\mu_{\text{IS}}, \hat\sigma_{\text{IS}})$ at length $T_{\text{IS}}$, tested against $R_{\text{IS}}$ (negative control) & 0.0 \\
\bottomrule
\end{tabular}
\end{table}

The $90.0\%$ ``nearly correct'' rate is the practical ceiling for the CHMM OoS KS pass rates reported in the main generator comparison ($63.6$ to $82.1\%$ across the CHMM rows of Table~\ref{tab:model_comparison}); the $0\%$ rate on the Gaussian misspecified generator confirms that the KS test retains ample power at IS length to reject clearly wrong generators.

\subsubsection{Block-Bootstrap KS Recalibration and Mean p-value Distribution}
\label{sec:ks_block_bootstrap}

Two complementary concerns can be raised about the main IS KS pass-rate metric. The asymptotic two-sample KS test assumes i.i.d.\ samples, but the simulated paths from CHMM and GARCH all carry volatility clustering, so the asymptotic null distribution may not apply. The binary ``pass/fail at $\alpha = 0.05$'' also reduces a continuous p-value to a one-bit summary. We computed two recalibrations addressing each concern (Table~\ref{tab:ks_block_bootstrap}).

To address the dependence concern, we generated the empirical null distribution of the two-sample KS statistic under a two-sample stationary-block-bootstrap null~\citep{politis1994stationary}: each of $B = 1{,}000$ replicates draws \emph{two independent} block resamples of the SPY IS series (mean block lengths $L \in \{5, 10, 20\}$) and computes the KS statistic between them, so the null replicate has the same structure as the actual observed-versus-simulated comparison: two independent length-$n$ series sharing the observed marginal and short-range dependence. The resulting $95\%$ critical values were $0.0433$, $0.0469$, and $0.0497$, all \emph{larger} than the asymptotic i.i.d.\ value $1.36 \sqrt{2/n_{\text{IS}}} = 0.0383$: accounting for serial dependence widens the null distribution of the two-sample KS statistic, so the dependence-aware test is \emph{less} strict than the asymptotic one, and the asymptotic pass rates reported in the main table are conservative in this respect. The threshold remains conditional on the empirical IS distribution (no generator refit per replicate), so all pass rates in this appendix are descriptive. For each generator we report the fraction of paths whose KS statistic falls below the block-bootstrap critical value.

To address the binarisation concern, we report the mean two-sample KS p-value across simulated paths plus its $5$/$25$/$50$/$75$/$95$ quantiles, giving a continuous picture of how distinguishable each generator's paths are from the observed IS series.

\begin{table}[t]
\centering
\small
\caption{\textbf{Block-bootstrap KS recalibration and p-value distribution}. All numbers under seed \texttt{20260422} with $N_{\text{paths}} = 500$; the CHMM rows are this diagnostic's own fits at $K = 18$ (CHMM-t is the unpenalised per-state-$\nu$ fit), so the pass rates are not the $K^\star = 3$ main-table values. ``asymp pass'' uses the metric definition of Table~\ref{tab:model_comparison} (two-sample KS p-value $\geq 0.05$), recomputed within this run. ``mean pv'' is the mean two-sample KS p-value across paths, with $q_5$ and $q_{95}$ as the lower / upper p-value quantiles. ``blkL\%'' is the fraction of paths whose KS statistic falls below the two-sample stationary-block-bootstrap $95\%$ critical value at mean block length $L$ (two independent IS resamples per null replicate). Critical values: $0.0433$ ($L = 5$), $0.0469$ ($L = 10$), $0.0497$ ($L = 20$); asymptotic i.i.d.\ value $0.0383$.}
\label{tab:ks_block_bootstrap}
\begin{tabular}{l c c c c c c c}
\toprule
Generator & asymp pass\% $\uparrow$ & mean pv $\uparrow$ & pv $q_{05}$ & pv $q_{95}$ & blk5\% $\uparrow$ & blk10\% $\uparrow$ & blk20\% $\uparrow$ \\
\midrule
i.i.d.\ bootstrap & $99.6$ & $0.820$ & $0.352$ & $0.999$ & $100.0$ & $100.0$ & $100.0$ \\
GARCH(1,1)        & $26.8$ & $0.048$ & $0.000$ & $0.227$ & $42.4$ & $54.4$ & $63.0$ \\
CHMM-N            & $94.8$ & $0.546$ & $0.047$ & $0.976$ & $98.4$ & $98.8$ & $99.2$ \\
CHMM-t            & $\mathbf{95.8}$ & $\mathbf{0.571}$ & $0.060$ & $0.987$ & $98.2$ & $99.4$ & $99.6$ \\
CHMM-L            & $94.4$ & $0.504$ & $0.047$ & $0.948$ & $98.0$ & $99.0$ & $99.6$ \\
\textbf{CHMM-GED} & $95.6$ & $0.547$ & $0.064$ & $0.968$ & $\mathbf{98.8}$ & $\mathbf{99.6}$ & $\mathbf{100.0}$ \\
\bottomrule
\end{tabular}
\end{table}

Under the dependence-aware null every CHMM variant passes on essentially all paths ($98$--$100\%$ across the three block lengths), the i.i.d.\ bootstrap passes on all, and GARCH(1,1) rises from $26.8\%$ asymptotic to $42$--$63\%$ block-aware while remaining clearly the weakest entry (Table~\ref{tab:ks_block_bootstrap}). The cross-generator ordering is unchanged from the asymptotic metric; the level shift reflects the wider dependence-aware null rather than any change in the fits.

\paragraph{OoS-anchored block-bootstrap recalibration.}
\label{sec:ks_block_bootstrap_oos}
The block-bootstrap recalibration of Table~\ref{tab:ks_block_bootstrap} is anchored on the IS series; we report here whether the same construction holds on the OoS window. We repeated the procedure (two independent block resamples per null replicate) re-anchored on $R_{\text{OoS}}$ ($n_{\text{OoS}} = 572$, $B = 1{,}000$, $L \in \{5, 10, 20\}$, $N_{\text{paths}} = 500$). The critical values were $0.0857$, $0.0909$, and $0.0927$ at $L = 5, 10, 20$ respectively, again above the asymptotic $0.0804$. The OoS-anchored panel is reported in Table~\ref{tab:ks_block_bootstrap_oos}; the source CSV is in the companion code repository.

\begin{table}[!ht]
\centering
\small
\caption{\textbf{OoS-anchored block-bootstrap KS recalibration.} Same construction as Table~\ref{tab:ks_block_bootstrap} (CHMM rows at $K = 18$, this run's own fits) but with the two-sample stationary-block-bootstrap null anchored on $R_{\text{OoS}}$ ($n = 572$). ``asymp pass'' is the main OoS metric (two-sample KS p-value $\ge 0.05$); ``blkL\%'' is the OoS block-bootstrap pass rate at mean block length $L$. The cross-generator ordering is preserved under the dependence-aware recalibration on both windows, with all pass rates shifted upward by the wider null.}
\label{tab:ks_block_bootstrap_oos}
\begin{tabular}{l c c c c c c}
\toprule
Generator & asymp pass\% $\uparrow$ & mean pv $\uparrow$ & pv $q_{05}$ & pv $q_{95}$ & blk5\% $\uparrow$ & blk10\% $\uparrow$ \\
\midrule
i.i.d.\ bootstrap & $90.4$ & $0.437$ & $0.030$ & $0.912$ & $94.2$ & $96.8$ \\
GARCH(1,1)        & $59.2$ & $0.164$ & $0.000$ & $0.645$ & $65.8$ & $75.6$ \\
CHMM-N            & $81.0$ & $0.318$ & $0.010$ & $0.876$ & $87.0$ & $91.4$ \\
\textbf{CHMM-t}   & $\mathbf{84.8}$ & $\mathbf{0.355}$ & $0.018$ & $\mathbf{0.911}$ & $\mathbf{90.8}$ & $\mathbf{94.0}$ \\
CHMM-L            & $77.4$ & $0.280$ & $0.008$ & $0.791$ & $84.2$ & $90.0$ \\
CHMM-GED          & $81.0$ & $0.327$ & $0.010$ & $0.837$ & $86.4$ & $90.8$ \\
\bottomrule
\end{tabular}
\end{table}

The cross-generator OoS ordering matched the IS ordering (bootstrap $>$ CHMM-t $>$ CHMM-N/GED $>$ CHMM-L $\gg$ GARCH), and on both windows the dependence-aware pass rates sit above the asymptotic ones, so the asymptotic metric's cross-generator ranking is the robust comparison while its absolute levels are conservative.

\paragraph{Default-state-count summary at $L = 20$.}
The OoS asymptotic pass rate alongside the OoS block-bootstrap pass rate at $L = 20$ (two-sample block-bootstrap $95\%$ critical value $0.0927$ vs.\ asymptotic $0.0804$) at the default $K^\star = 3$ is reported in Table~\ref{tab:ks_block_body}. Under the dependence-aware null every $K^\star = 3$ family's pass rate rises by $10$--$19$pp relative to the asymptotic value; the ranking across families is unchanged.

\begin{table}[t]
\centering
\small
\caption{\textbf{Block-aware OoS KS recalibration at $L = 20$} (two-sample stationary-block-bootstrap null of \citealp{politis1994stationary}, two independent OoS resamples per replicate, $B = 1{,}000$, $500$ OoS-length simulated paths per generator, seed $20260422$). The asymp column recomputes the Table~\ref{tab:model_comparison} OoS metric at $500$ paths under this table's seed and RNG stream, so its values differ from the $1{,}000$-path main-table column by path noise ($\le 3$pp); the block-bootstrap $95\%$ null KS critical value at $L = 20$ is $0.0927$ vs the asymptotic $0.0804$. State count $K^\star = 3$.}
\label{tab:ks_block_body}
\begin{tabular}{l c c}
\toprule
Model & OoS KS asymp.\ (\%) & OoS KS block $L = 20$ (\%) \\
\midrule
Bootstrap                                & $90.4$ & $97.4$ \\
GARCH(1,1)                               & $59.2$ & $75.8$ \\
CHMM-N                                   & $79.8$ & $90.8$ \\
CHMM-t pen.\ ($\lambda = 20$)            & $\mathbf{82.4}$ & $\mathbf{92.8}$ \\
CHMM-L                                   & $64.4$ & $83.6$ \\
CHMM-GED                                 & $79.2$ & $90.2$ \\
\bottomrule
\end{tabular}
\end{table}

\subsubsection{QuantGAN Deep-Generative Baseline}
\label{sec:quantgan_supp}

The QuantGAN deep-generative baseline is a repo-native approximation to the convolutional WGAN of \citet{wiese2020quantgan}: a 1D convolutional generator and critic (each three layers, channels $32$, kernel size $5$; latent dim $L = 8$, leaky-ReLU on the critic), Wasserstein loss with critic weight clipping at $\pm 0.01$~\citep{arjovsky2017wasserstein}, Adam at $\eta = 10^{-4}$, batch $64$, $15$ epochs $\times\,120$ steps with four critic updates per generator update, trained on rolling windows of length $W = 64$ on standardised SPY growth rates and stitched end-to-end at synthesis. The implementation is Julia with \texttt{Flux.jl}~\cite{innes2018flux}; seed $20260422$. In the expanded panel (Table~\ref{tab:extended_baselines}) this configuration reached $0\%$ IS / OoS KS and IS excess kurtosis $2.1$ vs.\ observed $7.68$; the QuantGAN row in the main generator comparison (Table~\ref{tab:model_comparison}) is a separate temporal-convolutional run, also at $0\%$ KS, with IS excess kurtosis $0.56$. Both fall far short of the observed heavy tail, consistent with documented GAN failure modes on volatility-clustering benchmarks~\citep{takahashi2019modeling, kwon2024can}; the runner is materially smaller than the reference seven-block TCN with Lambert-W pre-processing in \citet{wiese2020quantgan}, so the row is the panel's deep-generative negative control rather than a verdict on the QuantGAN literature.

\subsubsection{Expanded Baseline Panel: GARCH Family, Semi-Markov CHMM, and QuantGAN}
\label{sec:extended_baselines}

We added five conditional-volatility baselines, the three semi-Markov CHMM variants, and the QuantGAN deep-generative baseline to the single GARCH(1,1) Gaussian row of Table~\ref{tab:model_comparison}. The conditional-volatility models are fit by grid-initialised Nelder-Mead maximum likelihood on the SPY IS window ($T_{\text{IS}} = 2{,}516$); simulation follows the same recursion used in estimation.

\paragraph{Baselines fitted in this panel.}
The conditional-volatility rows comprise EGARCH(1,1) Gaussian~\citep{nelson1991conditional}, GJR-GARCH(1,1) \citep{glosten1993relation, zakoian1994threshold}, GARCH(1,1) with Student-$t$ innovations~\citep{bollerslev1987conditionally}, HAR-RV~\citep{corsi2009simple} on daily squared demeaned growth rates as the RV proxy, and Markov-switching GARCH(1,1) \citep{haas2004new} at $K \in \{2, 3, 4, 6\}$ with regime triples $(\omega_k, \alpha_k, \beta_k)$ plus softmax-normalised transition logits, fitted by grid-initialised Nelder-Mead and canonicalised by ascending unconditional variance. Higher-$K$ MS-GARCH did not close the gap to the CHMM family on KS (IS KS plateaued at $27$--$37\%$ across $K \in \{2, 3, 4, 6\}$ against CHMM-N's $94.1\%$ at $K = 18$); the binding constraint is the unimodal-innovation per-regime structure (each regime's emission is Gaussian rescaled by the regime's conditional variance), which adding regimes does not relax. The CHMM decouples this constraint by giving each state its own emission-mixture component. Implementation details for every conditional-volatility row are in the companion code repository.

\paragraph{MS-GARCH reference Bayesian re-run via \texttt{MSGARCH}.}
As a Bayesian counterpart to the in-house Nelder-Mead MS-GARCH fit, we re-ran the same Markov-switching GARCH(1,1) specification through the reference \texttt{MSGARCH} R package of \citet{ardia2019msgarch} (version $2.51$) at $K \in \{2, 3, 4\}$, driven from the Julia harness via \texttt{RCall.jl}. The fit is fully Bayesian (DEMC sampler with the package's default proper priors, $12{,}500$ MCMC draws, $2{,}500$ burn-in, thin $10$, single chain); all $1{,}000$ simulated paths are posterior-predictive (one path per retained posterior draw), so the simulated marginal integrates parameter uncertainty path-by-path. The reference Bayesian re-run reported lower KS pass rates than our in-house frequentist Nelder-Mead fit: $0.0\%$ ($K = 2$), $0.1\%$ ($K = 3$), $0.0\%$ ($K = 4$) IS, and $5.8\%$ ($K = 2$), $5.1\%$ ($K = 3$), $5.3\%$ ($K = 4$) OoS, against the in-house plateau of $27$--$37\%$ IS / $33$--$39\%$ OoS (Table~\ref{tab:extended_baselines}). The gap is methodological rather than estimator-quality: the in-house implementation generates all $1{,}000$ paths from one MLE point estimate, so the simulated marginal is tighter; the reference Bayesian implementation samples one set of parameters per path from the posterior, so the simulated marginal is wider and accordingly fewer paths fall within the asymptotic two-sample KS critical band. The posterior-mean log-likelihood at the in-sample data improved modestly from $K = 3$ to $K = 4$ but did not translate to a higher KS pass rate. The posterior-mean transition matrix was diagonally dominant at $K = 3$ and less so at $K = 4$, indicating the higher-$K$ posterior was searching for additional structure that the data do not strongly support. The reference Bayesian re-run is therefore consistent with the in-house frequentist plateau at the same conclusion under either estimator: vanilla Gaussian-innovation MS-GARCH at $K \le 4$ did not match the CHMM family on the main KS metric for this universe. Reproducibility: R $4.6.0$, MSGARCH $2.51$, all transitive R dependencies pinned via \texttt{renv} (\texttt{r\_msgarch/renv.lock} of the companion code repository), all seeds explicit; full per-$K$ parameter posteriors and fit logs are written to \texttt{results/msgarch\_reference/} of the companion code repository.

\paragraph{Semi-Markov CHMM (SM-CHMM, the explicit hidden semi-Markov model (HSMM) foil).}
The SM-CHMM is the conceptual foil of the present paper, in the spirit of the explicit-duration HSMM of \citet{yu2010hidden}: a hidden semi-Markov extension of the CHMM in which the geometric sojourn distribution of the Markov chain is replaced by a per-state explicit-duration family chosen by raw fitted log-likelihood from $\{$Pareto, negative binomial, geometric$\}$. The plug-in estimator takes the Viterbi state path on the IS series under the converged CHMM fit, then refits per-state AR(1) emissions and a sojourn family chosen by fitted log-likelihood. On SPY IS at $K = 18$ the selected sojourn family was Pareto for $17$ of the $18$ states (negative binomial for one) across all three emission families, indicating heavy-tailed sojourns that the geometric CHMM cannot capture by construction.

\medskip
\noindent\emph{Caveat on the SM-CHMM estimator.} This plug-in construction (single-shot Viterbi decoding under the converged CHMM fit, then AR(1)-residual + sojourn-family refit) is not maximum-likelihood for the underlying HSMM: it does not iterate the joint forward-backward over (state, duration) pairs in the spirit of \citet{yu2010hidden}. A full ML HSMM at $K = 18$ would refit the duration-augmented log-likelihood jointly. The plug-in's per-state variance estimates degenerate at $K = 18$, where several low-occupancy states collapse to near-zero-variance point masses, so we do not report its distributional pass-rates and retain only its VaR back-test behaviour in Table~\ref{tab:extended_baselines}, where the heavy-tailed sojourns sharpen tail calibration. We do not claim that a jointly estimated HSMM would do the same; the multistart local-EM explicit-duration fit of the main table (Appendix~\ref{sec:supp_hsmm_diagnostic}) covers SPY at $K \in \{3, 18\}$, and extending it across the ticker universe is left as a companion-paper direction. The low-state ML HSMM result of \citet{bulla2006stylized} at low $K$ on monthly returns is the closest published anchor, but its $K \in \{2, 3\}$ regime resolution and monthly aggregation do not directly transfer.

The IS / OoS distributional fidelity, excess kurtosis, ACF-MAE, and Kupiec breach diagnostics for the extended GARCH family, the QuantGAN deep-generative row, and the main CHMM panel, together with the SM-CHMM VaR back-test rows, are reported in Table~\ref{tab:extended_baselines}.

\begin{table}[t]
\centering
\small
\caption{\textbf{Extended baseline panel}. SPY IS / OoS, $T_{\text{IS}} = 2{,}516$, $T_{\text{OoS}} = 572$, $N_{\text{paths}} = 1{,}000$, $\alpha = 0.05$ for KS. Columns: IS / OoS KS pass rate (\%); IS / OoS AD pass rate (\%); simulated IS excess kurtosis; ACF-MAE on $|G_t|$ (volatility clustering) and on raw $G_t$ (linear autocorrelation, parallel column); pooled-archive Kupiec breach rate (\%) and unconditional likelihood-ratio statistic at the $1\%$ and $5\%$ tails. \emph{Convention note:} this panel's $|G_t|$ ACF-MAE is the \emph{pooled} population-curve estimand (simulated ACF averaged across paths before the MAE), whereas Table~\ref{tab:model_comparison} uses the per-path estimand (mean of per-path MAEs, which includes per-path sampling noise); the same canonical grid-initialised GARCH(1,1) fit scores $0.0309$ pooled and $0.0490$ per-path (one fit, two estimands), so values must not be compared across the two tables. The CHMM-t row is the unpenalised per-state-$\nu$ fit. The SM-CHMM rows report ``--'' for the KS, AD, excess kurtosis, and ACF columns because the plug-in estimator's per-state variances degenerate at $K = 18$, where several low-occupancy states collapse to near-zero-variance point masses; only their VaR back-test columns are retained. The QuantGAN raw-ACF-MAE is ``--'' pending a full re-run of that baseline script.}
\label{tab:extended_baselines}
\resizebox{\textwidth}{!}{%
\begin{tabular}{l cc cc c cc cc cc}
\toprule
& \multicolumn{2}{c}{KS (\%) $\uparrow$} & \multicolumn{2}{c}{AD (\%) $\uparrow$} & Kurt & \multicolumn{2}{c}{ACF-MAE} & \multicolumn{2}{c}{$1\%$ tail} & \multicolumn{2}{c}{$5\%$ tail} \\
\cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){7-8} \cmidrule(lr){9-10} \cmidrule(lr){11-12}
Model & IS & OoS & IS & OoS & IS & $|G_t|$ & $G_t$ & br\% & $\text{LR}_{\text{uc}}$ & br\% & $\text{LR}_{\text{uc}}$ \\
\midrule
GARCH(1,1) Gaussian & $28.0$ & $59.4$ & $12.5$ & $59.5$ & $7.0$  & $0.0309$ & $0.0173$ & $0.9$ & $0.10$ & $4.0$ & $1.23$ \\
EGARCH(1,1)         & $6.3$  & $34.9$ & $4.1$  & $41.7$ & $2.8$  & $0.0430$ & $0.0173$ & $1.0$ & $0.01$ & $3.5$ & $3.03$ \\
GJR-GARCH(1,1)      & $40.1$ & $57.7$ & $28.7$ & $61.4$ & $7.6$  & $0.0330$ & $0.0173$ & $1.2$ & $0.27$ & $4.7$ & $0.10$ \\
GARCH(1,1)-$t$      & $\mathbf{57.3}$ & $\mathbf{80.8}$ & $\mathbf{36.6}$ & $65.3$ & $15.1$ & $0.0316$ & $0.0173$ & $1.0$ & $0.01$ & $4.9$ & $0.01$ \\
HAR-RV              & $0.0$  & $0.0$  & $0.0$  & $0.0$  & $189$  & $0.0566$ & $0.0173$ & $0.2$ & $5.99$ & $3.5$ & $3.03$ \\
MS-GARCH $K = 2$    & $27.7$ & $38.7$ & $24.0$ & $44.4$ & $4.7$  & $0.0367$ & $0.0173$ & $1.0$ & $0.01$ & $4.2$ & $0.82$ \\
MS-GARCH $K = 3$    & $36.1$ & $33.1$ & $28.8$ & $38.0$ & $4.1$  & $\mathbf{0.0284}$ & $0.0173$ & $1.0$ & $0.01$ & $4.2$ & $0.82$ \\
MS-GARCH $K = 4$    & $37.6$ & $38.2$ & --     & --     & $3.8$  & $0.0437$ & --       & --    & --    & --    & --    \\
MS-GARCH $K = 6$    & $34.5$ & $33.4$ & --     & --     & $4.4$  & $0.0429$ & --       & --    & --    & --    & --    \\
MS-GARCH ref.\ B.\ $K = 2$ & $0.0$  & $5.8$  & --     & --     & $4.0$  & $0.0465$ & $0.0240$ & --    & --    & --    & --    \\
MS-GARCH ref.\ B.\ $K = 3$ & $0.1$  & $5.1$  & --     & --     & $4.5$  & $0.0433$ & $0.0241$ & --    & --    & --    & --    \\
MS-GARCH ref.\ B.\ $K = 4$ & $0.0$  & $5.3$  & --     & --     & $6.0$  & $0.0446$ & $0.0241$ & --    & --    & --    & --    \\
\midrule
CHMM-N ($K = 18$)   & $94.1$ & $81.8$ & $88.2$ & $74.2$ & $5.0$  & $0.0509$ & $0.0237$ & $0.7$ & $1.58$ & $4.0$ & $3.83$ \\
CHMM-t ($K = 18$)   & $95.6$ & $\mathbf{85.7}$ & $\mathbf{91.9}$ & $79.5$ & $14.4$ & $0.0549$ & $0.0234$ & $1.2$ & $0.58$ & $5.0$ & $3.83$ \\
CHMM-L ($K = 18$)   & $94.3$ & $81.6$ & $91.1$ & $\mathbf{82.0}$ & $6.7$  & $0.0567$ & $0.0234$ & $1.4$ & $3.26$ & $5.6$ & $3.03$ \\
\textbf{CHMM-GED ($K = 18$)} & $\mathbf{95.2}$ & $84.3$ & $91.1$ & $79.1$ & $5.2$  & $0.0548$ & $0.0234$ & $0.9$ & $0.58$ & $4.4$ & $3.83$ \\
SM-CHMM-N ($K = 18$) & --     & --     & --     & --     & --     & --       & --       & $1.1$ & $0.01$ & $5.4$ & $0.82$ \\
SM-CHMM-t ($K = 18$) & --     & --     & --     & --     & --     & --       & --       & $1.4$ & $0.10$ & $5.7$ & $1.74$ \\
SM-CHMM-L ($K = 18$) & --     & --     & --     & --     & --     & --       & --       & $1.4$ & $3.26$ & $5.5$ & $3.03$ \\
\midrule
QuantGAN            & $0.0$  & $0.0$  & $0.0$  & $0.0$  & $2.1$  & $0.0591$ & --       & $1.0$ & $0.01$ & $3.3$ & $3.83$ \\
\bottomrule
\end{tabular}%
}
\end{table}

No extended GARCH variant occupied the joint (KS, ACF) corner: GARCH(1,1)-$t$ was the strongest extended row on KS ($57.3\%$ IS, $80.8\%$ OoS) but its IS excess kurtosis of $15.1$ overshot observed $7.68$ by $2\times$; EGARCH and GJR-GARCH narrowed the asymmetry but not the main gap. The SM-CHMM rows under the Viterbi-AR(1) plug-in estimator contributed only their VaR back-test columns, where the heavy-tailed sojourns sharpened Kupiec tail calibration for the Gaussian and Student-$t$ variants relative to the flat CHMM; the jointly estimated explicit-duration HSMM on SPY is in the main table, and its cross-ticker extension is left as a companion-paper direction. The QuantGAN row is the panel's deep-generative negative control: $0\%$ IS / OoS KS, simulated excess kurtosis $2.1$ vs.\ observed $7.68$, ACF-MAE comparable to the i.i.d.\ baseline. The convolutional WGAN re-implementation here is materially smaller than \citet{wiese2020quantgan}'s reference seven-block TCN with Lambert-W pre-processing; a faithful reference re-run is a deferred follow-up.



\subsubsection{Stochastic-Volatility, Multifractal, and Jump-Diffusion Baselines}
\label{sec:sv_msm_jd_baselines}

For completeness alongside the Markov-switching family of the main generator comparison, we added the canonical state-space and jump baselines that Table~\ref{tab:model_comparison} does not exercise: a stochastic-volatility (SV-AR(1)) row in the lognormal log-AR(1) tradition~\citep{taylor1982financial,harvey1994multivariate}; a Markov-switching multifractal row~\citep{calvet2004regime} at $\bar k = 8$ multipliers; and a Poisson-Gaussian jump-diffusion row~\citep{merton1976option}. The three baselines are fit on the same SPY IS window and evaluated on the same $N_{\text{paths}} = 1{,}000$ pipeline used for the main panel, under the same global seed. Implementation lives in the companion \texttt{CHMM-Model} repository.

\begin{table}[!ht]
\centering
\small
\caption{\textbf{Stochastic-volatility / multifractal / jump-diffusion baselines on SPY.} Same IS / OoS windows, $N_{\text{paths}} = 1{,}000$, seed = 20260420 as Table~\ref{tab:model_comparison}. Observed excess kurtosis: $7.68$ IS, $5.29$ OoS.}
\label{tab:sv_msm_jd}
\begin{tabular}{l cc cc cc}
\toprule
& \multicolumn{2}{c}{KS (\%) $\uparrow$} & \multicolumn{2}{c}{Excess kurtosis} & \multicolumn{2}{c}{ACF-MAE $\downarrow$} \\
\cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7}
Model & IS & OoS & IS & OoS & $|G_t|$ & $G_t$ \\
\midrule
SV-AR(1)        & $38.2$ & $35.3$ & $7.52$ & $5.26$ & $0.0466$ & $0.0239$ \\
MSM ($\bar k = 8$) & $\phantom{0}0.0$ & $\phantom{0}7.2$ & $0.62$ & $0.60$ & $0.0605$ & $0.0235$ \\
Merton-JD       & $98.3$ & $91.1$ & $3.12$ & $2.98$ & $0.0627$ & $0.0236$ \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Reading.}
The stochastic-volatility row reproduced the IS excess kurtosis target almost exactly ($7.52$ vs. observed $7.68$) and matched the CHMM-N $|G_t|$ ACF-MAE within $0.0004$ ($0.0466$ vs $0.0462$ at $K^\star = 3$); its KS pass rate was poor ($38.2\%$ IS / $35.3\%$ OoS) because the Gaussian-conditional-on-volatility marginal cannot reproduce the bimodal-bulk / heavy-tail empirical density that the regime-mixture CHMM marginals can. The Merton jump-diffusion row was the inverse trade-off: KS at $98.3\%$ IS / $91.1\%$ OoS (the highest OoS KS in the panel except for the i.i.d.\ bootstrap), but excess kurtosis $3.12$ undershot $7.68$ by half and the absolute growth-rate ACF-MAE regressed to the i.i.d.\ baseline level ($0.0627$, the same level as the bootstrap row of Table~\ref{tab:model_comparison}) because the constant-volatility diffusion plus i.i.d.\ jumps carries no temporal persistence. The multifractal row under the moment-matching fit reported here (closed-form ACF-of-log$|r|$ matching with $\bar k = 8$ binomial multipliers) did not converge to a viable generator on this $T_{\text{IS}} = 2{,}516$ window: the marginal excess kurtosis collapsed to $0.62$ and KS dropped to $0\%$. A full HMM-style filter on the $2^{\bar k}$-state latent multiplier chain~\citep{calvet2004regime} is the standard exact fit and is deferred as a companion-paper direction; under the moment-fit reported here MSM was dominated.

The main conclusion survived the addition of these three rows: no single competitor occupied the joint-fit corner that the CHMM family at $K = 18$ spans (KS $\ge 80\%$ and $|G_t|$ ACF-MAE $\le 0.06$ on every family row, with CHMM-L closing the excess kurtosis gap to within $1$ unit of observed; Table~\ref{tab:extended_baselines}). The SV-AR(1) model was the closest competitor on excess kurtosis and ACF but lost on KS; Merton-JD won on KS but lost on excess kurtosis and ACF; MSM under this fit was dominated.

\subsubsection{Leverage-Effect Diagnostic}
\label{sec:leverage_effect}

This appendix reports the \citet{cont2001empirical} leverage-effect column: the lag-1 correlation $\mathrm{Corr}(G_t, |G_{t+1}|)$ between today's signed growth rate and tomorrow's absolute growth rate, which is negative on equity returns. We computed the per-path distribution under each generator at $K = 18$, $N_{\text{paths}} = 500$, seed = $20260420$.

\begin{table}[!ht]
\centering
\small
\caption{\textbf{Leverage effect $\mathrm{Corr}(G_t, |G_{t+1}|)$, per-path distribution.} Observed values: SPY IS $-0.135$, OoS $-0.214$. Per-generator median plus $[Q_5, Q_{95}]$ across $N_{\text{paths}} = 500$. A generator captures the leverage effect when its $[Q_5, Q_{95}]$ envelope brackets the observed value. Bold marks generators whose $[Q_5, Q_{95}]$ envelope brackets the observed IS value; the OoS observed value of $-0.214$ is more negative than every generator's $Q_5$ in the panel, indicating the OoS regime carries leverage stronger than any generator (including asymmetric GARCH) reproduces under static IS-fitted parameters.}
\label{tab:leverage_effect}
\begin{tabular}{l c c c c}
\toprule
& \multicolumn{2}{c}{IS (observed $-0.135$)} & \multicolumn{2}{c}{OoS (observed $-0.214$)} \\
\cmidrule(lr){2-3} \cmidrule(lr){4-5}
Generator & median & $[Q_5, Q_{95}]$ & median & $[Q_5, Q_{95}]$ \\
\midrule
i.i.d.\ bootstrap          & $+0.000$ & $[-0.034, +0.030]$ & $+0.004$ & $[-0.069, +0.068]$ \\
GARCH(1,1) Gaussian        & $-0.004$ & $[-0.066, +0.064]$ & $-0.003$ & $[-0.112, +0.107]$ \\
\textbf{EGARCH(1,1)}                & $\mathbf{-0.093}$ & $\mathbf{[-0.147, -0.047]}$ & $-0.087$ & $[-0.199, -0.002]$ \\
\textbf{GJR-GARCH(1,1)}             & $\mathbf{-0.079}$ & $\mathbf{[-0.153, -0.017]}$ & $-0.073$ & $[-0.174, +0.023]$ \\
\textbf{CHMM-N ($K = 18$)} & $\mathbf{-0.089}$ & $\mathbf{[-0.138, -0.037]}$ & $-0.087$ & $[-0.190, +0.015]$ \\
CHMM-t ($K = 18$)          & $-0.067$ & $[-0.116, -0.020]$ & $-0.067$ & $[-0.166, +0.019]$ \\
CHMM-L ($K = 18$)          & $-0.065$ & $[-0.109, -0.020]$ & $-0.064$ & $[-0.159, +0.021]$ \\
CHMM-GED ($K = 18$)        & $-0.072$ & $[-0.113, -0.029]$ & $-0.076$ & $[-0.163, +0.025]$ \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Reading.}
The CHMM family at $K = 18$ produced a non-trivial negative leverage signal despite symmetric per-state emissions: state-mixing on $\hat{\mathbf T}$ asymmetrically couples high-$\sigma_k$ states to negative-$\mu_k$ states, carrying $\sim 65\%$ of the IS observed leverage magnitude. The CHMM-N IS interval $[Q_5, Q_{95}] = [-0.138, -0.037]$ bracketed the IS observed value $-0.135$ at the lower boundary, as did the two asymmetric GARCH variants (EGARCH and GJR-GARCH), which carry an explicit leverage term; CHMM-GED was the second-strongest CHMM entry by median without bracketing. The OoS observed leverage of $-0.214$ exceeded every generator's $Q_5$ (including asymmetric GARCH); closing this residual gap requires skew-emission CHMM extensions or refit to the OoS distribution, both deferred to companion work.

\subsubsection{Iterated Pseudo-Likelihood Student-t Copula Estimator}
\label{sec:full_tcopula_mle}

A natural concern with the main text two-step copula estimator (Kendall's-$\tau$ inversion for $\hat\Sigma$, then profile MLE on $\nu$ with $\hat\Sigma$ held fixed) is that it may be biased toward the Gaussian limit when the marginal excess kurtosis differs across assets. We addressed this with an iterated pseudo-likelihood estimator that alternates over $(\Sigma, \nu)$ on the same six-asset US-equity universe (SPY, NVDA, JNJ, JPM, AAPL, QQQ; $T_{\text{IS}} = 2{,}516$, seed $20260420$): a bracketed golden-section maximisation of the copula log-likelihood over $\nu$ at the current $\Sigma$, alternating with a method-of-moments refit of $\Sigma$ at the current $\nu$ on the $t$-quantile-transformed pseudo-uniform sample with PSD projection. Iteration terminates at $|\Delta \log L| < 10^{-3}$. The $\nu$ step is a likelihood maximisation, but the $\Sigma$ step is a moment substitution rather than an argmax, so the alternation is not a joint MLE and need not increase the likelihood at every step; we report it as a robustness check on the two-step estimator rather than as a gold-standard joint maximiser.

\begin{table}[!ht]
\centering
\small
\caption{Two-step versus iterated pseudo-likelihood Student-$t$ copula estimation on the six-asset US-equity universe. The off-diagonal correlation differences are reported as the maximum absolute difference $\max_{ij} |\Delta \rho_{ij}|$ across the $\binom{6}{2} = 15$ pairs and the median absolute difference; both are sub-$0.025$, so the main text $\nu^\star = 6$ choice on $\hat\Sigma$ obtained by Kendall's-$\tau$ inversion is robust to the estimator switch.}
\label{tab:full_tcopula_mle}
\begin{tabular}{l c c}
\toprule
Estimator & $\hat\nu$ & log-likelihood \\
\midrule
Two-step construction & $6.00$ & $6157.47$ \\
\textbf{Iterated pseudo-likelihood}   & $\mathbf{6.40}$ & $\mathbf{6163.02}$ \\
\midrule
log-L improvement   & \multicolumn{2}{c}{$+5.55$} \\
$\max |\Delta \rho_{ij}|$ across $15$ pairs & \multicolumn{2}{c}{$0.024$} \\
median $|\Delta \rho_{ij}|$ across $15$ pairs & \multicolumn{2}{c}{$0.011$} \\
$\hat\nu_{\text{full}}$ inside the Wilks $95\%$ CI $[6, 7]$? & \multicolumn{2}{c}{Yes} \\
\bottomrule
\end{tabular}
\end{table}

The full MLE moved $\hat\nu$ from $6.00$ to $6.40$ ($+5.55$ log-L) with all $|\Delta \rho_{ij}| \le 0.025$ across the $15$ pairs (Table~\ref{tab:full_tcopula_mle}); the two-step estimator is therefore not detectably biased toward the Gaussian limit and $\nu^\star = 6$ is robust to the estimator switch.
