We presented a family of continuous hidden Markov models (CHMM-N, CHMM-t, CHMM-L) trained by EM for generating synthetic financial time series, together with two cross-asset dependence extensions for multi-asset synthesis.
All three variants share the same log-space forward-backward scaffold and quantile-based initialization; the M-step is the classical Baum-Welch update for Gaussian emissions, an ECM sequence with a golden-section search over $\nu_k$ for Student-t emissions, and closed-form weighted-median / weighted-MAD estimators for Laplace emissions.
The central univariate finding is that each CHMM variant reproduces all three canonical stylized facts of equity returns (heavy-tailed distributions, negligible linear autocorrelation, and persistent volatility clustering) at a single moderate state resolution ($K = 18$, selected from a $\{3, 6, 9, 12, 15, 18, 21\}$ sweep by winning or tying on every distributional pass-rate metric) within a standard HMM framework, as an alternative to (rather than a replacement for) the HSMM \citep{bulla2006stylized} and jump-augmented \citep{alswaidan2026hybrid} routes.
The heavy-tailed CHMM-t and CHMM-L variants additionally close the residual kurtosis gap of the Gaussian CHMM-N.

Our experiments are consistent with the broader regime-switching literature in which a transition matrix with heterogeneous diagonal entries across more than two or three regimes produces a mixture of geometric sojourn-time distributions that approximates slow ACF decay.
At $K \geq 3$ with quantile-based initialization, the Baum-Welch algorithm learns a transition matrix with sufficient diagonal dominance in high-volatility states to sustain persistence, achieving ACF-MAE of 0.047--0.053 across the $K$ sweep (comparable to GARCH(1,1) at 0.048) through a sum-of-exponentials approximation whose effect is comparable to the flexible sojourn-time distributions of hidden semi-Markov models \cite{bulla2006stylized}.
This moderate-$K$ route complements the HSMM \cite{bulla2006stylized} and jump-augmented \cite{alswaidan2026hybrid} corrections to the \citet{ryden1998stylized} $K = 2$--$3$ result rather than superseding them; the headline contribution of this paper is the integrative evaluation (unified EM scaffold, twelve-generator panel, regime-conditional VaR test), not a single-route resolution of the Ryd\'{e}n et al.\ limitation.

At the prior paper's state resolution ($K_{\text{disc}} = 90$) the discrete NJ and WJ variants of \citet{alswaidan2026hybrid} match the CHMM family on coarse KS/AD pass rates, so the continuous-emissions advantage is a claim about tail structure and downstream VaR usability: simulated kurtosis near $3.5$ against observed $7.68$, aggregational-Gaussianity collapse at horizons $h = 10, 21$, and an over-conservative $5\%$ VaR breach rate of $1.7\%$ that rejects Kupiec with $\text{LR}_{\text{uc}} = 16.81$ ($p < 10^{-3}$), each of which the continuous CHMM family clears (Section~\ref{sec:utility}).

The twelve-model benchmark of Table~\ref{tab:model_comparison} confirmed that no single generator dominates all metrics, but the CHMM at $K = 18$ achieves the most balanced performance: $93.2$--$93.8\%$ IS KS across the three emission families (vs.\ $25.9\%$ GARCH, $97.9$--$98.3\%$ for the $K_{\text{disc}} = 90$ discrete variants), ACF-MAE of $0.0507$ under CHMM-N (vs.\ $0.0484$ GARCH and $0.0598$--$0.0618$ for the discrete variants), Wasserstein-1 of $0.101$--$0.103$ under CHMM-t and CHMM-L (the smallest of any parametric model), and $100\%$ IS quantile coverage.
Cross-asset univariate generalization to NVDA ($96.7\%$ IS KS), JNJ ($99.4\%$), JPM ($97.8\%$), AAPL ($98.3\%$), and QQQ ($95.5\%$) confirmed adaptability to diverse equity risk profiles; OoS evaluation on held-out 2024-01-04 through 2026-04-20 data showed robust generalization on SPY/JNJ/AAPL/QQQ but substantial OoS KS drops on NVDA ($55.8\%$) and JPM ($49.8\%$), which we attribute to the 2024--2026 macroeconomic regime (the AI-driven high-volatility regime for tech, and the interest-rate-repricing cycle for financials) and which the walk-forward refit partially recovers.

The cross-asset extensions then showed that the CHMM marginals compose cleanly with standard cross-sectional dependence models.
On the six-asset universe, both the Gaussian and Student-t copulas substantially outperformed the Single Index Model factor baseline: copulas achieved a median IS KS pass rate of $97.8\%$/$96.0\%$ versus $78.5\%$ for SIM, and an off-diagonal correlation MAE of $0.031$ (Gaussian) and $0.027$ (Student-t) versus $0.076$ for SIM.
The Student-t copula's profile MLE selected $\nu^* = 6$ degrees of freedom, and the resulting tail-dependent copula produced the best correlation reproduction overall.
These findings replicate the ordering reported by \citet{alswaidan2026hybrid} with discrete HMM marginals, suggesting that the copula advantage is intrinsic to the rank-reordering construction and transfers across marginal families.

Option pricing, implied-volatility-surface calibration, volatility-index generalization, and leverage-effect modelling are explicitly out of scope for this paper and are deferred to future work.

Several directions for future work follow from these results.
\textit{Skew-heavy-tailed emissions} are the most immediate extension: this paper closes the kurtosis gap with Student-t and Laplace emissions, and skew-$t$ or skew-Laplace emissions would additionally accommodate within-regime asymmetry and match the observed negative skewness of equity returns ($-0.75$ for SPY).
\textit{Time-varying transition dynamics}, estimated via rolling windows or Bayesian online learning of the transition matrix, would relax the stationarity assumption and should directly address the OoS cliff observed on NVDA and JPM over the $2024$--$2026$ OoS window.
\textit{Full-universe multi-asset generation via vine copulas}, scaling the copula construction from six assets to the 424-asset universe of \citet{alswaidan2026hybrid}, requires vine or factor copulas for tractability and is a natural next step.
\textit{Principled state selection} is a further refinement: our $K = 18$ choice is a multi-metric score on top of a finite sweep, and combining it with BIC or held-out log-likelihood (walk-forward) would yield a formally principled selection that plugs directly into the appendix Table~\ref{tab:sensitivity}.
Finally, \textit{regime-aware dynamic asset allocation} is the most immediate downstream application of the synthetic-data engine, following the spirit of \citet{baekimmulvey2014} by using CHMM-decoded states as conditioning variables in a dynamic allocation model and evaluating the realized utility of the resulting policy against policies calibrated on i.i.d.\ or GARCH generators.

\paragraph{Data and Code Availability.}
The simulation scripts, training and test data, and all code to reproduce the results in this paper are available under an MIT license from the paper repository: \url{https://github.com/altashly1/CHMM-paper}.
The core continuous HMM implementation (all three emission families, SIM, and Gaussian / Student-t copula modules) is available in the code repository: \url{https://github.com/altashly1/CHMM-Model}.
The global random-seed policy for reproducibility is stated once in Section~\ref{sec:metrics}.
The software environment is Julia $\geq 1.12$ with package versions pinned in \texttt{Manifest.toml}; running \texttt{Pkg.instantiate()} inside the project directory reproduces the dependency graph, after which the analysis entry-point script regenerates every table and figure.

\paragraph{Conflict of Interest.}
The authors declare no competing interests.

\paragraph{Author Contributions.}
J.V.\ directed the study.
A.A.\ developed the continuous model, implemented the Baum-Welch algorithm, implemented the Single Index Model and copula-based cross-asset extensions, conducted the empirical analysis, and generated all figures and tables.
C.J.\ contributed to methodology review and validation.
All authors edited and reviewed the final manuscript.
