The results demonstrate that a continuous Gaussian HMM trained via Baum-Welch at a single operating point ($K = 18$) achieves strong distributional fidelity ($93.8\%$ IS KS, $84.2\%$ OoS KS, $100\%$ IS quantile coverage) while retaining competitive temporal fidelity (ACF-MAE of $0.0507$); the comparison is against, rather than a replacement for, the HSMM and jump-augmented routes already in the literature.
The direct comparison against the discrete HMM of \citet{alswaidan2026hybrid}, reproduced with the authoritative $K_{\text{disc}} = 90$ / $\epsilon = 5 \times 10^{-5}$ / $\lambda = 67$ hyperparameters of the prior-paper final SPY run, shows that at that state resolution the discrete NJ and WJ baselines match the CHMM family on coarse KS/AD pass rates but miss the tail-structure, aggregational-Gaussianity, and VaR-calibration dimensions that distinguish a synthetic-data generator for downstream use.
The cross-asset extension then shows that the univariate CHMM composes cleanly with rank-based copula dependence to produce multi-asset paths that preserve every per-asset marginal while outperforming a classical factor baseline on correlation reproduction.

\paragraph{Replicating the Slow-ACF Behaviour at Higher State Resolution.}
The CHMM achieves ACF-MAE of $0.047$--$0.053$ on $|G_t|$ across $K \in \{3,\ldots,21\}$, comparable to GARCH(1,1) ($0.049$). This is consistent with the \citet{ryden1998stylized} $K = 2$--$3$ Gaussian result rather than a contradiction of it: at higher state resolution the additional degrees of freedom in the transition matrix give the CHMM access to a qualitatively different sojourn-time mixture than the one available in the regimes they studied.
The mechanism is the spectral form derived in Theorem~\ref{thm:acf_mixture} of Section~\ref{sec:theory}: at $K \ge 3$ with quantile-initialised Baum-Welch, the absolute-return ACF is the population-level identity
\begin{equation}
    \rho_{|G|}(\tau) \;=\; \sum_{k=2}^{K} w_k\, \lambda_k^{\tau},
    \qquad w_k = c_k / \sigma_{|G|}^2,
    \label{eq:acf_mixture}
\end{equation}
with $\lambda_2, \ldots, \lambda_K$ the non-unit eigenvalues of $\mathbf T$ and the weights $w_k = c_k/\sigma_{|G|}^2$ given by equations~\eqref{eq:acf_coefficients}-\eqref{eq:acf_normalised}; Corollary~\ref{cor:ryden_separation} converts the empirical Ryd\'{e}n et al.\ low-$K$ failure into a $K$-rank statement, recovering the slow quasi-hyperbolic decay observed empirically without requiring modified sojourn times \citep{bulla2006stylized}.
Two issues compound at low $K$: a single self-transition rate produces a single exponential mode that cannot be simultaneously fast (good marginal) and slow (good ACF), and the random initialisations standard in the 1990s often converge to degenerate local optima with overlapping states.
Our $K = 2$ replication on SPY (Appendix~\ref{sec:ryden_replication}, Table~\ref{tab:ryden_init}) clarifies the mechanism: ACF-MAE on $|G_t|$ is essentially constant at $\approx 0.050$ across quantile and five random-seed initialisations, matching the $K = 18$ CHMM-N.
Our $K = 2$ replication with quantile initialisation indicates that the binding low-$K$ constraint in the \citet{ryden1998stylized} setup is \emph{distributional} rather than temporal (KS at $K = 2$ drops to $67$--$72\%$): with modern initialisation the ACF is reproduced at any $K \geq 2$, while distributional fidelity is what actually requires moderate state resolution.

\paragraph{OoS KS Pass Rates Against the Test-Power Ceiling.}
At $T_{\text{OoS}} = 572$ the two-sample KS test does not reach $100\%$ on truly-correct generators; CHMM pass rates must be read against this ceiling.
Appendix~\ref{sec:ks_power} calibrates it: $1{,}000$ i.i.d.\ resamples of length $572$ drawn from the IS distribution itself pass at $90.0\%$, the OoS series resampled against itself passes at $100\%$, and a Gaussian i.i.d.\ negative control is rejected at $0\%$ at IS length, so the test retains ample power to refuse clearly-wrong generators.
The CHMM-N / CHMM-t / CHMM-L OoS KS pass rates of $79$--$86\%$ sit just below the $90\%$ ceiling (a genuine fidelity claim, not a low-power artefact), while GARCH's $58\%$ is correctly rejected and the discrete NJ / WJ at $K_{\text{disc}} = 90$ join the CHMMs in the $83$--$86\%$ band.

\paragraph{Where the Discrete Baselines Still Fall Short at $K_{\text{disc}} = 90$.}
With the prior-paper hyperparameters, NJ / WJ reach $97.9$--$98.3\%$ IS KS, comparable to the CHMM family: at this state resolution a bin-centroid discrete HMM is not a straw man on coarse pass rates.
Correcting the bin count therefore re-centers the continuous-vs.-discrete claim on three structural dimensions where the centroid scaffold cannot follow the data (Tables~\ref{tab:model_comparison}, \ref{tab:agg_kurt}, \ref{tab:sm_var}): \emph{tail weight} (NJ / WJ kurtosis clamps near $3.5$ vs.\ observed $7.68$ and CHMM-t $14.57$); \emph{aggregational Gaussianity} (NJ / WJ kurtosis collapses to $\le 0.47$ at $h=10,21$, missing the slow-decay signature CHMM tracks); and \emph{$5\%$ VaR calibration} (both discrete variants produce a $1.7\%$ breach rate against a $5\%$ target, rejecting Kupiec with $\text{LR}_{\text{uc}} = 16.81$, $p < 10^{-3}$, while all CHMM variants pass). Proposition~\ref{prop:discrete_var_floor} grounds the $5\%$ VaR failure structurally: the bin-centroid VaR estimator has a deterministic absolute-error floor of half the local bin width near the $\alpha$-quantile, and the Poisson-jump augmentation only redistributes mass across the same centroid set, so it cannot move the estimator off that floor.
At $\epsilon = 5 \times 10^{-5}$ and $\lambda = 67$ the expected jump-trigger count is below one per IS or OoS window (Section~\ref{sec:model_comparison}), so WJ is a slight perturbation of NJ rather than a meaningful tail mechanism.
This is not a criticism of \citet{alswaidan2026hybrid} on its own terms (jumps were framed there as an ACF corrective and bin-conditional Student-t emissions mitigate the centroid tail loss), but it is a sign that for tail-aware return-space metrics the joint optimisation of transitions and continuous emissions is essential.

\paragraph{The Temporal--Distributional Tradeoff.}
The twelve-model comparison of Table~\ref{tab:model_comparison} confirms the temporal-distributional tradeoff identified in \citet{alswaidan2026hybrid}: GARCH wins ACF ($0.048$) but fails KS ($25.9\%$); the bootstrap wins KS ($99.9\%$) but is the worst on ACF ($0.063$).
The CHMM at $K = 18$ occupies a distinctive corner (IS KS $93.2$--$93.8\%$, AD $86.6$--$91.2\%$, ACF-MAE within $0.002$--$0.008$ of GARCH, smallest Wasserstein-1 ($0.102$) among parametric models, and $100\%$ IS quantile coverage), a combination no other row in the comparison achieves.

\paragraph{Closing the Kurtosis Gap with Heavy-Tailed Emissions.}
CHMM-N produces excess kurtosis of $5.02$ against an observed $7.68$, a gap inherent to Gaussian emissions: a Gaussian mixture's kurtosis is bounded by the variance of component means relative to their variances, and pushing tail states out degrades the central fit.
CHMM-t addresses this directly through per-state Student-t emissions whose $\nu_k$ is updated by ECM one-dimensional search, allowing tail states to carry arbitrarily high individual kurtosis; CHMM-L's Laplace emissions add per-component excess kurtosis of $3$ rather than $0$ at zero parameter cost.
Because all three variants share the same forward-backward scaffold, the KS/AD/ACF-MAE fidelity of CHMM-N is preserved or improved without retuning $K$ or the transition structure; the AD pass rate (which weights tails more) shows the largest gain, confirming that the kurtosis gap was the residual source of AD failures.

\paragraph{Diagnosing and Shrinking Away the CHMM-t IS Kurtosis Overshoot.}
CHMM-t's unpenalised IS mixture kurtosis of $14.30$ overshoots the observed $7.69$ by $\sim 86\%$.
The bracket sweep $\nu_{\min} \in \{2.1, 2.5, 3.0, 4.0\}$ in Appendix~\ref{sec:nu_diagnostics} shows the fit does \emph{not} pile against the lower bracket (median $\nu_k = 50$, only $2/18$ states near the lower edge), and lifting $\nu_{\min}$ does not reduce the overshoot because the fit shifts posterior mass into the newly-recentred tail states.
We implement a penalised ECM with an exponential shrinkage prior on $1/\nu_k$, giving the objective $Q_{\text{pen}}(\nu_k) = Q(\nu_k) - \lambda / \nu_k$ with a golden-section search unchanged in form (details in Appendix~\ref{sec:algorithms_appendix}).
A rate sweep $\lambda \in \{0, 5, 20, 50, 100, 200\}$ on SPY IS reduces simulated excess kurtosis from $14.30$ ($\lambda = 0$) to $11.19$ ($\lambda = 5$), $8.43$ ($\lambda = 20$), $5.41$ ($\lambda = 50$), $3.96$ ($\lambda = 100$), and $3.67$ ($\lambda = 200$), so the rate $\lambda \approx 20$ recommended here brings simulated kurtosis close to the observed $7.69$ without crashing distributional fidelity (IS KS pass rate stays at $94.7\%$ against the unpenalised $95.7\%$).
Only the $\nu_{\min}$ state parameter actually moves ($2.1 \to 4.89$ at $\lambda = 20$); the upper bracket and the median $\nu_k$ remain at $50$. The overshoot is therefore a single-state artefact of the two tail regimes hitting the lower bracket under an unpenalised ECM, resolvable within the Student-t family without resort to a hierarchical prior or the Gaussian limit.
The downstream consequence (Section~\ref{sec:utility}) is that CHMM-t's $5$--$95\%$ envelopes still bracket observed VaR / ES under both the unpenalised and penalised fits, so a risk consumer incurs a widened uncertainty band rather than a systematic over-statement of tail risk under either variant.

\paragraph{Why the Jump Correction Is Not Required in This Scaffold.}
The discrete framework of \citet{alswaidan2026hybrid} estimates emissions \emph{after} the transition matrix in a three-stage procedure (binning $\to$ frequency-counted $\mathbf{T}$ $\to$ within-bin Student-t), so $\mathbf{T}$ is not jointly optimised with emissions and the jump mechanism becomes a corrective for insufficient tail-state persistence.
Baum-Welch jointly optimises $\mathbf{T}$ and $\{(\mu_k, \sigma_k)\}$: emissions and transitions co-adapt so that tail states with distinctive emission shapes also accumulate the self-transition probability $T_{kk}$ that reflects empirical clustering of extremes.
Eliminating jumps removes two hyperparameters $(\epsilon, \lambda)$ and the $48 \times 200 = 9{,}600$ simulations per $K$ that the prior paper's grid search required; CHMM needs only $20$--$40$ Baum-Welch iterations to convergence.

\paragraph{State Resolution and the Choice of $K = 18$.}
Sensitivity analysis (Section~\ref{sec:k_sensitivity}) shows ACF-MAE flat across $K$ (the dominant eigenvalue $\lambda_2$ is determined by the maximum self-transition probability, which Baum-Welch always learns to be near unity) while KS / AD pass rates rise from $K = 3$ to $K = 18$ and saturate.
The practical recommendation is $K = 18$ for distributional-fidelity-sensitive applications; $K = 3$ only when the downstream task is regime identification.

\paragraph{Cross-Asset Robustness and OoS Degradation on NVDA and JPM (Pipeline~A).}
The Pipeline~A cross-asset univariate results confirm generalisation across diverse risk profiles (SPY $95.5\%$, NVDA $96.7\%$, JNJ $99.4\%$, AAPL $98.3\%$, QQQ $95.5\%$ IS KS).
At $K = 18$, OoS degradation concentrates on NVDA ($55.8\%$) and JPM ($49.8\%$), a stationarity artefact: the IS-fixed transitions and emissions cannot capture sector-specific regime shifts (the AI-driven NVDA regime; the rate-repricing cycle for financials).
The pattern persists at every $K$ in the sweep, so it is not a model-selection issue; rolling-window re-estimation or Bayesian online learning of $\mathbf{T}$ is the natural remedy.

\paragraph{Why Copulas Beat SIM on Cross-Asset Fidelity (Pipeline~B).}
The two copula constructions dominate the SIM factor baseline on per-asset KS ($97.5\%$ / $96.2\%$ vs.\ $72.1\%$) and correlation reproduction (off-diagonal MAE $0.031$ / $0.027$ vs.\ $0.076$); all three sit on the same Pipeline~A CHMM-N marginals so only the coupling differs.
Two structural reasons.
First, rank reordering preserves marginals by construction (the simulated $\hat{G}_{j,1:T}^{(p)}$ is a permutation of the stand-alone CHMM path so its empirical CDF coincides with the CHMM marginal), whereas SIM's affine combination $\alpha_j + \beta_j G_{m,t} + \eta_{j,t}$ convolves the market factor with idiosyncratic residuals, distorting the marginal whenever $\beta_j \ne 1$ or residuals are heavy (JPM's $\hat\beta = 1.22$ produces the dramatic $31.5\%$ IS KS).
Second, the Student-t copula's edge over Gaussian on correlation MAE ($0.027$ vs.\ $0.031$) reflects joint tail dependence: profile MLE selects $\nu^* = 6$, in the range typically reported for equity indices \citep{demarta2005tcopula, mcneil2015quantitative}, implying non-trivial tail-coupling that injects rare concordant extreme moves the Gaussian copula never produces.
The same ordering holds in \citet{alswaidan2026hybrid} with discrete marginals, evidence that the copula advantage is robust to the marginal family: practitioners can swap in richer marginals without revisiting the dependence structure.

\paragraph{Three-Way Operational Split: Distribution, Unconditional VaR, Conditional VaR.}
Synthesising the extended evaluation (Section~\ref{sec:extended_evaluation}), the SM ablation (Section~\ref{sec:sm_ablation}), the conditional-VaR subsection (Section~\ref{sec:conditional_var}), and the QuantGAN / diffusion / MS-GARCH rows (Section~\ref{sec:benchmarks}), the recommendation for synthetic-data consumers splits along three axes.
On \emph{distributional fidelity} (indistinguishability on marginal and short-horizon joint metrics), \textbf{flat CHMM-t at $K = 18$} wins with the smallest MMD ($2.0 \times 10^{-5}$ IS), discriminator AUC closest to $0.5$ ($0.607$), and cleanest monotone aggregational kurtosis decay; window-diffusion is competitive on the first two but collapses on global stylised-fact coverage ($\bar{pv}_{\text{OoS}} = 0.204$).
For \emph{unconditional VaR calibration} (long-run breach rate under Kupiec), \textbf{two-regime MS-GARCH and SM-CHMM-N} both sit at the coarse-grid floor of $1\%$ Kupiec ($\text{LR}_{\text{uc}} \approx 0.01$, breach rate exactly $1.0\%$); at $T_{\text{OoS}} = 572$ this is close to point-mass identity on the achievable LR grid rather than a resolved head-to-head ranking between them, with MS-GARCH edging at $5\%$ while SM-CHMM-N edges at OoS HAR-vol forecasting and supplies regime labels.
For \emph{regime-conditional VaR} (calibration without breach clustering), the \textbf{smoother-based (Viterbi-decoded) state-conditional quantile} of flat CHMM-t is the only combination in our panel that sits below the $\chi^2_1$ Kupiec and Christoffersen thresholds at both $\alpha = 0.01$ and $0.05$ on the $2024$--$2026$ OoS window, using the same fitted model that leads the distributional axis.
This is an IS goodness-of-fit diagnostic on the regime structure rather than a one-step-ahead backtest (the Viterbi state at time $t$ depends on future observations), and at $T_{\text{OoS}} = 572$ with $\alpha = 0.01$ the Christoffersen test sees only $\approx 5$--$7$ expected breaches so small $\text{LR}_{\text{ind}}$ values should be read as "consistent with independence" rather than as resolved rankings.
The operationally correct filter-based one-step-ahead backtest (forward posterior $P(s_t \mid r_{1:t})$ under IS-fit parameters, equation~\eqref{eq:filter_conditional_var}) fails Kupiec for all three CHMM families on this OoS window (breach rate $0.2\%$ vs.\ $1\%$ target and $1.2$--$1.4\%$ vs.\ $5\%$ target, Table~\ref{tab:conditional_var} filter rows): the filter mixture of heavy-tailed per-state emissions is systematically over-conservative, and Proposition~\ref{prop:filter_var} grounds the over-conservatism in the structural inequality $F^{-1}_{\sum_k \pi_k F_k}(\alpha) \le \min_k F_k^{-1}(\alpha)$ that holds whenever the filter posterior places non-trivial mass on at least one heavy-tailed tail-state.
Taken together, the smoother diagnostic validates the IS regime partition on the OoS stress window but the operational real-time estimator does not translate that regime information into a Kupiec-calibrated VaR. A consumer needing an operational three-axis pipeline today can assemble flat CHMM-t for paths and IS regime interpretation, plus MS-GARCH (or the SM-CHMM variant) for unconditional VaR; operational regime-conditional VaR is left open, with adaptive re-fitting, an explicit state-conditional variance overlay, and the $\nu_k$-shrinkage direction below as the candidate remedies.

\paragraph{Limitations.}
Several limitations temper these conclusions.
A \textit{residual tail gap under CHMM-N} ($5.02$ simulated vs.\ $7.68$ observed kurtosis at $K = 18$) remains; CHMM-t and CHMM-L close most of it within the same EM framework, but generalised hyperbolic or $\alpha$-stable emissions would be the next step for the most extreme quantiles.
A related constraint is \textit{emission symmetry}: all three variants are symmetric within state, so SPY's negative skewness ($-0.75$) is recovered through asymmetric state occupancies and means rather than within-state asymmetry. The plug-in skew-emission ablation quantifies this: the symmetric CHMM-t at $K = 18$ already produces simulated IS skewness of $-0.667$, close to the observed $-0.752$; replacing each per-state symmetric emission with a Fernandez-Steel skew variant (per-state $\gamma_k$ fit by weighted MLE on the EM posteriors with $(\mu_k, \sigma_k, \nu_k)$ held fixed) brings simulated skewness to $-0.728$, with a small IS KS cost ($96.4\% \to 95.0\%$). Crucially, the per-state $\gamma_k$ values cluster tightly around $1$ (range $0.96$ to $1.02$, median $1.00$, zero of $18$ states with $|\gamma_k - 1| > 0.05$), so multi-state symmetric mixing absorbs most of the observed asymmetry and the residual within-state-skew gap is genuinely small. A full joint EM with $\gamma_k$ inside the M-step (rather than the plug-in used here) is a natural follow-up but is unlikely to materially change this conclusion.
Our \textit{stationarity assumption} is a second structural limitation, since $\mathbf{T}$ and the emissions are estimated on the IS window and held constant, so structural breaks (e.g.\ March 2020) may not be captured if the IS window lacks comparable events, the leading explanation for the OoS KS drop on NVDA and JPM.
The single $572$-day OoS window is complemented by the rolling-origin evaluation of Section~\ref{sec:rolling_and_weekly}, which runs five rolling one-year OoS windows across $2022$--$2026$ and surfaces a direct counter-example: the $2022$ OoS window on an $2014$--$2021$ IS sees OoS KS collapse to $0.2$--$0.4\%$ and Kupiec $\text{LR}_{\text{uc}}^{5\%} = 31.6$--$41.4$ because the rate-hike cycle is a structural break the IS does not contain. The $2023$--$2026$ rolling windows sit in the $87$--$97\%$ OoS KS range consistent with the paper's main single-window result, so the central stationarity limitation is that the CHMM scaffold is well-calibrated when the IS window contains volatility regimes comparable to those of the OoS window and fails predictably when it does not.
The \textit{KS test i.i.d.\ assumption} is relevant as well, since volatility clustering in simulated paths may slightly inflate KS pass rates relative to a block-bootstrap calibration \citep{alswaidan2026hybrid}.
The cross-asset study is restricted to a \textit{small universe} of six tickers, and scaling to the 424-asset universe of \citet{alswaidan2026hybrid} would require vine or factor copulas for tractability.
Finally, our state selection is an \textit{information-criterion rather than held-out} procedure, so a walk-forward log-likelihood would provide a complementary purely-predictive check on $K$.
