Estimation of $(\mathbf{T}, \{\boldsymbol\theta_k\}, \boldsymbol\pi)$ proceeds by EM in log-space.
This section gives the shared scaffold and the family-specific M-step updates; the state-count selection that operationalises this scaffold is reported in Section~\ref{sec:results}.

\subsection{Shared EM Scaffold}
\label{sec:em_scaffold}

All four emission families share the same log-space forward-backward
recursions and the same quantile-based initialization \cite{bilmes1998gentle}:
observations are sorted, partitioned into $K$ equal-sized chunks, and each
state's initial location and scale are seeded from the corresponding chunk;
$T_{ij}^{(0)} = \pi_k^{(0)} = 1/K$.
EM iterates to tolerance $|\mathcal{L}^{(n)} - \mathcal{L}^{(n-1)}| < 10^{-4}$
with $\texttt{max\_iter} = 60$; convergence is typically reached in
20--40 iterations.
Quantile-based initialization is what makes Baum-Welch reach a non-degenerate $K \ge 3$ fit: random initialization standard in the 1990s often converges to overlapping local optima, which is one of the two compounding sources of the low-$K$ failure observed by \citet{ryden1998stylized} (the other is the spectral one of Section~\ref{sec:theory}).
Algorithmic details and the forward-backward recursion are in Appendix~\ref{sec:supp_algorithms}.

\subsection{Family-Specific M-Step Updates}
\label{sec:mstep}

\paragraph{CHMM-N (Gaussian).}
The classical Baum-Welch M-step \cite{baum1970maximization} is closed form:
$\mu_k \leftarrow \sum_t \gamma_t(k) O_t / \sum_t \gamma_t(k)$ and
$\sigma_k^2 \leftarrow \sum_t \gamma_t(k) (O_t - \mu_k)^2 / \sum_t \gamma_t(k)$.

\paragraph{CHMM-t (Student-t via ECM).}
We adopt the ECM formulation of \citet{peel2000robust} and \citet{liu1995ml}, the same heavy-tailed innovation choice studied by \citet{abantovalle2017svm} for stochastic-volatility-in-mean models.
The latent precision
\begin{equation}
    u_{t,k} = \frac{\nu_k + 1}{\nu_k + ((O_t - \mu_k)/\sigma_k)^2}
    \label{eq:tprecision}
\end{equation}
is treated as an augmented E-step quantity, and the M-step updates
\begin{equation}
    \mu_k \leftarrow \frac{\sum_t \gamma_t(k) u_{t,k} O_t}{\sum_t \gamma_t(k) u_{t,k}},
    \qquad
    \sigma_k^2 \leftarrow \frac{\sum_t \gamma_t(k) u_{t,k} (O_t - \mu_k)^2}{\sum_t \gamma_t(k)}
    \label{eq:tupdate}
\end{equation}
have closed form conditional on $\nu_k$.
The degrees of freedom $\nu_k$ are updated by a golden-section search
maximizing the per-state $Q$-function
$Q_k(\nu) = \sum_t \gamma_t(k) \log t_\nu(O_t; \mu_k, \sigma_k)$
over the bracket $(\nu_{\min}, \nu_{\max}) = (2.1, 50)$.
The two-block CM step preserves the EM monotonicity guarantee on the compact bracket, so the incomplete-data log-likelihood is non-decreasing across iterations.

\paragraph{CHMM-L (Laplace).}
The weighted maximum-likelihood estimators are in closed form: $\mu_k$ is the posterior-weighted median of the observations (weights $\gamma_t(k)$, computed by cumulative-weight bracketing with linear interpolation between neighbouring order statistics), and $\beta_k \leftarrow \sum_t \gamma_t(k) |O_t - \mu_k| / \sum_t \gamma_t(k)$.

\paragraph{CHMM-GED (Generalized Error Distribution via three-stage ECM).}
For $f_k(x) = \mathrm{GED}(x; \mu_k, \alpha_k, p_k)$ the M-step decomposes into three CM updates per state: a bracketed $\mu_k \leftarrow \arg\min_{\mu} \sum_t \gamma_t(k)\,|O_t - \mu|^{p_k}$ via golden-section over $[\mu_k^{(n-1)} - 5\alpha_k^{(n-1)}, \mu_k^{(n-1)} + 5\alpha_k^{(n-1)}]$ (centred on the previous EM iterate), then the closed-form scale $\alpha_k \leftarrow \big[(p_k / W_k) \sum_t \gamma_t(k)\,|O_t - \mu_k|^{p_k}\big]^{1/p_k}$ with $W_k = \sum_t \gamma_t(k)$, and finally a bracketed $p_k \leftarrow \arg\max_{p \in [p_{\min}, p_{\max}]} \sum_t \gamma_t(k)\, \log f_k(O_t; \mu_k, \alpha_k, p)$ via golden-section over $[p_{\min}, p_{\max}] = [0.5, 3.0]$. Each CM step is a partial maximization, so EM monotonicity holds on the compact bracket. CHMM-N (all $p_k = 2$) and CHMM-L (all $p_k = 1$) are the boundary cases.
