Synthetic generators of equity-return time series support stress testing, regulatory backtesting, scenario design, and the training of downstream machine-learning models on data that cannot be released for privacy or licensing reasons~\cite{jordon2022synthetic, assefa2020generating, alswaidan2026hybrid}. The bar for a useful generator is that it must reproduce, simultaneously, the canonical stylized facts of daily returns: heavy-tailed marginals, negligible linear autocorrelation, and slow decay of the autocorrelation function (ACF) of absolute returns~\cite{mandelbrot1963variation, cont2001empirical}. Generalised autoregressive conditional heteroskedasticity (GARCH) models~\cite{engle1982autoregressive, bollerslev1986generalized} capture volatility clustering but not the multi-regime structure of the marginal; i.i.d.\ generators match the marginal but destroy temporal persistence; deep generators~\cite{yoon2019timegan, wiese2020quantgan, rasul2021autoregressive} reproduce some stylized facts at the cost of interpretability and reproducibility. An influential negative result has shaped two decades of work on regime-switching models in this space. \citet{ryden1998stylized} fitted Gaussian-emission hidden Markov models (HMMs) at low state counts and concluded that the geometric sojourn-time distribution of a standard Markov chain produces exponential ACF decay too fast to match the observed slow decay; \citet{bulla2006stylized} restored ACF fidelity by replacing the Markov chain with a hidden semi-Markov chain at the cost of substantially increased complexity. The empirical Ryd\'{e}n et al.\ failure compounds two constraints: a temporal one (the absolute-return ACF must admit slow-decay modes) and a distributional one (the marginal mixture must carry enough components to match observed kurtosis). A textbook closed-form identity for the absolute-return ACF, written as a sum over the non-unit eigenvalues of the transition matrix~\cite{hamilton1994time, krolzig1997markov, timmermann2000moments}, bounds the number of decay modes by the rank of the centred transition matrix. The central question of this paper is whether that bound actually matters on equity-return data once enough latent states are admitted.

We instantiate this question on a decade of daily SPY exchange-traded fund (ETF) data with a multi-year held-out window. The model is a continuous HMM (CHMM) trained by Baum-Welch~\cite{baum1970maximization, rabiner1986introduction} under quantile-based initialisation, run in four emission variants (Gaussian; per-state Student-$t$ in the tradition of~\cite{peel2000robust, liu1995ml}; Laplace; and the Generalised Error Distribution, GED) that share identical forward-backward recursions and differ only in the M-step. The empirical scope is daily US equities in stable out-of-sample periods: a non-equity stress test on the GLD ETF collapses under static fitting, a documented scope limit rather than a deferred follow-up. We extend the single-asset construction along four independent generalisation axes: a six-fold rolling-origin walk-forward on the SPY decade, a sector-balanced 30-ticker panel spanning ten Global Industry Classification Standard (GICS) sectors, a Center for Research in Security Prices (CRSP) cross-decade transfer between a non-overlapping ten-year in-sample window and a two-year out-of-sample slice, and a six-asset US-equity Student-$t$ copula extension. The copula extension composes per-asset CHMM marginals into a joint generator through Sklar's theorem. We position the construction conservatively: the i.i.d.\ bootstrap and a maximum-likelihood hidden semi-Markov model (HSMM) both exceed the CHMM on raw single-window distributional fit, but neither supports a regime-conditional risk forecast, a parametric multi-asset coupling, or a privacy-preserving synthetic representation. The CHMM is useful for things those alternatives cannot do, and we evaluate the construction on each of them in turn.

Our contributions are empirical and structural. First, we give empirical evidence that the rank bound from the bilinear identity does not bind at the cross-ticker median once enough latent states are admitted, though the rank constraint still does visible work on a left tail of tickers that introduced new regimes. This recasts the median-case \citet{ryden1998stylized} failure as a limit on the marginal mixture rather than on the chain's temporal structure. Second, we develop a unified expectation-conditional-maximisation (ECM) framework for four emission families (Gaussian, Student-$t$, Laplace, GED) under shared forward-backward recursions in which the M-step is the only difference between variants, with the GED per-state shape parameter nesting Gaussian and Laplace as boundary cases. Third, we construct a regime-conditional Value-at-Risk (VaR) by averaging the family-appropriate predictive density over the one-step-ahead state forecast, and validate it under the Christoffersen joint conditional-coverage test with Monte Carlo power calibration and an Engle-Manganelli dynamic-quantile backstop. This is the use case that no exchangeable bootstrap or unconditional GARCH benchmark can serve. Finally, we establish an empirical scope along four independent generalisation axes (rolling-origin walk-forward, sector-balanced cross-ticker panel, CRSP cross-decade transfer, six-asset Student-$t$ copula) that together protect the construction against single-window over-fitting, and we document the limit beyond which periodic refit becomes the recommended approach.

In this study, we fitted a continuous HMM with quantile-based initialisation to a decade of daily SPY returns, a sector-balanced 30-ticker panel, a CRSP cross-decade transfer between a non-overlapping ten-year in-sample window and a two-year out-of-sample slice, and a six-asset US-equity copula, under four emission families sharing identical forward-backward recursions and differing only in the M-step. We observed that the four-family panel reproduced all three symmetric Cont stylized facts at the chosen $K^\star$, and that heavy-tailed emissions narrowed the kurtosis gap relative to the Gaussian baseline without disturbing the shared framework. The regime-conditional Value-at-Risk passed the Christoffersen joint conditional-coverage test cleanly at the standard significance level on the main window and on the majority of walk-forward folds, and the Student-$t$ copula composed a six-asset basket with off-diagonal correlation reproduction substantially tighter than the single-index alternative. The interpretation is that the Ryd\'{e}n low-$K$ failure is a limit on the marginal mixture rather than on the chain's temporal structure once initialisation is fixed and the emission family is allowed to carry heavy tails, that the CHMM is useful for things stronger single-window distributional baselines cannot do (regime-conditional risk forecasting, multi-asset copula composition, parametric synthetic-data privacy), and that periodic refit is the recommended approach within the stable out-of-sample scope.
