\subsection{Data and model}
\label{sec:data_and_model}

The single-asset analysis uses daily SPDR S\&P~500 ETF (SPY) prices from January~3, 2014 to January~3, 2024, a ten-year in-sample (IS hereafter) window of length $T_{\text{IS}} = 2{,}516$ trading days, with the period from January~4, 2024 through April~20, 2026 held out for out-of-sample (OoS hereafter) evaluation, $T_{\text{OoS}} = 572$ trading days. Prices are split-adjusted but not dividend-adjusted; the data sources are described in the Data Availability Statement. The 323-day Polygon versus Alpaca/IEX overlap check found no visible break at the vendor join. For ticker $i$ and trading day $j$, with $P_{i,j}$ the session volume-weighted average price (VWAP), $\Delta t = 1/252$ the annualised trading-day step, and $r_f$ the risk-free rate (set to zero here), the annualised excess growth rate is
\begin{equation}
    G_{i,j} \equiv \left(\frac{1}{\Delta t}\right)\; \ln\!\left(\frac{P_{i,j}}{P_{i,j-1}}\right) - r_f.
    \label{eq:growth_rate}
\end{equation}
We use daily data and the matching annualised growth rates. The probabilistic framework does not depend on the time step, although application at other frequencies requires frequency-specific preprocessing, state calibration, and evaluation horizons.

The model is fit one ticker at a time. We write $G_t$ for the daily excess growth rate on trading day $t$ for the ticker being fit (equation~\eqref{eq:growth_rate} with the ticker index dropped and the day index renamed from $j$ to $t$). The sequence $\{G_t\}$ is the observed output of a continuous hidden Markov model with $K$ latent states. We write the model as $\mathcal{M} = (\mathcal{S}, \mathbf{T}, \mathbf{F}, \boldsymbol{\pi})$~\cite{rabiner1986introduction}, where $\mathcal{S} = \{1, \ldots, K\}$ is the finite state space and $s_t \in \mathcal{S}$ is the (unobserved) latent state at time $t$. The transition matrix $\mathbf{T} \in \mathbb{R}^{K \times K}$ has its $(i,j)$ entry equal to the probability of moving from state $i$ at time $t$ to state $j$ at time $t+1$, $T_{ij} = \Prob(s_{t+1} = j \mid s_t = i)$; the rows therefore sum to one (each row is a valid probability distribution over next states). The initial-state distribution $\boldsymbol{\pi} \in \mathbb{R}^K$ has $k$th entry $\pi_k = \Prob(s_1 = k)$, the probability that the chain starts in state $k$. The per-state emission densities are collected in $\mathbf{F} = \{f_k(\,\cdot\,;\boldsymbol\theta_k)\}_{k=1}^K$, where $f_k(x; \boldsymbol\theta_k)$ is the density of the day's return $G_t$ when the chain is in state $k$, evaluated at $G_t = x$; the parameter vector $\boldsymbol\theta_k$ holds that state's location, scale, and any shape parameters. The per-state cumulative distribution function (CDF) is $F_k(x; \boldsymbol\theta_k) = \int_{-\infty}^x f_k(u; \boldsymbol\theta_k)\,du$, used in the regime-conditional VaR construction.

Over many days the chain settles into a stable long-run distribution over its states. Write $\bar{\boldsymbol\pi} = (\bar\pi_1, \ldots, \bar\pi_K) \in \mathbb{R}^K$ for it, the one state distribution that a transition step leaves unchanged ($\bar{\boldsymbol{\pi}} = \bar{\boldsymbol{\pi}}\mathbf{T}$), so $\bar\pi_k$ is the long-run fraction of days the chain spends in state $k$. Collect every return the model produces and ignore which state generated each one; the density of that pooled set of returns is a weighted average of the $K$ per-state densities, each weighted by how often its state is visited,
\begin{equation}
    f(x) = \sum_{k=1}^{K} \bar{\pi}_k\, f_k(x; \boldsymbol\theta_k), \qquad \bar{\boldsymbol{\pi}} = \bar{\boldsymbol{\pi}} \mathbf{T}.
    \label{eq:mixture}
\end{equation}
We call $f(x)$ the marginal mixture: \emph{marginal} because it averages over the hidden state to describe one day's return on its own, a \emph{mixture} because it is a weighted sum of the per-state densities. A unique $\bar{\boldsymbol\pi}$ exists when $\mathbf{T}$ is irreducible (every state can eventually be reached from every other) and aperiodic (returns to a state are not locked to a fixed period); the formal conditions are given in Assumption~\ref{ass:irred}. We learn $\mathbf{T}$, the per-state parameter vectors $\boldsymbol\theta_{1:K} = (\boldsymbol\theta_1, \ldots, \boldsymbol\theta_K)$, and $\boldsymbol{\pi}$ jointly by expectation-maximisation (EM; Figure~\ref{fig:chmm_architecture}). The fitted $\boldsymbol\pi$ enters the sequence likelihood; unconditional simulation instead initialises the chain from the stationary distribution $\bar{\boldsymbol\pi}$, which removes transient dependence on the first observed day. The CHMM-N variant uses Gaussian emissions, $f_k = \mathcal{N}(\mu_k, \sigma_k^2)$ with $\sigma_k > 0$. The CHMM-t variant uses Student-$t$, $f_k = t_{\nu_k}(\mu_k, \sigma_k)$ with $\sigma_k > 0$ and $\nu_k \in [\nu_{\min}, \nu_{\max}]$; an optional exponential penalty on $1/\nu_k$ is used only in the $\lambda=20$ sensitivity specification. The CHMM-L variant uses Laplace, $f_k = \mathrm{Laplace}(\mu_k, b_k)$ with $b_k > 0$. The CHMM-GED variant uses the generalised error distribution, $f_k = \mathrm{GED}(\mu_k, \alpha_k, p_k)$ with $\alpha_k > 0$ and $p_k \in [p_{\min}, p_{\max}]$; the Gaussian and Laplace variants sit at the boundary of CHMM-GED, with $p_k = 2$ recovering Gaussian and $p_k = 1$ recovering Laplace.

\begin{figure}[!ht]
\centering
\begin{tikzpicture}[
    font=\small,
    >={Stealth[length=2mm, width=1.6mm]},
    state/.style={draw, circle, minimum size=9mm, inner sep=0pt,
                  fill=black!4, thick},
    obs/.style={draw, rectangle, rounded corners=1pt, minimum size=9mm,
                inner sep=1pt, fill=black!10, thick},
    block/.style={draw, rectangle, rounded corners=1.5pt, align=center,
                  minimum height=11mm, minimum width=34mm, inner sep=1.5mm,
                  fill=black!6, thick},
    flow/.style={-Stealth, thick, black!75},
    lbl/.style={font=\footnotesize\itshape, black!65, inner sep=1pt},
    eq/.style={font=\footnotesize, inner sep=1pt},
]
\node[state] (s1) at (0, 0) {$s_{t-1}$};
\node[state, right=14mm of s1] (s2) {$s_t$};
\node[state, right=14mm of s2] (s3) {$s_{t+1}$};
\node[left=6mm of s1] (sdots) {$\cdots$};
\node[right=6mm of s3] (edots) {$\cdots$};
\draw[flow] (sdots) -- (s1);
\draw[flow] (s1) -- (s2) node[midway, above, eq] {$T_{ij}$};
\draw[flow] (s2) -- (s3) node[midway, above, eq] {$T_{ij}$};
\draw[flow] (s3) -- (edots);
\node[lbl, above=2mm of s2] {latent regime chain $s_t \in \{1,\ldots,K\}$};
\node[obs, below=12mm of s1] (o1) {$G_{t-1}$};
\node[obs, below=12mm of s2] (o2) {$G_t$};
\node[obs, below=12mm of s3] (o3) {$G_{t+1}$};
\draw[flow] (s1) -- (o1) node[midway, right, eq] {$f_{s_{t-1}}$};
\draw[flow] (s2) -- (o2) node[midway, right, eq] {$f_{s_t}$};
\draw[flow] (s3) -- (o3) node[midway, right, eq] {$f_{s_{t+1}}$};
\node[lbl, below=2mm of o2] {observed excess growth rate $G_t \in \mathbb{R}$};
\node[block, right=18mm of edots] (mstep)
    {M-step\\ update $\mathbf{T}$, $\{\boldsymbol\theta_k\}$\\ (family-specific)};
\node[block, below=6mm of mstep] (estep)
    {E-step\\ forward-backward\\ $\gamma_t(k),\ \xi_t(i,j)$};
\draw[flow, black!75] ([xshift=-3mm]mstep.south) -- ([xshift=-3mm]estep.north);
\draw[flow, black!75] ([xshift=3mm]estep.north) -- ([xshift=3mm]mstep.south);
\node[lbl, above=1mm of mstep] {EM until $|\Delta \mathcal{L}| < 10^{-4}$};
\end{tikzpicture}
\caption{\textbf{CHMM architecture and training loop.} A $K$-state latent Markov chain with transition matrix $\mathbf{T}$ generates the regime sequence $s_t$; each $s_t$ emits the observed excess growth rate $G_t$ through a state-specific density $f_{s_t}$. The four CHMM variants share everything except the per-state emission family (Gaussian, Student-$t$, Laplace, GED). The EM procedure alternates a log-space forward-backward E-step and a family-specific M-step until $|\Delta \mathcal L| < 10^{-4}$.}
\label{fig:chmm_architecture}
\end{figure}

\subsection{Estimation by EM}
\label{sec:estimation}

With the model specified, we estimate the parameters $(\mathbf{T}, \boldsymbol\theta_{1:K}, \boldsymbol\pi)$ by EM, with all probability computations carried out in log-space for numerical stability. From here on we write $O_t \equiv G_t$ for the observation at time $t$ ($O$ is the standard symbol for an HMM observation; here it just renames $G_t$), $O_{1:T} = (O_1, O_2, \ldots, O_T)$ for the full observed sequence over the fitting window of length $T$, and $\gamma_t(k) = \Prob(s_t = k \mid O_{1:T})$ for the smoothed posterior probability that the chain was in state $k$ at time $t$, given the entire observed sequence and the current parameter iterate (``smoothed'' means computed using both past and future observations, not just past). The pair posterior $\xi_t(i, j) = \Prob(s_t = i, s_{t+1} = j \mid O_{1:T})$ is the smoothed probability of an $i \to j$ transition at time $t$, used in the transition update. The collection $\{\gamma_t(k)\}_{t=1,\ldots,T;\, k=1,\ldots,K}$ is computed at every E-step by the standard forward-backward recursion (one forward pass and one backward pass through the data). We also use the superscript $(n)$ to mark the value of any quantity at EM iteration $n$, so $(0)$ is the initial value. All four emission families share the same log-space forward-backward recursions~\cite{bilmes1998gentle} and the same quantile-based initialisation: observations are sorted, split into $K$ equal-sized chunks, and state $k$'s initial location and scale are seeded from chunk $k$, with the transition matrix and initial distribution started uniform, $T_{ij}^{(0)} = \pi_k^{(0)} = 1/K$. The EM loop runs until the observed-data log-likelihood $\mathcal{L}^{(n)} = \log p(O_{1:T} \mid \mathbf{T}^{(n)}, \boldsymbol\theta_{1:K}^{(n)}, \boldsymbol\pi^{(n)})$, the likelihood of the returns alone with the latent states summed out, stops changing meaningfully between iterations, $|\mathcal{L}^{(n)} - \mathcal{L}^{(n-1)}| < \epsilon$ with $\epsilon = 10^{-4}$, or a maximum iteration cap is reached; the iteration cap, bracket widths, one-dimensional-search settings, and the full forward-backward recursion are reported in the algorithmic appendix. Quantile-based initialisation is chosen to reduce overlap among starting state components, with the aim of improving convergence to non-degenerate $K \ge 3$ fits.

The four emission families share the latent-state forward-backward recursion, the transition update, and the quantile-based initialisation; they differ in the emission density evaluated at each E-step and in the M-step update for the per-state parameter vector $\boldsymbol\theta_k$. Given the smoothed posteriors $\gamma_t(k)$, each family applies its own parameter update. For CHMM-N, the per-state parameters are the Gaussian mean and variance, $\boldsymbol\theta_k = (\mu_k, \sigma_k^2)$, and the classical Baum-Welch M-step~\cite{baum1970maximization} updates them in closed form using weighted sample-mean and sample-variance formulas, where each observation is weighted by the posterior probability $\gamma_t(k)$ that state $k$ generated it (all sums $\sum_t$ below run over $t = 1, \ldots, T$):
\[
\mu_k \leftarrow \frac{\sum_t \gamma_t(k)\, O_t}{\sum_t \gamma_t(k)},
\qquad
\sigma_k^2 \leftarrow \frac{\sum_t \gamma_t(k)\, (O_t - \mu_k)^2}{\sum_t \gamma_t(k)}.
\]

For CHMM-t, the per-state parameters are the Student-$t$ location, scale, and degrees of freedom, $\boldsymbol\theta_k = (\mu_k, \sigma_k, \nu_k)$. We use the expectation conditional maximisation (ECM) construction of \citet{peel2000robust} and \citet{liu1995ml}, which replaces the single M-step with a sequence of conditional maximisation (CM) sub-updates that each move one block of parameters at a time. The location and scale use the augmented-data closed-form CM updates, while the degrees of freedom use the expectation/conditional-maximisation-either (ECME) variant of \citet{liu1995ml}, which maximises the observed-data likelihood directly; this is the same heavy-tailed innovation choice used by \citet{abantovalle2017svm} for stochastic-volatility-in-mean models. The construction uses a standard representation of the Student-$t$ as a Gaussian with random variance: conditional on a latent positive scalar $u$, the observation is Gaussian, $G_t \mid s_t = k,\, u \sim \mathcal{N}(\mu_k, \sigma_k^2/u)$, and $u$ itself is drawn from a Gamma distribution, $u \sim \Gamma(\nu_k/2,\, \nu_k/2)$. Averaging over $u$ recovers the Student-$t$. In ECM, $u$ is treated as an additional latent variable, and the E-step adds its conditional expectation under the current parameter iterate,
\begin{equation}
    u_{t,k} \equiv \E[u \mid O_t, s_t = k] = \frac{\nu_k + 1}{\nu_k + ((O_t - \mu_k)/\sigma_k)^2},
    \label{eq:tprecision}
\end{equation}
and the M-step then updates
\begin{equation}
    \mu_k \leftarrow \frac{\sum_t \gamma_t(k) u_{t,k} O_t}{\sum_t \gamma_t(k) u_{t,k}},
    \qquad
    \sigma_k^2 \leftarrow \frac{\sum_t \gamma_t(k) u_{t,k} (O_t - \mu_k)^2}{\sum_t \gamma_t(k)}
    \label{eq:tupdate}
\end{equation}
for the location and scale, both in closed form given $\nu_k$. The degrees of freedom $\nu_k$ are then updated by a finite one-dimensional search on the bounded interval $[\nu_{\min}, \nu_{\max}]$ for a high-value point of the state-$k$ objective $Q_k(\nu) = \sum_t \gamma_t(k) \log t_\nu(O_t; \mu_k, \sigma_k)$, the observed-data state-weighted Student-$t$ log-likelihood rather than the augmented-data CM update in $\E[\log u]$. This direct maximisation is the ECME conditional-maximisation step; in exact form it preserves the monotone observed-data ascent~\cite{wu1983convergence, mengrubin1993ecm}, and the fixed-iteration numerical search used here approximates that maximiser, so we assess numerical convergence from the observed-data log-likelihood trace and stop at the stated tolerance or iteration cap rather than claim exact monotonicity for every finite-precision step.

For CHMM-L, the per-state parameters are the Laplace location and scale, $\boldsymbol\theta_k = (\mu_k, b_k)$, and both have closed-form updates. The location $\mu_k$ is a weighted median of the observations $\{O_t\}_{t=1}^T$, where each observation is weighted by the posterior probability $\gamma_t(k)$ that state $k$ generated it. The scale $b_k \leftarrow \sum_t \gamma_t(k)\, |O_t - \mu_k| / \sum_t \gamma_t(k)$ is the weighted mean absolute deviation of the observations from $\mu_k$, using the same weights.

For CHMM-GED, the per-state parameters are the GED location, scale, and shape, $\boldsymbol\theta_k = (\mu_k, \alpha_k, p_k)$, with conditional density $f_k(x; \mu_k, \alpha_k, p_k) = \mathrm{GED}(x; \mu_k, \alpha_k, p_k)$. The three parameters cannot all be updated together in closed form, so the M-step uses three sequential block updates. The location update searches a bounded interval for a low-value point of $\sum_t \gamma_t(k)\,|O_t - \mu|^{p_k}$. This objective is convex for $p_k \ge 1$ but need not be convex for $p_k < 1$, so the finite search is a numerical update rather than a certified global conditional minimisation. The scale then has a closed-form update at the just-updated location, $\alpha_k \leftarrow \big[(p_k / W_k) \sum_t \gamma_t(k)\,|O_t - \mu_k|^{p_k}\big]^{1/p_k}$, where $W_k = \sum_t \gamma_t(k)$ is the total posterior weight on state $k$. The shape is updated last by a bounded one-dimensional search over $[p_{\min}, p_{\max}]$ at the just-updated location and scale. We therefore describe CHMM-GED as an ECM-style numerical block-coordinate procedure and diagnose convergence from its observed-data log-likelihood trace; the exact ECM monotonicity theorem applies only when every block is conditionally maximised.

\subsection{Evaluation}
\label{sec:evaluation}

With the model fitted, we evaluate the CHMM at two levels, single-asset and multi-asset, across four datasets: the SPY in-sample and out-of-sample windows, a sector-balanced 30-ticker panel spanning ten GICS sectors with three large-cap representatives each, a CRSP cross-decade transfer between a 1994 to 2004 in-sample window and a 2004 to 2006 out-of-sample slice, and a six-asset US-equity basket (SPY, NVDA, JNJ, JPM, AAPL, QQQ). The \emph{single-asset} evaluation fits one CHMM per ticker and covers the stylized-fact diagnostics, $K$ selection, main generator comparison, cross-ticker generalisation panel, CRSP cross-decade transfer, walk-forward stress test, and Value-at-Risk panel. The \emph{multi-asset} evaluation keeps each asset's fitted CHMM marginal fixed and models only how the assets move together. A copula makes this split possible: Sklar's theorem~\cite{sklar1959fonctions} says any joint distribution factors into its one-asset marginals and a copula that carries the dependence between them, so the marginals and the dependence can be fitted separately. We use a Student-$t$ copula~\cite{demarta2005tcopula, mcneil2015quantitative} on the six-asset basket, with its correlation matrix fixed from Kendall's $\tau$ via $\rho=\sin(\pi\tau/2)$; its degrees of freedom $\nu^\star$ are selected by a grid search over the profile likelihood. To draw joint samples we use the rank-reordering scheme of \citet{iman1982distribution}, which reorders the per-asset CHMM draws into the copula's dependence pattern without altering any asset's own values, so the reordering preserves each asset's simulated marginal exactly. A single-index model (SIM), a Gaussian copula on the same CHMM marginals, and a truncated C-vine are reported in the appendix as multi-asset comparators.

The single-asset CHMM is benchmarked against one representative from each model class, all fit on the same IS window: a plain i.i.d.\ empirical bootstrap~\cite{efron1979bootstrap} as the strongest non-parametric baseline; Gaussian and Laplace i.i.d.\ generators as marginal baselines; GARCH(1,1) with Gaussian~\cite{bollerslev1986generalized} and Student-$t$~\cite{bollerslev1987conditionally} innovations as conditional-variance baselines; and Markov-switching GARCH (MS-GARCH)~\cite{haas2004new} as the closest regime-switching comparator. Extended GARCH-family, semi-Markov, stochastic-volatility, multifractal, and jump-diffusion baselines are reported in the appendix; a deep-generative QuantGAN row~\cite{wiese2020quantgan} appears in the main comparison table as a negative control, with its construction detailed in the appendix. For every model in the panel we simulate $P$ independent paths matching the window being scored: $T_{\text{IS}}$ in sample and $T_{\text{OoS}}=572$ out of sample~\cite{stenger2024thinking}; the path count and the deterministic seed used in each run are reported in the corresponding result table. We score the panel on three primary metrics. The two-sample Kolmogorov-Smirnov (KS) pass rate at significance level $\alpha = 0.05$~\cite{kolmogorov1933, smirnov1948} counts the fraction of simulated paths whose marginal distribution is not rejected against the observed window, measuring distributional fidelity at the marginal level. The asymptotic two-sample KS null assumes i.i.d.\ observations, whereas the observed and simulated paths carry serial dependence, so we read this pass rate as a descriptive marginal-fidelity score rather than a calibrated test and recalibrate it with the stationary block bootstrap in the appendix. Mean simulated excess kurtosis, read against the observed kurtosis on the same window, measures how well the heavy-tailed marginal is reproduced. The mean absolute error (MAE) between the observed and path-averaged simulated autocorrelation function (ACF) of absolute returns $|G_t|$ over $L = 252$ lags (one trading year) is
\begin{equation}
    \text{ACF-MAE} = \frac{1}{L}\sum_{\tau=1}^{L} \left| \hat\rho^{\text{obs}}_{|G|}(\tau) - \bar{\hat\rho}^{\,\text{sim}}_{|G|}(\tau) \right|,
    \label{eq:acf_mae}
\end{equation}
where $\tau$ is the lag index, $|G| = \{|G_t|\}_{t=1}^T$ is the absolute-return sequence, $\hat\rho^{\text{obs}}_{|G|}(\tau)$ is the sample autocorrelation of $|G|$ at lag $\tau$ on the observed window, and $\bar{\hat\rho}^{\,\text{sim}}_{|G|}(\tau)$ is the same quantity averaged across the $P$ simulated paths; this captures volatility clustering, the third Cont stylized fact~\cite{cont2001empirical}. The Value-at-Risk panel additionally reports the unconditional Kupiec coverage test~\cite{kupiec1995techniques}, the Christoffersen joint conditional-coverage test~\cite{christoffersen1998evaluating} as our primary VaR backtest, and an expected-shortfall (ES) envelope (the $[5\%, 95\%]$ band of simulated ES across paths). Auxiliary distributional metrics (block-bootstrap KS recalibration, Anderson-Darling (AD), Wasserstein-1 ($W_1$), Hellinger distance, and continuous ranked probability score (CRPS)) are reported in the appendix to confirm robustness of the cross-generator ranking. The models and evaluation pipeline were implemented in Julia in the companion repository \texttt{CHMM-Model}.
