\paragraph{Hidden Markov models in finance.}
\citet{hamilton1989new} introduced Markov-switching to economics; \citet{schaller1997regime} estimated regime-switching specifications on US monthly stock returns and documented the predictive content of the latent regime for return forecasting. The seminal evaluation against the Cont stylized facts is due to \citet{ryden1998stylized}: Gaussian-emission HMMs at $K = 2$--$3$ states fail the slow ACF decay of absolute returns because the geometric sojourn-time distribution of a standard Markov chain produces exponential decay too fast to match the observed signature. \citet{bulla2006stylized} restored fidelity with hidden semi-Markov models at substantially higher complexity. \citet{nystrup2017long} extends the same line to long-memory volatility. Maximum-likelihood fitting of heavy-tailed innovations follows \citet{peel2000robust} and \citet{liu1995ml}; the Generalized Error Distribution \cite{subbotin1923law, box1973bayesian} provides a continuous shape parameter that interpolates between Gaussian ($p = 2$) and Laplace ($p = 1$) and underlies our CHMM-GED variant in the same role that the per-state degrees of freedom $\nu_k$ plays in CHMM-t. The discrete-state predecessor of this paper, \citet{alswaidan2026hybrid}, reproduced the same three stylized facts via Laplace quantile binning plus a Poisson-jump duration mechanism; here we replace the discretization step with continuous emissions and recover the same target through the spectral mechanism of Section~\ref{sec:theory}.

\paragraph{GARCH and its extensions.}
GARCH \cite{engle1982autoregressive, bollerslev1986generalized} encodes volatility persistence through a single decay parameter but imposes a unimodal innovation distribution and fails distributional Kolmogorov--Smirnov on equity returns. Heavy-tailed innovations \cite{bollerslev1987conditionally}, asymmetric leverage extensions (EGARCH, \citealp{nelson1991conditional}; GJR-GARCH, \citealp{glosten1993relation}; threshold GARCH, \citealp{zakoian1994threshold}), and regime-switching extensions (MS-GARCH, \citealp{haas2004new}, with the \texttt{MSGARCH} R package of \citealp{ardia2019msgarch} the standard reference implementation) close some gaps but inherit the unimodal-innovation constraint. Realised-variance frameworks \cite{andersen1997heterogeneous, corsi2009simple} provide a complementary forecasting baseline.

\paragraph{Deep generative models for financial time series.}
Generative adversarial network (GAN)-based generators reproduce the visual appearance of return series but typically fail volatility-clustering tests unless architectures encode temporal dependence explicitly \cite{takahashi2019modeling, kwon2024can}; the TimeGAN family of \citet{yoon2019timegan}, the convolutional Wasserstein GAN (WGAN) QuantGAN of \citet{wiese2020quantgan}, and the autoregressive-diffusion approach of \citet{rasul2021autoregressive} are the standard references. We include QuantGAN as the deep-generative row in our extended panel (Appendix~\ref{sec:quantgan_supp}); the present paper's contribution is grounded against parametric and non-parametric baselines.

\paragraph{Cross-asset dependence and synthetic-data evaluation.}
For multi-asset synthesis, the Single Index Model \cite{sharpe1963simplified} propagates a market factor through a rank-one linear decomposition but distorts each non-market marginal; Sklar's theorem \cite{sklar1959fonctions, nelsen2006introduction} supports a cleaner decomposition in which each asset's fitted marginal is preserved and only the copula is estimated, with Gaussian and Student-$t$ copulas \cite{demarta2005tcopula, mcneil2015quantitative} the standard choices and rank-reordering simulation \cite{iman1982distribution} the standard injection. Dynamic-conditional-correlation models \cite{engle2002dynamic} and composite-likelihood estimators for vast covariance matrices \cite{pakel2021fitting} are the standard alternatives at higher $d$; the present paper's six-asset construction sits well below the dimensions where those estimators become essential, but the rank-based copula architecture composes with either at scale. Synthetic-data evaluation is anchored on the framework of \citet{stenger2024thinking}, the proper-scoring-rule literature \cite{gneiting2007strictly, gneiting2014probabilistic, diebold1995comparing}, kernel two-sample tests \cite{gretton2012kernel}, and signature-based metrics \cite{chevyrev2016primer, ni2020conditional}.

\paragraph{State-space stochastic volatility.}
Stochastic-volatility (SV) models \cite{taylor1982financial} are the continuous-latent counterpart of the discrete-latent CHMM: log-volatility evolves as a latent AR(1) and likelihood inference proceeds via simulation-based estimators \cite{jacquier1994bayesian, kim1998stochastic} or hidden-Markov approximations on a discretised volatility grid \cite{rossi2006volatility, abantovalle2017svm}. Markov-switching SV \cite{so1998stochastic} sits between the two formulations. We include a Markov-switching SV row in the extended baseline panel (Appendix~\ref{sec:extended_baselines}) for direct comparison; the architectural difference is that the CHMM treats the latent regime as a finite-state object with $K$ free per-state densities, whereas SV ties all states through a single AR(1) on a one-dimensional log-volatility latent. The former gives more representational capacity at the marginal level (the kurtosis points of Table~\ref{tab:variant_choice}); the latter gives a strictly stationary latent process by construction.

\paragraph{Regime breaks and online change-point detection.}
The W2 (COVID 2020) and W4 (2022 rate-hike onset) walk-forward stress folds (Section~\ref{sec:results}) are regime introductions that the IS distribution does not span: any IS-fitted symmetric-emission CHMM rejects them by KS, and faster refit cadences do not close the gap (Appendix~\ref{sec:walkforward_refit_cadence}). The natural complements are Bayesian online change-point detection \cite{adams2007bayesian} and the particle-filter recursion for multiple change-points of \citet{fearnhead2007online}, both of which model the run-length distribution of an underlying segmentation explicitly and would allow the transition matrix or per-state densities to refit at detected break points rather than on a fixed quarterly cadence. We do not implement this here; it is one of the companion-paper directions discussed in Section~\ref{sec:discussion}.

\paragraph{Signature- and diffusion-based market generators.}
A second wave of deep market generators sits beyond the GAN-based panel surveyed above. Path-signature methods \cite{chevyrev2016primer} encode an entire return path as a graded tensor of iterated integrals; \citet{ni2021sig}, \citet{buehler2020data}, and \citet{boursin2022deep} use the signature either as a moment-matching loss or as the conditioning variable for a generative model and report tighter rough-path statistics than convolutional baselines on equity tape. Score-based and diffusion-based generators \cite{ho2020denoising, rasul2021autoregressive, tashiro2021csdi, liao2024diffusion} replace adversarial training with a denoising objective that is more stable on heavy-tailed marginals. The two classes share the deep-generative tradeoff against the present paper: tighter visual fidelity to the empirical tape, opaque per-state semantics, and no parametric copula handle for multi-asset composition.
