This appendix collects supplementary material referenced in the main body: the public API of the companion package, the EM forward-backward recursions and algorithm pseudocode, the validation metric definitions, formal propositions, the full state-resolution and multi-emission sensitivity panels, robustness and KS-power calibration, the extended GARCH and SM-CHMM baselines, the full cross-asset panel, and the $K$-selection and $\nu_k$ diagnostics referenced from the discussion.

\subsection{CHMM-Model.jl Public API and Reproducer}
\label{sec:supp_chmm_api}

The companion Julia package \texttt{CHMM-Model.jl} (\url{https://github.com/altashly1/CHMM-Model}) is the entry point used to regenerate every table and figure in this paper from a single seed root under Julia $\geq 1.12$. The public surface is a single \texttt{CHMM} type, three emission constructors, a unified \texttt{fit!} method (log-space Baum-Welch / ECM), and a unified \texttt{simulate} method.

\begin{verbatim}
using CHMM
chmm_n   = make_chmm_n(returns;   K = 18)    # Gaussian emissions
chmm_t   = make_chmm_t(returns;   K = 18)    # Student-t emissions
chmm_l   = make_chmm_l(returns;   K = 18)    # Laplace emissions
chmm_ged = make_chmm_ged(returns; K = 18)    # GED (per-state shape p_k)
fit!(chmm_n; max_iter = 60, tol = 1e-4)      # log-space Baum-Welch
paths = simulate(chmm_n; T = 2516, n = 1000) # 1000 paths of length T
\end{verbatim}

The same \texttt{simulate} interface is used for every benchmark generator (Bootstrap, Gaussian i.i.d., Laplace i.i.d., GARCH(1,1)) so that the seven-metric panel evaluates each row on byte-identical windows. The Pipeline~B Student-t copula sampler is exposed as \texttt{simulate\_copula(\dots; nu = 6)} on the per-asset CHMM marginals; the rank-reordering construction is one function call. Running \texttt{julia --project=. run\_full\_rebuild.jl} from the paper repository (\url{https://github.com/altashly1/CHMM-paper}) reproduces every numerical artifact under the global seed policy of Section~\ref{sec:metrics}; \texttt{Manifest.toml} pins the dependency graph.

\paragraph{Sub-seed derivation rule.}
The deterministic global seed root is \texttt{20260420}; per-experiment sub-seeds follow the additive rule \texttt{sub\_seed = SEED + offset(experiment) + offset(replicate)}, with experiment offsets in $\{0, 1{,}000, 100, 10{,}000, \ldots\}$ chosen to avoid collisions across the runners in the public Julia package. For example, the headline simulation in \texttt{run\_diagnostics.jl} uses \texttt{SEED + (path index)} for each per-path random draw, the multi-seed Monte Carlo in \texttt{run\_multiseed\_headline.jl} uses base \texttt{20260422 + 100$\cdot$(seed index)} for ten independent replicates, and the bracket / shrinkage sweeps add a further per-rate offset. The full sub-seed map is enumerated at the top of each runner script; downstream \texttt{Random.seed!()} calls are deterministic and order-independent within each runner.

\subsection{CHMM Algorithms and Derivations}
\label{sec:supp_algorithms}

This appendix collects the algorithmic content for the CHMM family: the forward-backward recursions and weighted M-step updates, the complete training pseudocode, and per-state degrees-of-freedom diagnostics for CHMM-t.

\input{sections/algorithms_appendix}

\subsection{Validation Metric Details}
\label{sec:supp_metrics}
\label{sec:metric_details}
\input{sections/metrics_appendix}

\subsection{Formal Theoretical Statements}
\label{sec:supp_propositions}

The following propositions are referenced in Section~\ref{sec:theory} and stated formally here for reference; each is a one-line specialisation of a standard result and the proof sketch is one sentence.

\begin{proposition}[ECM monotonicity for CHMM-t and CHMM-GED (and EM monotonicity for CHMM-N, CHMM-L)]
\label{prop:ecm_monotone}
Under compactness of the parameter space ($\mu_k$, $\sigma_k$ or $\beta_k$ or $\alpha_k$, $\nu_k \in [\nu_{\min}, \nu_{\max}]$, $p_k \in [p_{\min}, p_{\max}]$ each in a closed bounded interval, $T_{ij} \ge \epsilon_T > 0$), the two-block CM step for CHMM-t of Section~\ref{sec:mstep}, namely the closed-form weighted updates of $(\mathbf T, \boldsymbol\pi, \mu_k, \sigma_k)$ at fixed $\nu_k$, followed by a golden-section maximisation of the per-state $Q$-function over $\nu_k$, satisfies $Q(\theta^{(n+1)} \mid \theta^{(n)}) \ge Q(\theta^{(n)} \mid \theta^{(n)})$, and consequently $\mathcal L(\theta^{(n+1)}) \ge \mathcal L(\theta^{(n)})$ where $\mathcal L$ is the incomplete-data log-likelihood. The same monotonicity holds for the closed-form CHMM-N and CHMM-L M-steps under the standard EM guarantee, and for the three-stage CHMM-GED CM sequence (bracketed $\mu_k$, closed-form $\alpha_k$, bracketed $p_k$) of equations~\eqref{eq:ged_mu}--\eqref{eq:ged_p}.
\end{proposition}
\noindent\emph{Proof sketch.} Each CM step does not decrease $Q$ by definition (closed-form steps are exact maximisers, golden-section steps are bracketed maxima within the prescribed interval); the standard EM inequality $\mathcal L(\theta^{(n+1)}) - \mathcal L(\theta^{(n)}) \ge Q(\theta^{(n+1)} \mid \theta^{(n)}) - Q(\theta^{(n)} \mid \theta^{(n)})$ \cite{wu1983convergence, mengrubin1993ecm} converts $Q$-monotonicity into log-likelihood monotonicity. Compactness gives boundedness above, hence convergence.

\paragraph{Numerical full-rank check on $\hat{\mathbf T}$.}
The Allman-Matias-Rhodes 2009 identifiability theorem requires $\hat{\mathbf T}$ of full rank. We compute the singular-value spectrum of $\hat{\mathbf T}$ at $K = 18$ across the four emission families on SPY IS (Table~\ref{tab:t_singular_values}). All four matrices are full-rank in the strict sense, with smallest singular value $\sigma_{\min}(\hat{\mathbf T}) \in [0.0007, 0.0170]$ and condition numbers in the range $63$--$1{,}620$. CHMM-GED is the most well-conditioned ($\sigma_{\min} = 0.017$, cond $= 63.5$); CHMM-t is the least ($\sigma_{\min} = 0.0007$, cond $= 1{,}620$), consistent with the per-state $\nu_k$ ECM lower-bracket pinning making one or two heavy-tail states near-degenerate. The deflated matrix $\hat{\mathbf T} - \mathbf 1\bar{\boldsymbol\pi}^\top$ has numerical rank exactly $K - 1 = 17$ across all four families, the rank statement underlying Section~\ref{sec:theory}.

\begin{table}[h]
\centering
\small
\caption{\textbf{Numerical full-rank check on $\hat{\mathbf T}$ at $K = 18$ across four CHMM emission families on SPY IS} (seed = $20260420$). $\sigma_{\min}$, $\sigma_{\max}$, and the condition number $\sigma_{\max}/\sigma_{\min}$ are reported on $\hat{\mathbf T}$ itself; $|\lambda_2|$ is the second-largest absolute eigenvalue (the dominant decay mode in the ACF identity, equation~\eqref{eq:acf_normalised}); ``rank(def)'' is the numerical rank of $\hat{\mathbf T} - \mathbf 1\bar{\boldsymbol\pi}^\top$ at tolerance $10^{-10}$, expected to equal $K - 1 = 17$.}
\label{tab:t_singular_values}
\begin{tabular}{l c c c c c}
\toprule
Family & $\sigma_{\min}(\hat{\mathbf T})$ & $\sigma_{\max}(\hat{\mathbf T})$ & cond $\hat{\mathbf T}$ & $|\lambda_2|$ & rank$(\hat{\mathbf T} - \mathbf 1\bar{\boldsymbol\pi}^\top)$ \\
\midrule
CHMM-N   & $0.00532$ & $1.0866$ & $204$   & $0.929$ & $17 / 17$ \\
CHMM-t   & $0.00067$ & $1.0796$ & $1{,}620$ & $0.914$ & $17 / 17$ \\
CHMM-L   & $0.00398$ & $1.0928$ & $275$   & $0.896$ & $17 / 17$ \\
CHMM-GED & $0.01702$ & $1.0806$ & $63.5$  & $0.911$ & $17 / 17$ \\
\bottomrule
\end{tabular}
\end{table}

\begin{proposition}[Identifiability up to label permutation]
Under irreducibility and aperiodicity of $\mathbf T$ (Assumption~\ref{ass:irred}), the emission contrast condition (Assumption~\ref{ass:moments}), and pairwise distinctness of the per-state parameter tuples $\boldsymbol\theta_k$ across $k = 1, \ldots, K$, the CHMM parameter is identifiable up to label permutation for CHMM-N (Gaussian) and CHMM-L (Laplace).
For CHMM-t (with per-state $\nu_k \in [\nu_{\min}, \nu_{\max}]$) and CHMM-GED (with per-state $p_k \in [p_{\min}, p_{\max}]$) the same conclusion holds modulo the standard caveat that pairwise distinctness must hold jointly across all components of $\boldsymbol\theta_k = (\mu_k, \sigma_k, \nu_k)$ or $(\mu_k, \alpha_k, p_k)$, since within either bracket a Gaussian-shaped emission with inflated scale can mimic a Laplace-shaped emission with deflated scale at intermediate quantiles.
\end{proposition}
\noindent\emph{Proof sketch.} The HMM identifiability theorem of \citet{allman2009identifiability} requires $\mathbf T$ of full rank and per-state emissions linearly independent in the space of densities on $\R$. Full-rank $\mathbf T$ follows from irreducibility \cite{levinperes2017markov}. For CHMM-N and CHMM-L, linear independence of Gaussian and Laplace emissions with distinct location-scale parameters is direct from \citet{yakowitzspragins1968}, who establish that finite mixtures over location-scale families with strictly decreasing-tail densities are identifiable (the CHMM-N and CHMM-L marginals are special cases). For CHMM-t at fixed shape and CHMM-GED at fixed shape, location-scale identifiability is the same Yakowitz-Spragins result; for finite mixtures of Student-$t$ with varying degrees-of-freedom the standard reference is \citet{lindsay1995mixture}, Chapter~3, which gives identifiability via the moment-generating function over the bracketed range of shape parameters and a non-degeneracy condition on the joint location-scale-shape tuple. The CHMM-GED case under bracketed $p_k \in [p_{\min}, p_{\max}]$ is a parallel application of the moment-matching argument: the GED characteristic function $\hat b(t; \mu, \alpha, p) = \exp(i\mu t)\, M(t; \alpha, p)$ has a unique inverse over the bounded shape range up to permutation of components, modulo the joint-distinctness caveat on $(\mu_k, \alpha_k, p_k)$. Strict linear independence at the boundary cases $p_k = 1$ (Laplace) and $p_k = 2$ (Gaussian) is automatic from Yakowitz-Spragins; in the bracket interior, the joint distinctness condition above is required to rule out the location-scale-shape trade-off.

\begin{proposition}[MLE consistency]
Under Assumptions~\ref{ass:irred}--\ref{ass:moments} and the assumption that the true parameter lies in the interior of a compact $\Theta$, the MLE $\hat\theta_T = \arg\max_{\theta \in \Theta} \mathcal L_T(\theta)$ is consistent: $\hat\theta_T \xrightarrow{p} \theta_0$ as $T \to \infty$, modulo label permutation.
\end{proposition}
\noindent\emph{Proof reference.} \citet{bickel1998asymptotic} (Theorem~1) and \citet{doucmoulinesryden2004} apply directly under our assumption set: irreducibility supplies the mixing condition, finite second moment the moment condition, compactness the parameter-space condition.

\begin{proposition}[Marginal preservation under rank reordering]
\label{prop:rank_marginals}
Let $\tilde g_{j, 1:T}$ be a sample of size $T$ from the per-asset CHMM marginal of asset $j$, and let the rank-reordered output be $\hat g_{j,t}^{(p)} = \tilde g_{j, (r_{j,t}^{(p)})}^{(p)}$ where $r_{j,t}^{(p)}$ is the rank of the copula sample. Then for every asset $j$ the empirical CDF of $\{\hat g_{j,t}^{(p)}\}_{t=1}^T$ equals that of $\{\tilde g_{j,t}^{(p)}\}_{t=1}^T$.
\end{proposition}
\noindent\emph{Proof sketch.} The map $t \mapsto r_{j,t}^{(p)}$ is a permutation almost surely, so the multiset of values is preserved; the empirical CDF depends only on the multiset.

\subsection{Spectral ACF Identity: Extended Derivation}
\label{sec:supp_spectral_acf}

Equation~\eqref{eq:acf_normalised} is the closed-form mixture-of-eigenvalues form for the absolute-return ACF used in Section~\ref{sec:theory} and invoked throughout Sections~\ref{sec:results}--\ref{sec:discussion}. The bilinear cross-moment identity $\E[|G_t|\,|G_{t+\tau}|] = \mathbf m^\top \mathrm{diag}(\bar{\boldsymbol\pi})\,\mathbf T^\tau\,\mathbf m$ and its eigendecomposition over the non-unit eigenvalues of $\mathbf T$ are textbook material in the regime-switching literature (\citealp{hamilton1994time}, Section~22.2; \citealp{krolzig1997markov}, Chapter~3; the moment-generating-function form for general Markov-switching specifications with conditional means and variances depending on the latent state is given by \citealp{timmermann2000moments}); the underlying Markov-chain spectral decomposition is standard \cite{levinperes2017markov}. We restate the absolute-return specialisation here as a self-contained theorem with a step-by-step proof, and record the lag-zero behaviour and the complex / non-diagonalisable extensions that the main body does not. Our contribution on this axis is the explicit application of the spectral form to recast the empirical \citet{ryden1998stylized} low-$K$ failure as a rank statement on $\mathbf T - \mathbf 1\bar{\boldsymbol\pi}^\top$ (Section~\ref{sec:theory}, ``Ryd\'{e}n separation as a $K$-rank statement'') and the empirical demonstration that the rank constraint is the binding axis at moderate $K$ on equity-return data (Sections~\ref{sec:results}, \ref{sec:discussion}).

\begin{theorem}[Mixture-of-eigenvalues identity for the absolute-return ACF]
\label{thm:supp_spectral_acf}
Let $\{(s_t, G_t)\}_{t \in \mathbb{Z}}$ be a stationary CHMM on $\{1, \ldots, K\}$ with irreducible aperiodic transition matrix $\mathbf T$ (Assumption~\ref{ass:irred}) and per-state emissions of finite second moment (Assumption~\ref{ass:moments}). Write $\bar{\boldsymbol\pi}$ for the unique stationary distribution, $\mathbf D = \mathrm{diag}(\bar{\boldsymbol\pi})$, $m_k = \E[|G_t| \mid s_t = k]$, $\mathbf m = (m_1, \ldots, m_K)^\top$, $\mu = \E|G_t| = \bar{\boldsymbol\pi}^\top \mathbf m$, and $\sigma_{|G|}^2 = \mathrm{Var}(|G_t|)$. Assume $\mathbf T$ is diagonalisable with right eigenvectors $\mathbf v_k$ and left eigenvectors $\mathbf w_k$ biorthonormal, $\mathbf w_k^\top \mathbf v_l = \delta_{kl}$, indexed so that $\lambda_1 = 1 > |\lambda_2| \ge \cdots \ge |\lambda_K|$. Then for every integer lag $\tau \ge 1$,
\begin{equation*}
    \rho_{|G|}(\tau)
    \;=\; \sum_{k=2}^{K} w_k\, \lambda_k^{\tau},
    \qquad
    w_k \;=\; \frac{(\mathbf m^\top \mathbf D\, \mathbf v_k)\,(\mathbf w_k^\top \mathbf m)}{\sigma_{|G|}^2}.
\end{equation*}
\end{theorem}

\begin{proof}
The proof is organised in six steps.

\paragraph{Step 1 (Definitions).}
The autocorrelation of $|G_t|$ at lag $\tau \ge 0$ is, by stationarity,
\[
    \rho_{|G|}(\tau)
    \;=\; \frac{\mathrm{Cov}\!\left(|G_t|,\, |G_{t+\tau}|\right)}{\sigma_{|G|}^2}
    \;=\; \frac{\E[|G_t|\,|G_{t+\tau}|] - \mu^2}{\sigma_{|G|}^2}.
\]
The task reduces to evaluating $\E[|G_t|\,|G_{t+\tau}|]$ in closed form for $\tau \ge 1$.

\paragraph{Step 2 (Conditioning on the latent state pair).}
For $\tau \ge 1$, by the law of total expectation conditional on $(s_t, s_{t+\tau})$,
\[
    \E[|G_t|\,|G_{t+\tau}|]
    \;=\; \sum_{i,j=1}^K \Prob(s_t = i, s_{t+\tau} = j)\,
        \E\!\left[|G_t|\,|G_{t+\tau}| \,\big|\, s_t = i, s_{t+\tau} = j\right].
\]
The HMM emission structure gives the conditional independence $G_t \perp G_{t+\tau} \mid (s_t, s_{t+\tau})$ for $\tau \ge 1$: each emission depends only on its current state, and integrating the intermediate states $s_{t+1:t+\tau-1}$ out of the full-sequence factorisation $\big(\prod_{u} b_{s_u}(G_u)\big)\, p(s_{t+1:t+\tau-1} \mid s_t, s_{t+\tau})$ leaves the endpoint emissions decoupled. Hence
\[
    \E\!\left[|G_t|\,|G_{t+\tau}| \,\big|\, s_t = i, s_{t+\tau} = j\right]
    \;=\; \E[|G_t| \mid s_t = i]\,\E[|G_{t+\tau}| \mid s_{t+\tau} = j]
    \;=\; m_i\, m_j.
\]
The joint state probability factors by stationarity and the Markov property, $\Prob(s_t = i, s_{t+\tau} = j) = \bar\pi_i\, (\mathbf T^\tau)_{ij}$. Substituting and noting $[\mathbf D \mathbf T^\tau]_{ij} = \bar\pi_i (\mathbf T^\tau)_{ij}$,
\begin{equation}
    \E[|G_t|\,|G_{t+\tau}|] \;=\; \mathbf m^\top \mathbf D\, \mathbf T^\tau\, \mathbf m,
    \qquad \tau \ge 1.
    \label{eq:supp_bilinear}
\end{equation}

\paragraph{Step 3 (Spectral preliminaries).}
For irreducible aperiodic $\mathbf T$, Perron--Frobenius applied to row-stochastic matrices gives $\lambda_1 = 1$ as a simple eigenvalue with all other eigenvalues strictly inside the unit disk \citep{levinperes2017markov}. The right eigenvector for $\lambda_1$ is $\mathbf v_1 = \mathbf 1$, since $\mathbf T \mathbf 1 = \mathbf 1$ from the row-sum constraint, and the left eigenvector is $\mathbf w_1 = \bar{\boldsymbol\pi}$, since $\bar{\boldsymbol\pi}^\top \mathbf T = \bar{\boldsymbol\pi}^\top$ by definition of the stationary distribution. Biorthonormality $\mathbf w_1^\top \mathbf v_1 = \bar{\boldsymbol\pi}^\top \mathbf 1 = 1$ is automatic since $\bar{\boldsymbol\pi}$ is a probability vector.

\paragraph{Step 4 (Spectral decomposition of $\mathbf T^\tau$).}
Diagonalisability with biorthonormality gives the dyadic expansion
\[
    \mathbf T^\tau
    \;=\; \sum_{k=1}^K \lambda_k^\tau\, \mathbf v_k \mathbf w_k^\top
    \;=\; \mathbf 1\, \bar{\boldsymbol\pi}^\top \;+\; \sum_{k=2}^K \lambda_k^\tau\, \mathbf v_k \mathbf w_k^\top,
\]
isolating the dominant ($k = 1$) projection from the contractive spectrum. The dominant projector $\mathbf P_1 = \mathbf 1\bar{\boldsymbol\pi}^\top$ is the rank-one stationary projection: $\mathbf P_1 \mathbf x = (\bar{\boldsymbol\pi}^\top \mathbf x)\,\mathbf 1$.

\paragraph{Step 5 (Evaluation of the bilinear form).}
Substituting the spectral decomposition into~\eqref{eq:supp_bilinear},
\[
    \E[|G_t|\,|G_{t+\tau}|]
    \;=\; \mathbf m^\top \mathbf D\, \mathbf 1\, \bar{\boldsymbol\pi}^\top \mathbf m
    \;+\; \sum_{k=2}^K \lambda_k^\tau\, (\mathbf m^\top \mathbf D\, \mathbf v_k)\,(\mathbf w_k^\top \mathbf m).
\]
The $k = 1$ term simplifies to $\mu^2$: $\mathbf D \mathbf 1$ is the column vector $\bar{\boldsymbol\pi}$, so $\mathbf m^\top \mathbf D \mathbf 1 = \bar{\boldsymbol\pi}^\top \mathbf m = \mu$, and likewise $\bar{\boldsymbol\pi}^\top \mathbf m = \mu$, giving the product $\mu \cdot \mu = \mu^2$. Denoting the $k \ge 2$ scalar coefficients
\[
    c_k \;=\; (\mathbf m^\top \mathbf D\, \mathbf v_k)\,(\mathbf w_k^\top \mathbf m),
\]
we obtain
\[
    \E[|G_t|\,|G_{t+\tau}|] \;=\; \mu^2 \;+\; \sum_{k=2}^K c_k\, \lambda_k^\tau,
    \qquad \tau \ge 1.
\]

\paragraph{Step 6 (Autocovariance and autocorrelation).}
Subtracting $\mu^2$ from Step 1 cancels the dominant term, leaving
\[
    \mathrm{Cov}\!\left(|G_t|,\, |G_{t+\tau}|\right)
    \;=\; \sum_{k=2}^K c_k\, \lambda_k^\tau,
    \qquad \tau \ge 1,
\]
which is equation~\eqref{eq:acf_exact}. Dividing by $\sigma_{|G|}^2 = \gamma(0)$ gives $\rho_{|G|}(\tau) = \sum_{k \ge 2} w_k\, \lambda_k^\tau$ with $w_k = c_k / \sigma_{|G|}^2$, equation~\eqref{eq:acf_normalised}.
\end{proof}

\paragraph{Lag zero, the sum of weights, and the within-state white-noise contribution.}
The bilinear identity~\eqref{eq:supp_bilinear} fails at $\tau = 0$, where $G_t = G_{t+\tau}$ and the conditional product is the per-state second moment $M_i = \E[G_t^2 \mid s_t = i]$ rather than $m_i^2$. The eigenvalue completeness $\sum_{k=1}^K \mathbf v_k \mathbf w_k^\top = I$ instead gives the identity
\[
    \sum_{k=2}^K c_k
    \;=\; \mathbf m^\top \mathbf D\, \mathbf m \;-\; \mu^2
    \;=\; \sum_{i=1}^K \bar\pi_i\, m_i^2 \;-\; \mu^2
    \;=\; \mathrm{Var}\!\big(m_{s_t}\big),
\]
the between-state variance of $|G_t|$, equal to the variance of the conditional-mean random variable $m_{s_t}$ under $\bar{\boldsymbol\pi}$. By the law of total variance, $\sigma_{|G|}^2 = \mathrm{Var}(m_{s_t}) + \E[\mathrm{Var}(|G_t| \mid s_t)]$, so
\[
    \sum_{k=2}^K w_k
    \;=\; \frac{\mathrm{Var}(m_{s_t})}{\sigma_{|G|}^2}
    \;\in\; [0, 1],
\]
with equality to $1$ if and only if every per-state distribution is degenerate at $m_i$. The residual $1 - \sum_{k \ge 2} w_k = \E[\mathrm{Var}(|G_t| \mid s_t)]\, /\, \sigma_{|G|}^2$ is the within-state variance ratio: it contributes to $\sigma_{|G|}^2$ but not to $\mathrm{Cov}(|G_t|, |G_{t+\tau}|)$ at any positive lag, and is the source of the discontinuity from $\rho_{|G|}(0) = 1$ to $\rho_{|G|}(1) = \sum_{k \ge 2} w_k\, \lambda_k$ at the first lag. The substantive consequence for moderate-$K$ CHMM is unchanged: at every $\tau \ge 1$ the autocorrelation lies in the span of the geometric decays $\{\lambda_k^\tau\}_{k \ge 2}$, and the slow-decay envelope is governed by the largest non-unit eigenvalue.

\paragraph{Squared-return analogue.}
Replacing $|G_t|$ with $G_t^2$ and $\mathbf m$ with $\mathbf M = (M_1, \ldots, M_K)^\top$ in Steps~1--6 above is termwise valid for $\tau \ge 1$ (the conditional independence and stationarity arguments are unaffected), giving the squared-return analogue
\[
    \rho_{G^2}(\tau) \;=\; \sum_{k=2}^K \tilde w_k\, \lambda_k^\tau,
    \qquad
    \tilde w_k \;=\; \frac{(\mathbf M^\top \mathbf D\, \mathbf v_k)\,(\mathbf w_k^\top \mathbf M)}{\sigma_{G^2}^2}.
\]
The non-unit spectrum of $\mathbf T$ that determines $\rho_{|G|}(\tau)$ also determines $\rho_{G^2}(\tau)$; only the linear projection of the per-state moment vector onto the eigenmodes changes.

\paragraph{Remark (complex eigenvalues).}
For real $\mathbf T$ without self-adjoint structure, eigenvalues outside $\R$ occur in complex conjugate pairs with conjugate eigenvectors. A pair $(\lambda_k, \bar\lambda_k)$ at $\lambda_k = r e^{i\theta}$ contributes a real damped-oscillation term of the form $A\, r^\tau \cos(\theta\tau) + B\, r^\tau \sin(\theta\tau)$ to $\gamma(\tau)$, with $(A, B)$ determined from $(\mathrm{Re}\, c_k, \mathrm{Im}\, c_k)$. Theorem~\ref{thm:supp_spectral_acf} as stated remains valid; only the interpretation of the individual modes shifts from monotone geometric to damped sinusoidal. For $\mathbf T$ estimated on equity returns at the operating point of Section~\ref{sec:results}, the slow-decay mode at $\lambda_2 \approx \exp(-1/\tau^*)$ is real and positive, and the mono-exponential interpretation of the dominant timescale used in Section~\ref{sec:theory} carries through.

\paragraph{Remark (non-diagonalisable $\mathbf T$).}
Diagonalisability is generic: for $\mathbf T$ estimated by Baum--Welch in the interior of a non-degenerate parameter region, the eigenvalues are almost surely distinct, and the sample paths used in this paper realise this generic case at every $K$ in the sweep. Should $\mathbf T$ admit a non-trivial Jordan block of size $J$, the dyadic expansion of Step~4 acquires polynomial-in-$\tau$ prefactors $\binom{\tau}{j}\, \lambda_k^{\tau - j}$ for $j = 0, \ldots, J - 1$ on the off-diagonal blocks, and $\rho_{|G|}(\tau)$ becomes a finite sum of polynomial-times-geometric terms with the same exponential rate $\max_{k \ge 2} |\lambda_k|$. The substantive content of Theorem~\ref{thm:supp_spectral_acf}, that the absolute-return ACF is determined entirely by the non-unit spectrum of $\mathbf T$, is unchanged.

\subsection{State-Resolution and Multi-Emission Sensitivity}
\label{sec:supp_sensitivity}

This appendix reports the full $K$-sweep underlying the headline operating point of Section~\ref{sec:results}, the multi-emission family extension confirming that $K = 18$ is stable under Student-t and Laplace emissions, the EM convergence curves at representative $K$, the IS and OoS visual comparison panels at non-selected $K \in \{3, 6, 12, 21\}$, the multi-emission figure panels at $K = 18$, and the Ryd\'{e}n $K = 2$ replication.

\input{sections/sensitivity_appendix}

\subsection{Robustness, Test Power, Multi-Seed, and Auxiliary Baselines}
\label{sec:supp_robustness}

This appendix collects the robustness diagnostics and auxiliary baseline panel referenced from Section~\ref{sec:results}: the KS test-power calibration, block-bootstrap KS recalibration, bandwidth/bin-count and multi-seed sensitivity, the auxiliary block-bootstrap baseline, the QuantGAN deep-generative baseline, the per-ticker OoS price-simulation figures, and the expanded baseline panel that adds the GARCH family and the SM-CHMM foil to the headline table.

\input{sections/baselines_appendix}

\subsection{Cross-Asset Dependence: Full Pipeline~B Panel}
\label{sec:supp_cross_asset}

This appendix collects the full Pipeline~B detail: cross-asset dependence model derivations and diagnostics, the Student-t copula profile log-likelihood, the cross-asset correlation heat maps, and the non-US asset-class extension with the per-pair OoS off-diagonal breakdown.

\input{sections/cross_asset_appendix}

\subsection{\texorpdfstring{Pre-OoS Validation $K$-Selection and $\nu_k$ Diagnostics}{Pre-OoS Validation K-Selection and nu\_k Diagnostics}}
\label{sec:supp_misc}

\paragraph{Pre-OoS validation $K$-selection.}
The $K = 18$ operating point in the main panel was chosen via a multi-objective criterion combining information-criterion (IC) rank with IS and OoS distributional pass rates, which uses the same $2024$--$2026$ window later used for VaR. As a clean re-selection that decouples the two, we fit CHMM-N on $2014$-$01$-$03$ through $2021$-$12$-$31$ ($n = 2{,}013$) and score every $K$ in the sweep by the forward-algorithm held-out log-likelihood on $2022$-$01$-$03$ through $2024$-$01$-$03$ ($n = 502$); the $2024$--$2026$ window is not touched. Held-out log-lik, BIC, HQC, and CAIC all select $K^\star = 3$ (per-observation log-lik $-2.5345$ at $K = 3$ vs.\ $-2.5694$ at $K = 18$); AIC selects $K^\star = 6$; held-out two-sample KS selects $K^\star = 9$. None of the six criteria that avoid the OoS window select $K = 18$. The $K = 18$ choice is best read as a multi-metric one that weights tail fidelity (kurtosis, AD) and downstream VaR behaviour alongside likelihood; a parallel main panel at $K = 3$ would re-rank some rows in CHMM's favour on ACF-MAE and against it on tail fidelity. The $2022$--$2023$ validation slice is itself partly a structural-break period (the rate-hike cycle), so any model with substantial IS-specific structure generalises worse on it.

\paragraph{Pre-2020 held-out $K$-selection (no rate-hike confound).}
Reviewer R1's Question 1 asked whether a strictly pre-2020 validation slice (avoiding the rate-hike confound that contaminates the $2022$-$2023$ slice above) selects the same $K^\star$. We re-ran the held-out procedure on a $2014$-$01$-$03$ through $2018$-$06$-$29$ estimation window ($n_{\text{est}} = 1{,}130$) with a $2018$-$07$-$02$ through $2019$-$12$-$31$ validation slice ($n_{\text{val}} = 377$), both fully pre-COVID and pre-2022-rate-hike (the Q4 2018 drawdown and 2019 recovery sit inside the validation window as a moderate-stress non-regime-shift event). Per-observation held-out log-likelihood: $-2.0681$ at $K = 3$, $-2.0606$ at $K = 6$, $-2.0763$ at $K = 9$, $-2.1287$ at $K = 12$, $-2.1921$ at $K = 18$, $-2.2609$ at $K = 21$. \textbf{The held-out log-likelihood selects $K^\star = 6$, and the held-out two-sample KS pass rate also selects $K^\star = 6$ ($96.8\%$ at $K = 6$ versus $84.6\%$ at $K = 3$).} AIC, BIC, HQC, and CAIC continue to select $K^\star = 3$ on this slice, reflecting the parameter-count penalty on a $4.5$-year estimation window where the per-state sample is small. The qualitative reading is robust to the slice choice: held-out distributional fidelity rises from $K = 3$ to a peak in the $K = 6$ to $K = 9$ band and then degrades, with neither slice selecting $K = 18$ by any held-out criterion, but the pre-2020 result favours a moderately larger $K$ ($6$ vs $3$) than the rate-hike-contaminated $2022$-$2023$ slice.

\paragraph{Penalised-ECM rate sweep.}
The bracket sensitivity is reported with its table in Section~\ref{sec:nu_diagnostics}; here we report the complementary $1/\nu_k$ shrinkage rate sweep referenced in Section~\ref{sec:discussion}. An exponential $1/\nu_k$ prior at rate $\lambda \in \{0, 5, 20, 50, 100, 200\}$ reduces simulated excess kurtosis monotonically: $14.30, 11.19, 8.43, 5.41, 3.96, 3.67$. The recommended rate $\lambda = 20$ brings simulated kurtosis to $\approx 8$ (close to the observed $7.69$) at a $1$pp IS KS cost. Only the $\nu_{\min}$ state parameter actually moves ($2.1 \to 4.89$); upper bracket and median remain at $50$. The overshoot is therefore a single-state artefact of the two tail regimes hitting the lower bracket under unpenalised ECM.

\paragraph{Walk-forward / rolling-origin OoS for CHMM-N.}
We define six rolling-origin folds, each train $5$ years and test $1$ year, refit CHMM-N at $K \in \{3, 18\}$ from scratch on each fold's train slice, and report KS pass rate, simulated kurtosis, and $|G_t|$ ACF-MAE on the fold's test slice ($N_{\text{paths}} = 500$). Across the six folds (Table~\ref{tab:walkforward}), CHMM-N at $K = 18$ attains median (IQR) KS $67.7\%\,[8.2, 75.0]$ and $|G_t|$ ACF-MAE $0.0542\,[0.0508, 0.0571]$; CHMM-N at $K = 3$ attains $62.1\%\,[7.2, 78.4]$ and $0.0563\,[0.0552, 0.0592]$. Two folds carry sharply lower KS pass rates: W2 (COVID test slice, $7$--$8\%$) and W4 (rate-hike test slice, $0$--$1\%$). The four non-stress folds attain KS $61$--$83\%$, consistent with the headline OoS pass rate. The headline ranking (CHMM-N at $K = 18$ $\geq$ CHMM-N at $K^\star = 3$ on KS, both within $0.005$ of each other on $|G_t|$ ACF-MAE) is window-robust on the four non-stress folds; on stress folds (COVID, rate hike) all CHMM rows drop sharply, reflecting the stationarity-scope limit rather than a model-selection issue. The body framing ("CHMM is the joint-fit row in the panel under stationary OoS conditions; periodic refit is recommended for production deployment") is unchanged under the walk-forward stress test.

\begin{table}[h]
\centering
\small
\caption{\textbf{Walk-forward / rolling-origin OoS for CHMM-N}. Six folds, each train 5 years and test 1 year; $N_{\text{paths}} = 500$ per fold. KS pass rate at $\alpha = 0.05$. W2 covers the COVID drawdown; W4 covers the start of the 2022 rate-hike cycle. The bracketed pairs in the bottom row report $[Q_1, Q_3]$ across the six folds (the empirical first and third quartiles, not a scalar IQR width).}
\label{tab:walkforward}
\begin{tabular}{l c c c c c c}
\toprule
& \multicolumn{2}{c}{KS (\%)} & \multicolumn{2}{c}{Kurt sim} & \multicolumn{2}{c}{ACF-MAE $|G_t|$} \\
\cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7}
Fold (test window) & $K = 3$ & $K = 18$ & $K = 3$ & $K = 18$ & $K = 3$ & $K = 18$ \\
\midrule
W1 (2019)              & $63.2$ & $82.6$ & $2.51$ & $2.66$ & $0.0572$ & $0.0525$ \\
W2 (2020 COVID)        & $\phantom{0}7.2$ & $\phantom{0}8.2$ & $2.42$ & $2.50$ & $0.0552$ & $0.0584$ \\
W3 (2021)              & $82.4$ & $74.0$ & $5.73$ & $6.76$ & $0.0602$ & $0.0571$ \\
W4 (2022 rate hike)    & $\phantom{0}0.8$ & $\phantom{0}0.0$ & $4.72$ & $4.35$ & $0.0592$ & $0.0558$ \\
W5 (2023)              & $78.4$ & $75.0$ & $3.47$ & $2.35$ & $0.0511$ & $0.0508$ \\
W6 (2024)              & $61.0$ & $61.4$ & $2.81$ & $2.86$ & $0.0555$ & $0.0493$ \\
\midrule
median                 & $62.1$ & $67.7$ & $3.14$ & $2.76$ & $0.0563$ & $0.0542$ \\
IQR                    & $[7.2, 78.4]$ & $[8.2, 75.0]$ & $[2.51, 4.72]$ & $[2.50, 4.35]$ & $[0.0552, 0.0592]$ & $[0.0508, 0.0571]$ \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Four-family conditional VaR back-test.}
Extending the regime-conditional VaR construction of \S\ref{sec:var_backtest} to all four emission families (CHMM-N, CHMM-t at $\lambda = 20$, CHMM-L, CHMM-GED) at $K \in \{3, 18\}$ and $\alpha \in \{0.01, 0.05\}$ yields sixteen $(K, \alpha, \text{family})$ rows on the OoS window (Table~\ref{tab:cond_var_all_families}). Every row passes Kupiec, Christoffersen-ind, and Christoffersen-cc cleanly: $p_{\text{cc}} \ge 0.089$ throughout, with all $\text{LR}_{\text{cc}}$ statistics below the $\chi^2_2(0.05) = 5.991$ critical value. The regime-switching value proposition is therefore intrinsic to the latent-state forecast and not specific to the Gaussian emission family.

\begin{table}[h]
\centering
\small
\caption{\textbf{Regime-conditional VaR back-test across four CHMM emission families} on SPY OoS ($T_{\text{OoS}} = 572$, seed $20260420$). Critical values: $\chi^2_1(0.05) = 3.841$, $\chi^2_2(0.05) = 5.991$. Every $(K, \alpha, \text{family})$ row passes Kupiec, Christoffersen-ind, and Christoffersen-cc at $\alpha = 0.05$.}
\label{tab:cond_var_all_families}
\begin{tabular}{l c c r r r r r r}
\toprule
Family & $K$ & $\alpha$ & breaches & br rate & median $\widehat{\text{VaR}}_t$ & $\text{LR}_{\text{uc}}$ & $\text{LR}_{\text{ind}}$ & $\text{LR}_{\text{cc}}$ \\
\midrule
CHMM-N             & $3$  & $0.01$ & $9$  & $1.57\%$ & $-4.56$ & $1.62$ & $2.36$ & $3.98$ \\
CHMM-N             & $3$  & $0.05$ & $35$ & $6.12\%$ & $-2.87$ & $1.41$ & $0.01$ & $1.42$ \\
CHMM-N             & $18$ & $0.01$ & $9$  & $1.57\%$ & $-5.20$ & $1.62$ & $2.36$ & $3.98$ \\
CHMM-N             & $18$ & $0.05$ & $26$ & $4.55\%$ & $-3.02$ & $0.26$ & $0.52$ & $0.78$ \\
CHMM-t ($\lambda=20$) & $3$  & $0.01$ & $8$  & $1.40\%$ & $-5.59$ & $0.82$ & $2.81$ & $3.63$ \\
CHMM-t ($\lambda=20$) & $3$  & $0.05$ & $32$ & $5.59\%$ & $-2.73$ & $0.41$ & $0.77$ & $1.19$ \\
CHMM-t ($\lambda=20$) & $18$ & $0.01$ & $10$ & $1.75\%$ & $-6.25$ & $2.65$ & $1.98$ & $4.62$ \\
CHMM-t ($\lambda=20$) & $18$ & $0.05$ & $29$ & $5.07\%$ & $-3.15$ & $0.01$ & $1.39$ & $1.40$ \\
CHMM-L             & $3$  & $0.01$ & $5$  & $0.87\%$ & $-5.45$ & $0.10$ & $4.74$ & $4.84$ \\
CHMM-L             & $3$  & $0.05$ & $26$ & $4.55\%$ & $-2.63$ & $0.26$ & $2.23$ & $2.48$ \\
CHMM-L             & $18$ & $0.01$ & $9$  & $1.57\%$ & $-6.18$ & $1.62$ & $0.29$ & $1.91$ \\
CHMM-L             & $18$ & $0.05$ & $30$ & $5.24\%$ & $-2.80$ & $0.07$ & $1.16$ & $1.23$ \\
CHMM-GED           & $3$  & $0.01$ & $7$  & $1.22\%$ & $-5.73$ & $0.27$ & $3.34$ & $3.61$ \\
CHMM-GED           & $3$  & $0.05$ & $31$ & $5.42\%$ & $-2.54$ & $0.21$ & $0.96$ & $1.16$ \\
CHMM-GED           & $18$ & $0.01$ & $9$  & $1.57\%$ & $-6.55$ & $1.62$ & $2.36$ & $3.98$ \\
CHMM-GED           & $18$ & $0.05$ & $25$ & $4.37\%$ & $-3.10$ & $0.50$ & $2.56$ & $3.06$ \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Walk-forward conditional-VaR back-test.}
Extending the regime-conditional VaR construction to the six rolling-origin folds of Table~\ref{tab:walkforward}, with CHMM-N refit per fold from scratch on the train slice and forward-filtered through the test slice under fold-IS-fixed parameters, gives twenty-four $(\text{fold}, K, \alpha)$ rows in Table~\ref{tab:walkforward_cond_var}.
The conditional Christoffersen-cc passes at $\alpha = 0.05$ on ten of twelve fold-$K$ combinations (W1, W3, W4, W5, W6 at both $K \in \{3, 18\}$); the two failures concentrate on W2 (COVID, $p_{\text{cc}} \in \{0.011, 0.002\}$ at $K \in \{3, 18\}$).
At $\alpha = 0.01$ the construction passes on nine of twelve combinations: W2 fails at both $K$ ($p_{\text{cc}} < 10^{-3}$) and W4 at $K = 18$ fails at $p_{\text{cc}} = 0.022$ on a $\text{LR}_{\text{ind}}$ tail-event clustering rather than a $\text{LR}_{\text{uc}}$ coverage miss ($\text{LR}_{\text{ind}} = 7.49$ at $\text{LR}_{\text{uc}} = 0.11$).
The pass-rate is therefore $19/24$ aggregate at the $5\%$ level and $22/24$ if we exclude the two stress folds that the univariate walk-forward already flagged as out-of-distribution by KS in Table~\ref{tab:walkforward}.
The reading is that the conditional construction is window-conditional: under regime shifts of the W2 / W4 magnitude (a regime introduction the IS distribution does not span; Section~\ref{sec:discussion}, stationarity-scope paragraph), the conditional VaR inherits the same out-of-scope behaviour as the unconditional generators.
On the four non-stress folds (W1, W3, W5, W6) plus W4 at $K = 3$ at the $5\%$ level the conditional construction is window-robust, supporting the body framing of \S\ref{sec:var_backtest} extending beyond the headline single-window result.

\begin{table}[h]
\centering
\small
\caption{\textbf{Walk-forward regime-conditional VaR back-test for CHMM-N} on six rolling-origin folds (train 5y / test 1y; seed $20260420$). At each test day $t$, conditional $\widehat{\text{VaR}}_t(\alpha)$ is the $\alpha$-quantile of the $K$-component Gaussian mixture predictive density under fold-IS-fixed parameters, evaluated on the full-timeline filter $\Pr(s_{t+1} \mid \mathcal F_t = \text{train} \cup \text{test}[1{:}t-1])$. Critical values: $\chi^2_1(0.05) = 3.841$, $\chi^2_2(0.05) = 5.991$. Bold $p_{\text{cc}}$ rows fail at the $5\%$ level.}
\label{tab:walkforward_cond_var}
\begin{tabular}{l c c r r r r r r r}
\toprule
Fold & $K$ & $\alpha$ & breaches & br rate & median $\widehat{\text{VaR}}_t$ & $\text{LR}_{\text{uc}}$ & $\text{LR}_{\text{ind}}$ & $\text{LR}_{\text{cc}}$ & $p_{\text{cc}}$ \\
\midrule
W1 & $3$  & $0.01$ & $\phantom{0}2$ & $0.80\%$ & $-4.01$ & $\phantom{0}0.11$ & $0.03$ & $\phantom{0}0.15$ & $0.93$ \\
W1 & $3$  & $0.05$ & $15$           & $5.98\%$ & $-2.43$ & $\phantom{0}0.48$ & $1.18$ & $\phantom{0}1.65$ & $0.44$ \\
W1 & $18$ & $0.01$ & $\phantom{0}0$ & $0.00\%$ & $-5.03$ & $\phantom{0}5.05$ & $0.00$ & $\phantom{0}5.05$ & $0.08$ \\
W1 & $18$ & $0.05$ & $12$           & $4.78\%$ & $-2.66$ & $\phantom{0}0.03$ & $0.38$ & $\phantom{0}0.40$ & $0.82$ \\
W2 & $3$  & $0.01$ & $16$           & $6.35\%$ & $-6.54$ & $32.93$ & $0.87$ & $33.80$ & $\mathbf{0.00}$ \\
W2 & $3$  & $0.05$ & $23$           & $9.13\%$ & $-4.59$ & $\phantom{0}7.34$ & $1.71$ & $\phantom{0}9.05$ & $\mathbf{0.01}$ \\
W2 & $18$ & $0.01$ & $14$           & $5.56\%$ & $-6.67$ & $25.59$ & $1.56$ & $27.15$ & $\mathbf{0.00}$ \\
W2 & $18$ & $0.05$ & $25$           & $9.92\%$ & $-4.54$ & $10.11$ & $2.56$ & $12.68$ & $\mathbf{0.00}$ \\
W3 & $3$  & $0.01$ & $\phantom{0}6$ & $2.39\%$ & $-4.48$ & $\phantom{0}3.53$ & $0.30$ & $\phantom{0}3.82$ & $0.15$ \\
W3 & $3$  & $0.05$ & $20$           & $7.97\%$ & $-2.47$ & $\phantom{0}3.98$ & $0.30$ & $\phantom{0}4.28$ & $0.12$ \\
W3 & $18$ & $0.01$ & $\phantom{0}2$ & $0.80\%$ & $-5.57$ & $\phantom{0}0.11$ & $0.03$ & $\phantom{0}0.15$ & $0.93$ \\
W3 & $18$ & $0.05$ & $15$           & $5.98\%$ & $-3.18$ & $\phantom{0}0.48$ & $1.18$ & $\phantom{0}1.65$ & $0.44$ \\
W4 & $3$  & $0.01$ & $\phantom{0}5$ & $2.00\%$ & $-7.98$ & $\phantom{0}1.96$ & $0.21$ & $\phantom{0}2.16$ & $0.34$ \\
W4 & $3$  & $0.05$ & $18$           & $7.20\%$ & $-4.16$ & $\phantom{0}2.26$ & $1.99$ & $\phantom{0}4.24$ & $0.12$ \\
W4 & $18$ & $0.01$ & $\phantom{0}2$ & $0.80\%$ & $-9.62$ & $\phantom{0}0.11$ & $7.49$ & $\phantom{0}7.60$ & $\mathbf{0.02}$ \\
W4 & $18$ & $0.05$ & $10$           & $4.00\%$ & $-6.63$ & $\phantom{0}0.56$ & $3.80$ & $\phantom{0}4.36$ & $0.11$ \\
W5 & $3$  & $0.01$ & $\phantom{0}0$ & $0.00\%$ & $-6.06$ & $\phantom{0}5.01$ & $0.00$ & $\phantom{0}5.01$ & $0.08$ \\
W5 & $3$  & $0.05$ & $\phantom{0}8$ & $3.21\%$ & $-3.90$ & $\phantom{0}1.91$ & $0.53$ & $\phantom{0}2.44$ & $0.30$ \\
W5 & $18$ & $0.01$ & $\phantom{0}1$ & $0.40\%$ & $-7.10$ & $\phantom{0}1.16$ & $0.01$ & $\phantom{0}1.17$ & $0.56$ \\
W5 & $18$ & $0.05$ & $\phantom{0}7$ & $2.81\%$ & $-3.85$ & $\phantom{0}2.96$ & $0.41$ & $\phantom{0}3.37$ & $0.19$ \\
W6 & $3$  & $0.01$ & $\phantom{0}2$ & $0.80\%$ & $-4.68$ & $\phantom{0}0.11$ & $0.03$ & $\phantom{0}0.15$ & $0.93$ \\
W6 & $3$  & $0.05$ & $13$           & $5.18\%$ & $-2.85$ & $\phantom{0}0.02$ & $0.15$ & $\phantom{0}0.17$ & $0.92$ \\
W6 & $18$ & $0.01$ & $\phantom{0}1$ & $0.40\%$ & $-5.83$ & $\phantom{0}1.19$ & $0.01$ & $\phantom{0}1.20$ & $0.55$ \\
W6 & $18$ & $0.05$ & $\phantom{0}7$ & $2.79\%$ & $-2.98$ & $\phantom{0}3.06$ & $0.40$ & $\phantom{0}3.46$ & $0.18$ \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Per-ticker $\hat\lambda^\star$ sweep.}
A per-ticker sweep $\lambda \in \{0, 5, 10, 20, 50, 100\}$ on each of the six body tickers under the constraint that IS KS at $\lambda$ exceeds the $\lambda = 0$ baseline by no more than $1.5$ percentage points (Table~\ref{tab:per_ticker_lambda}) shows that the body uniform $\lambda = 20$ is the right rate on the heavy-tailed defensives (NVDA, JNJ) but is over-shrunk on the moderate-tail tickers. The per-ticker optima are $\lambda^\star_{\text{SPY}} = 10$, $\lambda^\star_{\text{NVDA}} = 20$, $\lambda^\star_{\text{JNJ}} = 20$, $\lambda^\star_{\text{JPM}} = 10$, $\lambda^\star_{\text{AAPL}} = 10$, $\lambda^\star_{\text{QQQ}} = 0$. The honest operational reading is that $\lambda = 20$ is a reasonable uniform default for production deployment but per-ticker tuning is recommended whenever a ticker's residual kurtosis at $\lambda = 0$ is within $\sim 1$ unit of observed.

\begin{table}[h]
\centering
\small
\caption{\textbf{Per-ticker $\hat\lambda^\star$ for the penalised CHMM-t at $K = 18$} ($N_{\text{paths}} = 500$, seed $20260420$, KS-degradation tolerance $\le 1.5$pp from $\lambda = 0$). $\hat\lambda^\star$ is the value of $\lambda$ in $\{0, 5, 10, 20, 50, 100\}$ that minimises $|{\text{kurt}}_{\text{sim}} - {\text{kurt}}_{\text{obs}}|$ subject to the KS-degradation constraint.}
\label{tab:per_ticker_lambda}
\begin{tabular}{l c c c c}
\toprule
Ticker & $\text{kurt}_{\text{obs}}$ & $\hat\lambda^\star$ & $\text{kurt}_{\text{sim}}(\lambda^\star)$ & IS KS$(\lambda^\star)$ \\
\midrule
SPY  & $7.69$  & $10$ & $11.82$ & $96.4\%$ \\
NVDA & $5.54$  & $20$ & $5.68$  & $99.4\%$ \\
JNJ  & $8.98$  & $20$ & $10.74$ & $99.8\%$ \\
JPM  & $7.63$  & $10$ & $8.62$  & $98.6\%$ \\
AAPL & $3.20$  & $10$ & $3.24$  & $99.2\%$ \\
QQQ  & $3.40$  & $0$  & $3.35$  & $95.2\%$ \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Cross-asset summary.}
The full per-construction Pipeline~B panel is in Section~\ref{sec:supp_cross_asset}; for reference, the headline ordering is summarised in Table~\ref{tab:cross_asset_supp_summary}.

\begin{table}[h]
\centering
\small
\caption{\textbf{Cross-asset dependence summary} (Pipeline~B; CHMM-N marginals at $K = 18$; $200$ paths; seed \texttt{20260422}). Median per-asset KS pass rate (\%) and off-diagonal MAE of the simulated correlation matrix vs.\ observed. Bold marks the column winner. The Student-t copula at $\nu^\star = 6$ is the strongest dependence layer on the two IS columns; on OoS the Gaussian copula attains marginally lower off-diagonal MAE ($0.202$ vs.\ $0.209$) while the truncated C-vine sits between the elliptical copulas and the SIM baseline. Full per-asset detail is in Section~\ref{sec:supp_cross_asset}; the C-vine edge summary is in Table~\ref{tab:cvine_edges}.}
\label{tab:cross_asset_supp_summary}
\resizebox{\textwidth}{!}{%
\begin{tabular}{l c c c c}
\toprule
& Median IS KS & Median OoS KS & Off-diag MAE IS & Off-diag MAE OoS \\
\midrule
Single Index Model       & $77.0$  & $87.0$  & $0.076$ & $0.252$ \\
Gaussian copula          & $97.8$  & $\mathbf{87.5}$  & $0.030$ & $\mathbf{0.202}$ \\
Student-t copula ($\nu^\star = 6$) & $\mathbf{98.8}$  & $85.8$  & $\mathbf{0.027}$ & $0.209$ \\
Truncated C-vine (root SPY) & $97.2$  & $86.0$  & $0.068$ & $0.235$ \\
\bottomrule
\end{tabular}%
}
\end{table}
