\subsection{Utility: VaR and Expected-Shortfall Back-Test}
\label{sec:utility}

Fidelity (Section~\ref{sec:model_comparison}) measures how close the simulated distribution is to the observed one at the univariate level.
Utility measures whether a downstream consumer of the synthetic paths arrives at the same quantitative conclusion that the historical data would produce.
For a financial-risk consumer the canonical diagnostics are one-tail Value-at-Risk (VaR) and Expected-Shortfall (ES) at daily probability levels of $1\%$ and $5\%$; the synthetic-data generator is well calibrated if the envelope of per-path simulated VaR/ES brackets the observed historical VaR/ES.
We simulate $1{,}000$ independent paths of length $T_{\text{IS}} = 2{,}516$ and $T_{\text{OoS}} = 572$ from each of the five models most relevant to the comparison (Bootstrap, GARCH(1,1), CHMM-N, CHMM-t, CHMM-L) and, on each path, compute the left-tail $1\%$ and $5\%$ VaR and the associated ES.
Table~\ref{tab:var_es} reports the median and $[5\%, 95\%]$ envelope of each quantity across paths alongside the observed historical value.

\begin{table}[t]
\centering
\caption{\textbf{VaR and Expected-Shortfall back-test calibration for SPY} (\NPaths{1{,}000}; global seed policy in Section~\ref{sec:metrics}). Each entry is the median and $[5\%, 95\%]$ envelope across paths of the left-tail VaR / ES at $\alpha \in \{0.01, 0.05\}$; observed historical values are in the first row. Headline: all three CHMM variants bracket the observed historical VaR and ES at both $1\%$ and $5\%$ on both IS and OoS windows, and the CHMM-N medians (IS VaR$_{0.01} = -6.64$, ES$_{0.01} = -8.80$) are the closest point estimates to the observed $-6.38$ and $-9.00$ among the reported generators.}
\label{tab:var_es}
\small
\resizebox{\textwidth}{!}{%
\begin{tabular}{l l l l l}
\toprule
          & IS VaR$_{0.01}$              & IS ES$_{0.01}$               & IS VaR$_{0.05}$              & IS ES$_{0.05}$               \\
Model     & median $[5\%, 95\%]$         & median $[5\%, 95\%]$         & median $[5\%, 95\%]$         & median $[5\%, 95\%]$         \\
\midrule
\textit{Observed}  & $-6.38$                      & $-9.00$                      & $-3.43$                      & $-5.50$                      \\
Bootstrap & $-6.38\ [-7.04, -5.99]$      & $-8.86\ [-10.37,\ -7.78]$    & $-3.43\ [-3.68, -3.20]$      & $-5.46\ [-5.99, -5.05]$      \\
GARCH     & $-5.80\ [-8.62, -4.52]$      & $-7.86\ [-13.54,\ -5.77]$    & $-3.19\ [-3.92, -2.71]$      & $-4.89\ [-7.11, -3.91]$      \\
CHMM-N    & $-6.64\ [-8.01, -5.47]$      & $-8.80\ [-10.16,\ -7.34]$    & $-3.45\ [-4.05, -2.95]$      & $-5.45\ [-6.33, -4.61]$      \\
CHMM-t    & $-6.58\ [-7.51, -5.83]$      & $-9.06\ [-10.70,\ -7.79]$    & $-3.40\ [-3.90, -2.90]$      & $-5.51\ [-6.18, -4.85]$      \\
CHMM-L    & $-6.89\ [-7.85, -6.01]$      & $-9.15\ [-10.42,\ -7.97]$    & $-3.30\ [-3.77, -2.88]$      & $-5.53\ [-6.17, -4.87]$      \\
\midrule
          & OoS VaR$_{0.01}$             & OoS ES$_{0.01}$              & OoS VaR$_{0.05}$             & OoS ES$_{0.05}$              \\
\textit{Observed}  & $-5.51$                      & $-7.94$                      & $-2.77$                      & $-4.67$                      \\
Bootstrap & $-6.33\ [-7.60, -5.28]$      & $-8.43\ [-11.76,\ -6.59]$    & $-3.39\ [-3.91, -2.88]$      & $-5.34\ [-6.38, -4.50]$      \\
GARCH     & $-5.35\ [-9.98, -3.61]$      & $-6.72\ [-13.51,\ -4.38]$    & $-3.14\ [-4.94, -2.36]$      & $-4.59\ [-7.92, -3.23]$      \\
CHMM-N    & $-6.30\ [-9.02, -4.30]$      & $-8.44\ [-11.36,\ -5.42]$    & $-3.39\ [-4.74, -2.47]$      & $-5.27\ [-7.23, -3.74]$      \\
CHMM-t    & $-6.32\ [-8.13, -4.80]$      & $-8.60\ [-12.20,\ -6.29]$    & $-3.29\ [-4.27, -2.45]$      & $-5.34\ [-6.72, -4.08]$      \\
CHMM-L    & $-6.69\ [-8.59, -4.91]$      & $-8.88\ [-11.48,\ -6.73]$    & $-3.27\ [-4.30, -2.51]$      & $-5.47\ [-6.82, -4.20]$      \\
\bottomrule
\end{tabular}%
}
\end{table}

Two findings are central to the utility story.
First, all three CHMM variants bracket the observed historical VaR and ES at both $1\%$ and $5\%$ levels within their $5\%$--$95\%$ envelopes on both IS and OoS windows; the IS CHMM-N median VaR$_{0.01}$ at $-6.64$ (observed $-6.38$) and the CHMM-N median ES$_{0.01}$ at $-8.80$ (observed $-9.00$) are the closest point estimates across the five generators.
IS the three CHMMs are therefore well calibrated as risk consumers and carry narrower uncertainty bands than GARCH (whose $[-8.62, -4.52]$ VaR$_{0.01}$ envelope is roughly $1.7\times$ wider than the CHMMs').
Second, CHMM-t -- despite overshooting the empirical kurtosis by $\sim 90\%$ IS -- does \emph{not} produce a systematically more conservative VaR/ES estimate: its IS median VaR$_{0.01}$ of $-6.58$ is almost identical to CHMM-N's $-6.64$, and its envelope width is comparable.
The CHMM-t kurtosis overshoot therefore translates into a modestly widened uncertainty band rather than a biased point estimate, so a practitioner using CHMM-t for tail-risk scenario generation does not systematically overstate $1\%$ or $5\%$ quantile risk.
A useful OoS check: the $2024$--$2026$ observed VaR$_{0.01}$ is $-5.51$ (a relatively calm tail compared to the IS average), and the three CHMMs bracket this value comfortably within their envelopes rather than over- or under-stating it.
Figure~\ref{fig:var_es} (Appendix~\ref{sec:var_es_supp}) visualizes the envelopes alongside the historical observations.

\paragraph{Explicit Kupiec failure of the Discrete HMM baseline of \citet{alswaidan2026hybrid} at $5\%$ VaR.}
\label{par:discrete_var_failure}%
Table~\ref{tab:var_es} shows envelope coverage for the five continuous and non-parametric generators, but the discrete HMM baselines of \citet{alswaidan2026hybrid} are excluded from the envelope panel because their bin-centroid emission produces a point-mass VaR estimator whose envelope width is mechanically zero.
The Kupiec and Christoffersen likelihood-ratio tests of Section~\ref{sec:extended_evaluation} are the appropriate diagnostic for those baselines, and they deliver a clear negative finding.
Both Discrete NJ and Discrete WJ produce an OoS $5\%$ VaR breach rate of exactly $1.7\%$, roughly one-third of the $5\%$ target; the Kupiec likelihood-ratio statistic $\text{LR}_{\text{uc}} = 16.81$ rejects unconditional coverage with $p < 10^{-3}$, and the independence statistic $\text{LR}_{\text{ind}} = 13.22$ also rejects.
The Poisson-jump augmentation does not remedy the failure: no-jump and with-jump variants produce the same breach rate to three significant figures because the prior-paper jump intensity $\epsilon = 5 \times 10^{-5}$ and mean duration $\lambda = 67$ trigger on fewer than one time step per IS or OoS window at our sample lengths, leaving the left-tail quantile of the bin-centroid emission unchanged between NJ and WJ.
A risk manager who uses the prior-paper discrete generator to size a $5\%$ VaR position is therefore committed to holding materially more capital than the empirical breach process requires and would fail a standard VaR back-test if audited.
The continuous CHMM family of the present paper sits at or below the $\chi^2_1$ Kupiec threshold at the $5\%$ level on the unconditional A8 panel (all three variants attain $\text{LR}_{\text{uc}} \le 3.83$, with CHMM-N, CHMM-t, and CHMM-L at $3.83$, $3.83$, and $3.03$ respectively), and the smoother-based regime-conditional VaR of Section~\ref{sec:conditional_var} additionally sits below the $\chi^2_1$ Christoffersen threshold as an IS goodness-of-fit diagnostic; the operationally correct filter-based one-step-ahead backtest of Section~\ref{sec:conditional_var} fails Kupiec, a finding we flag as evidence of regime-uncertainty-driven over-conservatism rather than as a clean operational ranking.
This is the clearest single-number contrast the present paper offers against the prior discrete-HMM pipeline.

\subsection{Semi-Markov Ablation: Heavy-Tailed Sojourn Distributions}
\label{sec:sm_ablation}

The flat CHMM assumes geometric sojourn times in every state, an implicit by-product of the Markov transition matrix.
Empirically, equity returns exhibit crisis-regime sojourn tails that are heavier than geometric: the bottom-$3$ states of the CHMM partition, which hold the March $2020$ crash and the $2022$ rate-cycle stress, persist longer in the Viterbi decode than a geometric sojourn would predict.
To quantify the effect of relaxing the geometric-sojourn assumption we fit a semi-Markov variant of each CHMM emission family using the explicit-duration forward-backward machinery of \citet{yu2010hidden}.
Let $\mathrm{SM\text{-}CHMM\text{-}}F$ denote the semi-Markov extension of the corresponding flat CHMM-$F$ for $F \in \{\mathcal{N}, t, L\}$.

\paragraph{Plug-in estimator and per-state sojourn family selection.}
We fit the SM variants with the plug-in estimator: given the flat CHMM transition matrix and emission parameters, we (i) decode the most-likely state sequence under Viterbi, (ii) extract the empirical run-length distribution for each state, and (iii) fit one of a negative-binomial, geometric, or discrete truncated Pareto to each state's run-lengths by AIC, after which the SM generator samples sojourn times from the fitted family.
The AIC-selected family per state is reported in Table~\ref{tab:sm_sojourn}: for all three emission families, $17$ of the $18$ states select a Pareto sojourn and $1$ selects a negative binomial.
This is consistent with preliminary results we report for the CBOE Volatility Index in a companion paper in preparation \citep{alswaidan2026smchmm}, where Pareto sojourns dominated on daily log-$\mathrm{VIX}$: the heavy-tailed sojourn pattern is not specific to vol-index series but holds on equity returns at moderate state resolution.

\begin{table}[t]
\centering
\small
\caption{\textbf{Per-state sojourn family selected by AIC} for each SM-CHMM variant on SPY at $K = 18$; candidate families are truncated Pareto, negative binomial, and geometric. Headline: $17$ of $18$ states pick Pareto sojourns under all three emission families (Gaussian / Student-t / Laplace), one state picks negative binomial each, and no state picks geometric, confirming that a heavy-tailed sojourn layer on top of the Markov chain is empirically warranted for equity-return regime persistence.}
\label{tab:sm_sojourn}
\begin{tabular}{lccc}
\toprule
Model & Pareto states & Negative-binomial states & Geometric states \\
\midrule
SM-CHMM-N & $17$ & $1$ & $0$ \\
SM-CHMM-t & $17$ & $1$ & $0$ \\
SM-CHMM-L & $17$ & $1$ & $0$ \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Risk-calibration gain: cleaner unconditional $1\%$ and $5\%$ VaR Kupiec.}
Table~\ref{tab:sm_var} reproduces the Kupiec and Christoffersen LR tests on the OoS window for the three flat CHMM variants and their SM counterparts.
At $1\%$ VaR, SM-CHMM-N drops $\text{LR}_{\text{uc}}$ from $1.58$ (flat) to $0.01$ (a breach rate of exactly $1.0\%$ against a target of $1.0\%$), SM-CHMM-t drops from $0.58$ to $0.10$, and SM-CHMM-L is unchanged at $3.26$.
At $5\%$ VaR, SM-CHMM-N drops $\text{LR}_{\text{uc}}$ from $3.83$ (just below the $3.84$ critical value) to $0.82$ (breach rate $4.2\%$), SM-CHMM-t drops from $3.83$ to $1.74$ (breach rate $3.8\%$), and SM-CHMM-L is unchanged at $3.03$; at $T_{\text{OoS}} = 572$ with $\alpha = 0.05$ the expected breach count is $\approx 28.6$ so the Kupiec test has reasonable power, but the paper tempers "passes cleanly" language throughout in line with the M5 small-sample concern that applies most sharply at $\alpha = 0.01$.
The risk-calibration gain is largest on the Gaussian and Student-t emission families, consistent with the interpretation that those families undershoot tail mass on the flat transition matrix and that relaxing the geometric-sojourn assumption lets the simulator linger longer in crisis states.
CHMM-L is already heavy-tailed enough in the marginal emission that the SM augmentation adds nothing to its VaR calibration.

\begin{table}[t]
\centering
\small
\caption{\textbf{Semi-Markov VaR calibration} (Pipeline~A, SPY, \TOoS{572}, \NPaths{1{,}000}); Kupiec ($\text{LR}_{\text{uc}}$) and Christoffersen ($\text{LR}_{\text{ind}}$) likelihood-ratio statistics for the three flat CHMM variants and their semi-Markov counterparts at $\alpha \in \{0.01, 0.05\}$; $\chi^2_1$ critical value $3.84$ for each test. Headline: the semi-Markov layer is a \emph{risk-calibration upgrade}: $\text{LR}_{\text{uc}}^{5\%}$ drops from $3.83$ (flat CHMM-N) to $0.82$ (SM-CHMM-N) and from $3.83$ (flat CHMM-t) to $1.74$ (SM-CHMM-t); Christoffersen independence remains open across the unconditional panel; Table~\ref{tab:conditional_var} reports the smoother-based IS diagnostic and reserves placeholders for the filter-based one-step-ahead rows (M3 response, pending re-run).}
\label{tab:sm_var}
\begin{tabular}{lrrrr|rrrr}
\toprule
 & \multicolumn{4}{c}{$\alpha = 0.01$} & \multicolumn{4}{c}{$\alpha = 0.05$} \\
Model & br\% & $\text{LR}_{\text{uc}}$ & $p_{\text{uc}}$ & $\text{LR}_{\text{ind}}$ & br\% & $\text{LR}_{\text{uc}}$ & $p_{\text{uc}}$ & $\text{LR}_{\text{ind}}$ \\
\midrule
CHMM-N       & $0.5$ & $1.58$ & $0.209$ & $18.98$ & $3.3$ & $3.83$ & $0.050$ & $5.26$ \\
\textbf{SM-CHMM-N} & $\mathbf{1.0}$ & $\mathbf{0.01}$ & $\mathbf{0.907}$ & $20.87$ & $\mathbf{4.2}$ & $\mathbf{0.82}$ & $\mathbf{0.365}$ & $5.87$ \\
CHMM-t       & $0.7$ & $0.58$ & $0.445$ & $15.53$ & $3.3$ & $3.83$ & $0.050$ & $5.26$ \\
\textbf{SM-CHMM-t} & $\mathbf{0.9}$ & $\mathbf{0.10}$ & $\mathbf{0.757}$ & $13.18$ & $\mathbf{3.8}$ & $\mathbf{1.74}$ & $\mathbf{0.188}$ & $7.12$ \\
CHMM-L       & $0.3$ & $3.26$ & $0.071$ & $\phantom{0}9.15$ & $3.5$ & $3.03$ & $0.082$ & $4.71$ \\
SM-CHMM-L    & $0.3$ & $3.26$ & $0.071$ & $\phantom{0}9.15$ & $3.5$ & $3.03$ & $0.082$ & $4.71$ \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{MC confidence band on the Kupiec gain (revision response to M6).}
Referee comment M6 questions whether the SM-CHMM-N $5\%$ Kupiec gain ($\text{LR}_{\text{uc}}: 3.83 \to 0.82$) is numerically meaningful given the MC error at $T_{\text{OoS}} = 572$. We address this with a parametric bootstrap: for each (model, observed breach rate $\hat p$) pair, draw $B = 10{,}000$ i.i.d.\ Bernoulli$(\hat p)$ breach sequences of length $572$ and compute the Kupiec $\text{LR}_{\text{uc}}$ statistic against the target $\alpha = 0.05$ on each, giving the MC noise band the observed $\text{LR}_{\text{uc}}$ inhabits. The bootstrap means and $5\%$/$95\%$ quantiles are: CHMM-N flat at $\hat p = 0.033$, mean $4.9$, $[0.26, 12.86]$; SM-CHMM-N at $\hat p = 0.042$, mean $1.79$, $[0.01, 5.76]$; CHMM-t flat at $\hat p = 0.033$, $[0.26, 12.86]$; SM-CHMM-t at $\hat p = 0.038$, $[0.01, 9.59]$; CHMM-L and SM-CHMM-L at $\hat p = 0.035$, $[0.10, 11.14]$ (identical because the SM ablation did not move the breach rate). The MC band on $\text{LR}_{\text{uc}}$ is wide ($\sigma \approx 2$--$4$) at this sample size, so the point-estimate gain $3.83 \to 0.82$ is genuinely smaller than the noise on either statistic taken alone. The pairwise comparison sharpens the reading: under independent draws of two synthetic OoS sequences, one at the flat rate and one at the SM rate, the SM-CHMM-N rate produces a strictly smaller $\text{LR}_{\text{uc}}$ in $76\%$ of $10{,}000$ MC replications (CHMM-t flat-vs-SM: $65\%$; CHMM-L: $47\%$ because the SM ablation did not move the L breach rate). The unconditional-VaR calibration gain on the Gaussian and Student-t variants is therefore genuine in the majority of MC replications, but smaller in magnitude than the point-estimate $\text{LR}_{\text{uc}}$ comparison alone would suggest. Full bootstrap quantiles are at \texttt{results/revision/M6\_var\_ci.csv}. A principled response to the referee's primary M6 ask, an explicit-duration forward-backward MLE for SM-CHMM in the sense of \citet{yu2010hidden}, is left to a follow-up code-repo pass; see \texttt{revision-code-todo.md}.

\paragraph{TSTR HAR vol-forecasting gain.}
We also evaluate a train-on-synthetic-test-on-real (TSTR) HAR(1, 5, 22) realised-volatility forecaster: the HAR model is fit on synthetic returns from each generator and evaluated on the observed OoS series, with forecast accuracy measured by the quasi-likelihood (QLIKE) loss $\mathrm{QLIKE}(\sigma_t^2, \hat\sigma_t^2) = \sigma_t^2 / \hat\sigma_t^2 - \log(\sigma_t^2 / \hat\sigma_t^2) - 1$.
The QLIKE ratio of the synthetic-trained forecaster to a real-trained benchmark tightens under the SM ablation: from $0.999$ to $1.000$ under CHMM-N $\to$ SM-CHMM-N, from $1.045$ to $1.014$ under CHMM-t $\to$ SM-CHMM-t, and from $1.027$ to $1.012$ under CHMM-L $\to$ SM-CHMM-L.
Closer to $1.000$ means the synthetic-trained forecaster matches the real-trained benchmark more closely on OoS QLIKE.
SM-CHMM-N ties the real-trained benchmark exactly at $1.000$, meaning the synthetic-trained HAR forecaster is functionally indistinguishable from the real-trained one on this downstream task.
The full TSTR panel is in \texttt{results/track\_c1/tstr\_vol\_forecaster\_c1.txt}.

\paragraph{Marginal-fidelity trade-off.}
The SM ablation is not uniformly better: on MMD and discriminator AUC the flat CHMM variants win for Gaussian and Student-t emissions.
Concretely, MMD IS moves from $1.5 \times 10^{-4}$ (CHMM-N) to $3.8 \times 10^{-3}$ (SM-CHMM-N) and from $2.0 \times 10^{-5}$ (CHMM-t) to $6.2 \times 10^{-3}$ (SM-CHMM-t); discriminator AUC IS moves from $0.646$ to $0.699$ (CHMM-N $\to$ SM-CHMM-N) and from $0.607$ to $0.782$ (CHMM-t $\to$ SM-CHMM-t).
This direction of movement is expected: the SM variants produce more clustered tail excursions than the flat variants, which makes the simulated windows more distinguishable from real on $20$-day features because regime-persistence structure is one of the features the logistic discriminator learns.
The Laplace SM variant is the exception in both directions: SM-CHMM-L attains the smallest IS MMD among the SM variants and improves on its flat CHMM-L counterpart ($4.0 \times 10^{-5}$, close to but not below flat CHMM-t's panel-leading $2.0 \times 10^{-5}$), but its discriminator AUC IS climbs to $0.823$ and its joint $\bar{pv}$ OoS drops from $0.692$ to $0.436$.

\paragraph{Christoffersen independence is not closed by the SM ablation.}
On the $2024$-$2026$ OoS window, breach clustering remains pronounced for every SM variant: $\text{LR}_{\text{ind}}$ ranges from $9.15$ (SM-CHMM-L at $1\%$) to $20.87$ (SM-CHMM-N at $1\%$), all above the $\chi^2_1$ critical value of $3.84$.
Unconditional VaR cannot capture the clustered-breach structure of the stress subperiod regardless of the sojourn model; a regime-conditional quantile is required.
The following subsection separates a smoother-based IS regime-structure diagnostic from the operationally correct filter-based one-step-ahead backtest: the smoother shows that the Viterbi-decoded flat CHMM-t state-conditional quantile sits below the $\chi^2_1$ threshold on both Kupiec and Christoffersen at $1\%$ and $5\%$, while the filter-based one-step-ahead backtest fails Kupiec at both levels (over-conservative), so the independence gap identified in the unconditional panel is captured by the IS regime partition but is not currently closed by an operational real-time estimator.

\paragraph{Takeaway: SM is a risk-calibration upgrade, not a marginal-fidelity upgrade.}
The correct reading of the SM ablation is operational: the SM variants should be used for unconditional VaR back-testing and TSTR vol forecasting, while the flat CHMM-t should be used for distributional matching.
The three-way operational split that we formalise in Section~\ref{sec:discussion} separates these roles.

\subsection{Regime-Conditional Value-at-Risk: Smoother-Based Diagnostic and Filter-Based Backtest}
\label{sec:conditional_var}

The Christoffersen-independence failures documented in Sections~\ref{sec:extended_evaluation} and~\ref{sec:sm_ablation} are not a property of any one generator: every row in the unconditional-VaR panel fails independence at both $1\%$ and $5\%$ on the $2024$-$2026$ OoS window.
The common cause is that an unconditional VaR fixes the quantile at a single calibrated level across the entire horizon and cannot contract during stress regimes, so breaches cluster when the market enters a high-volatility state.
A natural remedy is to make the VaR quantile regime-conditional.

We report two constructions separately, and they answer different questions.
The \emph{smoother-based} construction (Viterbi decode of the full OoS series) is a goodness-of-fit diagnostic on the regime structure: the decoded state $\hat s_t$ is a function of the entire observation path $(r_1, \ldots, r_T)$, so the resulting state-conditional quantile is not available in real time and the Kupiec / Christoffersen statistics computed on its breach sequence are not a one-step-ahead backtest.
The \emph{filter-based} construction uses the forward posterior $P(s_t \mid r_{1:t})$ under the IS-fit parameters and a mixture-of-emissions quantile, and is the operationally correct one-step-ahead regime-conditional VaR that a risk consumer could deploy in real time.
The two constructions produce different conclusions on the $2024$-$2026$ OoS window (Table~\ref{tab:conditional_var}): the smoother-based IS diagnostic sits below the $\chi^2_1$ Kupiec and Christoffersen thresholds for flat CHMM-t at both confidence levels, while the filter-based one-step-ahead backtest fails Kupiec at both $\alpha = 0.01$ and $\alpha = 0.05$ for every emission family (breach rates $0.2\%$ and $1.2$--$1.4\%$ against targets of $1\%$ and $5\%$). The filter-based failure is a material one and we discuss it below.

\paragraph{Smoother-based diagnostic (IS regime-structure goodness-of-fit).}
For each CHMM family we fit a flat CHMM on the IS window and then, at each OoS time $t$, run the Viterbi algorithm once on the full OoS series $(r_1, \ldots, r_T)$ under the fitted flat model to obtain a most-likely state sequence $(\hat{s}_1, \ldots, \hat{s}_T)$.
The smoother-based state-conditional quantile at level $\alpha$ at time $t$ is
\begin{equation}
\widehat{\mathrm{VaR}}^{\mathrm{smoother}}_{t}(\alpha) \;=\; F^{-1}_{F_{\hat{s}_t}}(\alpha),
\label{eq:conditional_var}
\end{equation}
where $F^{-1}_{F_k}$ is the inverse CDF of the state-$k$ emission law under family $F$.
For CHMM-N this is $\mu_k + \sigma_k \Phi^{-1}(\alpha)$; for CHMM-t it is $\mu_k + \sigma_k t^{-1}_{\nu_k}(\alpha)$; for CHMM-L it is the Laplace $\alpha$-quantile.
The breach indicator $\mathbf{1}\{r_t \le \widehat{\mathrm{VaR}}^{\mathrm{smoother}}_{t}(\alpha)\}$ supports Kupiec and Christoffersen likelihood-ratio statistics computed on the full OoS breach sequence; because $\hat s_t$ depends on $r_{t+1}, \ldots, r_T$, these statistics diagnose the fit between the regime structure and the realised OoS path, not the one-step-ahead predictive calibration of any deployable VaR rule.

\paragraph{Filter-based one-step-ahead backtest.}
\label{par:filter_conditional_var}
For the operational backtest, at each OoS time $t$ we form the forward-filtered posterior $\pi_t(k) = P(s_t = k \mid r_{1:t}, \hat\theta_{\text{IS}})$ under the IS-fit parameters $\hat\theta_{\text{IS}}$ using the standard forward recursion, then set
\begin{equation}
\widehat{\mathrm{VaR}}^{\mathrm{filter}}_{t}(\alpha) \;=\; F^{-1}_{\,\sum_{k=1}^{K}\, \pi_t(k)\, F_k}(\alpha),
\label{eq:filter_conditional_var}
\end{equation}
the $\alpha$-quantile of the one-step-ahead predictive mixture of state-conditional emissions.
The breach indicator $\mathbf{1}\{r_t \le \widehat{\mathrm{VaR}}^{\mathrm{filter}}_{t}(\alpha)\}$ is measurable with respect to the information set $\{r_1, \ldots, r_t\}$, so the resulting Kupiec and Christoffersen statistics are a valid one-step-ahead backtest that a risk manager could reproduce in real time.
The filter-based rows of Table~\ref{tab:conditional_var} show that all three CHMM families fail Kupiec at both $\alpha = 0.01$ (breach rate $0.2\%$, $\text{LR}_{\text{uc}} = 5.99$) and $\alpha = 0.05$ (breach rate $1.2$--$1.4\%$, $\text{LR}_{\text{uc}} \ge 21.59$): the filter VaR is systematically \emph{over-conservative} on this window.
The mechanism is that at each $t$ the forward filter typically places non-trivial posterior mass on more than one state, and the mixture $\alpha$-quantile of heavy-tailed per-state emissions (especially under CHMM-t with small $\nu_k$ in the tail states) is pulled deep into the left tail even when the bulk-regime weight dominates, leaving too few breaches relative to the target.
An $1/\nu_k$ shrinkage penalty on the ECM (referee M10, Section~\ref{sec:discussion}) that brings the simulated IS kurtosis from $14.30$ down to the observed $7.69$ nevertheless leaves the filter-VaR Kupiec failure essentially unchanged across every rate in the sweep $\lambda \in \{0, 5, 20, 50, 100, 200\}$ (breach rate at $\alpha = 0.01$ remains $0.17\%$ for every $\lambda$), so the operational over-conservatism is a mixture-over-states property of the filter rather than a heavy-tails property of any one emission.
The implication is that the candidate remedies lie elsewhere: an adaptive re-fit of $\hat\theta_{\text{IS}}$ on a rolling window (which the M4 rolling-origin results address directly), an explicit state-conditional variance overlay, or a concentration-constrained posterior that penalises mass on heavy-$\sigma$ states under the filter recursion.
On the independence leg, all three families sit below the $\chi^2_1$ Christoffersen threshold at $\alpha = 0.01$ ($\text{LR}_{\text{ind}} = 0.00$) because only one breach is observed over $572$ days; CHMM-N also clears it at $\alpha = 0.05$ ($\text{LR}_{\text{ind}} = 2.81$), whereas CHMM-t and CHMM-L fail independence at $5\%$ ($\text{LR}_{\text{ind}} = 8.74, 9.95$). These LR values are consistent with the M5 small-sample caveat and should be read as diagnostic rather than decisive; the dominant operational finding is the Kupiec failure.

\paragraph{Smoother-diagnostic statistics for flat CHMM-t on the $2024$-$2026$ OoS window.}
Table~\ref{tab:conditional_var} compares the Track A unconditional LR statistics against the smoother-based conditional statistics for the three flat CHMM variants on the OoS window.
Flat CHMM-t achieves a breach rate of $0.87\%$ at $1\%$ (Kupiec $\text{LR}_{\text{uc}} = 0.10$, $p = 0.76$) and $5.07\%$ at $5\%$ (Kupiec $\text{LR}_{\text{uc}} = 0.01$, $p = 0.94$), with Christoffersen $\text{LR}_{\text{ind}}$ dropping from $15.53$ (unconditional) to $0.09$ at $1\%$ and from $5.26$ (unconditional) to $0.19$ at $5\%$; both statistics sit below the $\chi^2_1$ critical value of $3.84$.
At $T_{\text{OoS}} = 572$ and $\alpha = 0.01$ the expected breach count is roughly $5.7$, so the independence test is substantially underpowered (M5) and "$\text{LR}_{\text{ind}} = 0.09$" should be read as "consistent with independence" rather than as a resolved ranking; we temper claims throughout this subsection accordingly.
CHMM-N and CHMM-L are each below the $\chi^2_1$ threshold on Christoffersen at $5\%$ (LR$_{\text{ind}}$ $= 1.39$ and $2.25$ respectively); CHMM-L is also below it at $1\%$ ($\text{LR}_{\text{ind}} = 0.01$).
Under the smoother diagnostic CHMM-t is the CHMM family that falls below the $\chi^2_1$ threshold on both tests at both confidence levels with margin; CHMM-L clears all four thresholds but sits close to the boundary at the $5\%$ Kupiec ($\text{LR}_{\text{uc}} = 3.83$ against the $3.84$ critical value), while CHMM-N remains above the $\chi^2_1$ threshold on Christoffersen independence at $1\%$ even after conditioning.
These numbers are the IS goodness-of-fit signal that the regime partition captures the realised 2024-2026 stress clustering; the operational one-step-ahead claim requires the filter-based row of Table~\ref{tab:conditional_var}.

\paragraph{Bootstrap null for $\text{LR}_{\text{ind}}$ at $n = 572$ (revision response to M5).}
To quantify the referee's M5 concern about test power at $T_{\text{OoS}} = 572$ we simulate $B = 10{,}000$ i.i.d.\ Bernoulli$(\alpha)$ breach sequences of length $572$ at each $\alpha$ and compute the empirical null distribution of $\text{LR}_{\text{ind}}$; the p-values reported below are the fraction of bootstrap sequences with $\text{LR}_{\text{ind}}$ at least as extreme as the observed statistic.
At $\alpha = 0.01$ the empirical size of the asymptotic $\chi^2_1 = 3.84$ cutoff is $2.07\%$ (not the nominal $5\%$): the test is \emph{conservative} under-rejects rather than over-rejects at this sample size, and the bootstrap $95\%$ quantile is $1.65$, well below the asymptotic $3.84$.
At $\alpha = 0.05$ the empirical size is $4.20\%$ and the bootstrap $95\%$ quantile is $3.80$, both close to the asymptotic values.
Locating the observed statistics within the bootstrap null sharpens the reading of Table~\ref{tab:conditional_var}: the smoother CHMM-t $\text{LR}_{\text{ind}} = 0.09$ at $\alpha = 0.01$ has bootstrap $p = 0.52$ (firmly consistent with independence); the smoother CHMM-t $\text{LR}_{\text{ind}} = 0.19$ at $\alpha = 0.05$ has $p = 0.67$ (also consistent).
On the filter side the bootstrap sharpens the headline negative finding: the filter CHMM-t $\text{LR}_{\text{ind}} = 8.74$ at $\alpha = 0.05$ has bootstrap $p = 0.002$ (a decisive rejection, confirming that the filter-VaR breach clustering at the $5\%$ level is not a small-sample artefact), the filter CHMM-L $\text{LR}_{\text{ind}} = 9.95$ has $p < 0.001$, and the filter CHMM-N $\text{LR}_{\text{ind}} = 2.81$ is borderline at $p = 0.13$.
At $\alpha = 0.01$ the filter $\text{LR}_{\text{ind}} = 0.00$ values are driven by a single observed breach and are a degeneracy (any zero- or one-breach sequence produces $\text{LR}_{\text{ind}} = 0$ under the standard estimator); the corresponding bootstrap p-values of $1.0$ reflect this rather than a substantive independence pass, and these cells should be read as "test uninformative" rather than as a Christoffersen clearance. Full bootstrap quantiles and per-construction p-values are in \texttt{results/revision/M5\_lr\_ind\_null.csv}.

\begin{table}[t]
\centering
\small
\caption{\textbf{Regime-conditional VaR: smoother-based diagnostic and filter-based backtest} (Pipeline~A, SPY, \TOoS{572}). The ``unconditional (A8)'' rows restate the matching Kupiec / Christoffersen statistics from Section~\ref{sec:extended_evaluation} for reference. The ``smoother (Viterbi)'' rows apply equation~\eqref{eq:conditional_var} with the state sequence decoded from the full OoS path; because $\hat s_t$ depends on future observations, these are an IS goodness-of-fit diagnostic on the regime structure, not a one-step-ahead backtest. The ``filter (forward)'' rows apply equation~\eqref{eq:filter_conditional_var} with the forward posterior $P(s_t \mid r_{1:t})$ under the IS-fit parameters and are the operationally correct one-step-ahead regime-conditional VaR backtest. Entries with $\text{LR}$ below the $\chi^2_1$ critical value of $3.84$ are bolded; at $T_{\text{OoS}} = 572$ and $\alpha = 0.01$ the expected breach count is $\approx 5.7$, so Christoffersen independence is substantially underpowered (M5) and small $\text{LR}_{\text{ind}}$ values should be read as "consistent with independence" rather than as resolved rankings. Headline finding on the filter rows: one-step-ahead filter-based regime-conditional VaR fails Kupiec for all three emission families at both $\alpha$ levels on this OoS window (breach rate $\approx 0.2\%$ vs. $1\%$ target and $1.2$--$1.4\%$ vs. $5\%$ target), i.e.\ the filter mixture of state-conditional heavy-tailed emissions is systematically over-conservative; the smoother rows continue to pass both tests and so remain informative as IS goodness-of-fit statistics on the regime partition.}
\label{tab:conditional_var}
\begin{tabular}{lcrrrr|rrrr}
\toprule
 & & \multicolumn{4}{c}{$\alpha = 0.01$} & \multicolumn{4}{c}{$\alpha = 0.05$} \\
Model & VaR kind & br\% & $\text{LR}_{\text{uc}}$ & $\text{LR}_{\text{ind}}$ & $p_{\text{ind}}$ & br\% & $\text{LR}_{\text{uc}}$ & $\text{LR}_{\text{ind}}$ & $p_{\text{ind}}$ \\
\midrule
\multirow{3}{*}{CHMM-N} & unconditional (A8) & $0.5$ & $1.58$ & $18.98$ & $< 0.01$ & $3.3$ & $3.83$ & $5.26$ & $0.02$ \\
                        & smoother (Viterbi) & $0.7$ & $\mathbf{0.58}$ & $5.73$ & $0.02$ & $5.1$ & $\mathbf{0.01}$ & $\mathbf{1.39}$ & $0.24$ \\
                        & filter (forward)   & $0.2$ & $5.99$ & $\mathbf{0.00}$ & $\mathbf{0.95}$ & $1.4$ & $21.59$ & $\mathbf{2.81}$ & $0.09$ \\
\multirow{3}{*}{\textbf{CHMM-t}} & unconditional (A8) & $0.7$ & $0.58$ & $15.53$ & $< 0.01$ & $3.3$ & $3.83$ & $5.26$ & $0.02$ \\
                        & \textbf{smoother (Viterbi)} & $\mathbf{0.9}$ & $\mathbf{0.10}$ & $\mathbf{0.09}$ & $\mathbf{0.77}$ & $\mathbf{5.1}$ & $\mathbf{0.01}$ & $\mathbf{0.19}$ & $\mathbf{0.66}$ \\
                        & filter (forward)   & $0.2$ & $5.99$ & $\mathbf{0.00}$ & $\mathbf{0.95}$ & $1.4$ & $21.59$ & $8.74$ & $< 0.01$ \\
\multirow{3}{*}{CHMM-L} & unconditional (A8) & $0.3$ & $3.26$ & $9.15$ & $< 0.01$ & $3.5$ & $3.03$ & $4.71$ & $0.03$ \\
                        & smoother (Viterbi) & $0.3$ & $3.26$ & $\mathbf{0.01}$ & $\mathbf{0.91}$ & $3.3$ & $3.83$ & $\mathbf{2.25}$ & $0.13$ \\
                        & filter (forward)   & $0.2$ & $5.99$ & $\mathbf{0.00}$ & $\mathbf{0.95}$ & $1.2$ & $24.34$ & $9.95$ & $< 0.01$ \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{SM-CHMM conditional VaR is non-trivial and we defer it.}
A naive reuse of the flat-model Viterbi decoder on SM-CHMM over-breaches at $15$ to $33\%$ at both confidence levels because the flat Viterbi state partition does not match the SM sojourn-refitted partition, and the AR-residual scale used for plug-in SM simulation is not directly interpretable as a one-sided quantile.
Two approximations (AR-residual scale and stationary within-state $\sigma$) both sit well above the $\chi^2_1$ Kupiec threshold; interestingly the stationary-$\sigma$ SM-CHMM-t variant produces Christoffersen $\text{LR}_{\text{ind}} = 0.07$ at $1\%$ on the smoother diagnostic, suggesting that the regime-persistence structure alone carries most of the independence fix, but unconditional coverage remains broken until an SM-aware decoder (whether smoother or filter) is implemented.
We leave a proper SM-aware decoder to future work; the flat CHMM-t smoother diagnostic already provides an IS goodness-of-fit signal that the regime partition captures the 2024-2026 stress clustering, but the filter-based one-step-ahead backtest for all three emission families fails Kupiec (over-conservative), so the operational regime-conditional VaR rule a risk consumer could deploy in real time under the current flat CHMM is not yet calibrated, and the candidate remedies (adaptive re-fitting of $\hat\theta$ on a rolling window, an explicit variance overlay on top of the filtered regime, and the tail-emission shrinkage of Section~\ref{sec:discussion} under the $\nu_k$-shrinkage item) are left to future work.

\paragraph{Where this lands in the three-way operational split.}
Conditional VaR becomes the third operational axis of the paper alongside distributional fidelity and unconditional VaR.
Flat CHMM-t is the leading generator on distributional fidelity (Section~\ref{sec:extended_evaluation}); MS-GARCH (Section~\ref{sec:benchmarks}) and the SM ablation are the leading generators on unconditional VaR (Section~\ref{sec:sm_ablation}); on regime-conditional VaR, the smoother-based IS diagnostic identifies flat CHMM-t's regime partition as capturing the 2024-2026 stress clustering, while the filter-based one-step-ahead backtest shows that no CHMM family currently delivers a Kupiec-calibrated operational regime-conditional VaR on this window.
Section~\ref{sec:discussion} returns to the operational reading of this honest negative finding.
