\documentclass[11pt,a4paper]{article}

% --- Packages ---
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage[numbers,sort&compress]{natbib}
\usepackage[margin=1in]{geometry}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{multirow}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{xcolor}
\usepackage{float}

% --- Hyperref setup ---
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    citecolor=blue,
    urlcolor=blue
}

% --- Custom commands ---
\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Prob}{\mathbb{P}}

\title{Continuous Hidden Markov Models for Equity Return Dynamics:\\Gaussian, Student-t, and Laplace Emissions Trained by EM, with Cross-Asset SIM and Copula Extensions}

\author{
    Abdulrahman Alswaidan\thanks{Cornell University, Robert Frederick Smith School of Chemical and Biomolecular Engineering, Ithaca, NY, USA. Email: \texttt{aa2725@cornell.edu}.}
    \and
    Cade Jin\thanks{Cornell University, Ithaca, NY, USA. Email: \texttt{cj383@cornell.edu}.}
    \and
    Jeffrey D.\ Varner\thanks{Cornell University, Robert Frederick Smith School of Chemical and Biomolecular Engineering, Ithaca, NY, USA. Email: \texttt{jdv27@cornell.edu}.}
}

\date{\today}

\begin{document}

\maketitle

% ========================================================================================= %
\begin{abstract}
We present a family of continuous hidden Markov models (CHMMs) for generating synthetic financial time series that preserve the canonical stylized facts of asset returns.
Three variants share log-space Expectation--Maximization (EM) training over a common quantile-initialized transition matrix and differ only in their per-state emission law: Gaussian (CHMM-N), Student-t with data-selected per-state degrees of freedom (CHMM-t), and Laplace (CHMM-L).
At a single moderate state resolution, the continuous models reproduce heavy tails, negligible linear autocorrelation, and persistent volatility clustering without any jump augmentation or return-space discretization.

On ten years of daily SPDR S\&P~500 exchange-traded fund (ETF) data with ticker SPY, we benchmark the three variants against a suite of non-parametric, classical-parametric, discrete hidden Markov model (HMM), and deep-generative alternatives under seven distributional and autocorrelation metrics, in both in-sample and out-of-sample windows.
The heavy-tailed variants close the residual kurtosis gap left by CHMM-N while retaining its autocorrelation and tail pass-rate fidelity, and a $1\%$ and $5\%$ Value-at-Risk (VaR) and Expected Shortfall (ES) back-test validates the improvement at both levels.
We extend the framework to multi-asset synthesis through a Single Index Model and Gaussian / Student-t copula generators; on six equities the copulas substantially outperform the single-index baseline on both per-asset distributional fidelity and off-diagonal correlation reproduction.

\bigskip\noindent
\textbf{Keywords:} Continuous Hidden Markov Model, Baum-Welch, Student-t Emissions, Stylized Facts, Volatility Clustering, Copula Dependence, Cross-Asset Simulation, Value-at-Risk.
\end{abstract}

% ========================================================================================= %
\section{Introduction}
\label{sec:introduction}
\input{sections/introduction_v8}

\section{Related Work}
\label{sec:related}
\input{sections/related_v8}

\section{Methods}
\label{sec:methods}
\input{sections/methods_v8}

\section{Empirical Study}
\label{sec:results}
\input{sections/results_v8}

\section{Discussion}
\label{sec:discussion}
\input{sections/discussion_v8}

\section{Conclusion}
\label{sec:conclusion}
\input{sections/conclusion_v8}

% ========================================================================================= %
\bibliographystyle{unsrtnat}
\bibliography{References_v8}

% ========================================================================================= %
\appendix
\section{Supplementary Material}
\label{sec:supplemental}
\input{sections/supplemental_v8}

\end{document}
